---
title: "barisolidact"
author: "A.Amstutz"
date: "2023-07-29"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)
library(haven) # Read sas files

```

# Load Data
```{r, include=FALSE}
df <- read_sas("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/Bari-Solidact/bari_ipdma.sas7bdat")

```

# Define ITT set
```{r}
# 9 participants (3 in int. / 6 in cont.) did not receive a single dose of baricitinib, or placebo, respectively.
# Take them out, as done in results publication (mITT set)
df <- df %>% 
  filter(imp_adm_yn == 1)

```

# Baseline Characteristics
```{r echo=TRUE}

df <- df %>% ## no missing data
  rename(id_pat = PARTICIPANT_ID,
         trt = arm,
         age = D1_AGE,
         randdate = RANDODATED1,
         sex = GENDER,
         country = COUNTRY,
         icu = ICU_YN)

# Days with symptoms prior to randomization
df <- df %>% ## table(df$sympdur, useNA = "always") / ## no missing data / equals delay_symprando
  mutate(sympdur = randdate - SCR_SYMPSTDATED1,
         sympdur = as.numeric(sympdur))
ggplot(df, aes(x = sympdur)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "Symptom Duration",
       y = "Density")

# Severity of COVID-19 with respect to respiratory support at randomisation / Bari-Solidact used WHO score, transform
df <- df %>% ## no missing data // publication: 139 bari vs 136 placebo. In this dataset, slightly more (142 bari vs 138 placebo), due to additional randomized participants during extension of the trial
  mutate(clinstatus_baseline = case_when(whoscore_D1 == 0 | whoscore_D1 == 1 | whoscore_D1 == 3 ~ 1,
                                         whoscore_D1 == 4 ~ 2,
                                         whoscore_D1 == 5 ~ 3,
                                         whoscore_D1 == 6 ~ 4,
                                         whoscore_D1 == 7 | whoscore_D1 == 8 | whoscore_D1 == 9 ~ 5,
                                         whoscore_D1 == 10 ~ 6))
df$clinstatus_baseline <- factor(df$clinstatus_baseline, levels = 1:6) ## no missing data
addmargins(table(df$clinstatus_baseline, df$trt, useNA = "always"))

# Co-medication at baseline
df <- df %>% 
  mutate(comed_dexa = case_when(corticoid_yn == 1 | dexa_yn == 1 ~ 1,
                                corticoid_yn == 0 & dexa_yn == 0 ~ 0))
df <- df %>% 
  rename(comed_rdv = remdes_yn,
         comed_toci = toci_yn,
         comed_ab = antibio_yn,
         comed_acoa = anticoag_yn)
df$comed_interferon <- NA ## no interferon used / but monoclonal Abs and plasma -> other
df <- df %>%
  mutate(comed_other = case_when(monocloAb_yn == 1 | immunplasma_yn == 1 ~ 1,
                                monocloAb_yn == 0 & immunplasma_yn == 0 ~ 0))

## group them for the subgroup analysis, according to protocol // tocilizumab as rescue medication - incorporate?
df <- df %>% 
  mutate(comed_cat = case_when(comed_dexa == 0 & comed_toci == 0 ~ 1, # patients without Dexamethasone nor Tocilizumab
                               comed_dexa == 1 & comed_toci == 1 ~ 2, # patients with Dexamethasone and Tocilizumab
                               comed_dexa == 1 & comed_toci == 0 ~ 3, # patients with Dexamethasone but no Tocilizumab
                               comed_dexa == 0 & comed_toci == 1 ~ 4)) # patients with Tocilizumab but no Dexamethasone (if exist)

# Comorbidity at baseline, including immunocompromised
df <- df %>% ## 4 missing
  mutate(comorb_lung = case_when(D1_PULMO_YNK == 1 ~ 1,
                                 D1_PULMO_YNK == 0 ~ 0))
df <- df %>% 
  mutate(comorb_liver = case_when(D1_LIVER_YNK == 1 ~ 1,
                                D1_LIVER_YNK == 0 ~ 0))
df <- df %>% ## no missing. D1_CARDIO_YNK includes all CVDs but also 2 HTA that do not have anything else -> take out.
  mutate(comorb_cvd = case_when(D1_CARDIO_YNK == 1 & (D1_CARDIO_SP != "HTA" & D1_CARDIO_SP != "HIGH BLOOD PRESSURE") ~ 1,
                                 D1_CARDIO_YNK == 0 | D1_CARDIO_SP == "HTA" | D1_CARDIO_SP == "HIGH BLOOD PRESSURE" ~ 0))
df <- df %>% ## 1 missing. The 2 HTA from above were already counted in.
  mutate(comorb_aht = case_when(D1_HBP_YNK == 1 | D1_CARDIO_SP == "HTA" | D1_CARDIO_SP == "HIGH BLOOD PRESSURE" ~ 1,
                                D1_HBP_YNK == 0 ~ 0))
df <- df %>% ## 4 missing. D1_DIABETE not needed, correct.
  mutate(comorb_dm = case_when(d1_diabete_ynk == 1 ~ 1,
                                d1_diabete_ynk == 0 ~ 0))
df <- df %>% ## 0 missing. BMI not needed, correct.
  mutate(comorb_obese = case_when(obesity == 1 ~ 1,
                                obesity == 0 ~ 0))
df <- df %>% ## 0 missing. D1_SMOKING not needed, correct.
  mutate(comorb_smoker = case_when(smoker == 1 ~ 1,
                                smoker == 0 ~ 0))
df <- df %>% ## 0 missing.
  mutate(immunosupp = case_when(immunodef_yn == 1 ~ 1,
                                immunodef_yn == 0 ~ 0))
df <- df %>% 
  mutate(any_comorb = case_when(comorb_lung == 1 | comorb_liver == 1 | comorb_cvd == 1 |
                                  comorb_aht == 1 | comorb_dm == 1 | comorb_obese == 1 | comorb_smoker == 1
                                | immunosupp == 1
                                  ~ 1,
                                comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0
                                ~ 0))

## group them for the subgroup analysis, according to protocol // count all pre-defined comorbidities per patient first
comorb <- df %>% 
  select(id_pat, comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp)
comorb$comorb_count <- NA
for (i in 1:dim(comorb)[[1]]) {
  comorb$comorb_count[i] <- ifelse(
    sum(comorb[i, ] %in% c(1)) > 0,
    sum(comorb[i, ] %in% c(1)),
    NA
  )
}
comorb <- comorb %>% 
  mutate(comorb_count = case_when(comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 ~ 0,
                                TRUE ~ comorb_count))
df <- left_join(df, comorb[, c("comorb_count", "id_pat")], by = join_by(id_pat == id_pat)) ## merge imputed variable back
df <- df %>% 
  mutate(comorb_cat = case_when(immunosupp == 1 ~ 4, # immunocompromised
                                comorb_count == 0 ~ 1, # no comorbidity
                                comorb_count == 1 ~ 2, # one comorbidity
                                comorb_count >1 & immunosupp == 0 ~ 3)) # multiple comorbidities
# table(df$comorb_cat, useNA = "always") # 10 missing, have just 1 of the comorb categories missing...re-classify or multiple imputation?

# CRP
df$crp <- as.numeric(df$lab_v_conv_crp_D1) ## 7 missing

df %>% 
  drop_na(crp) %>% 
  ggplot(aes(x = crp)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of CRP",
       x = "CRP",
       y = "Density")

# Vaccination
df <- df %>% ## 4 missing
  mutate(vacc = case_when(D1_VACCIN_YNK == 1 ~ 1,
                          D1_VACCIN_YNK == 0 ~ 0))
# Viremia // Viral load value <LOQ and/or undectectable
df$vl_baseline <- df$vloqundet_yn_D1

# Variant
df <- df %>% 
  mutate(variant = case_when(SCR_COVARIANT == 1 ~ "Delta",
                             SCR_COVARIANT == 2 ~ "Omicron",
                             SCR_COVARIANT == 50 ~ "PRESENCE OF MUTATIONS E484Q AND L452R"))

# Serology
df$sero <- df$Serostatus
# df %>%
#    select(id_pat, Serostatus, Anti_Spike_wt, Anti_Nucl_wt, anti_RBDwt, Anti_RBD_Omicron) %>%
#   View()

```
Discussion points BASELINE data:
1) Ethnicity: Only country of birth available => missing for Bari-Solidact
2) Rescue therapy: Tocilizumab (n=12) or increased steroid dose (n=91). steroids_dosechang_yn, steroids_dosechang_date, rescue_yn, rescuedate -> Discuss for sens-analysis
3) Re-discuss the comorbidity list (e.g. autoimmune diseases, cancer, chronic kidney disease // smokers separate?)
4) Missing data for comorbidity subgroup analyses: MICE?
5) No-one received any monoclonal Abs? No-one received any plasma? Any interferons reported?
6) Serology (Serostatus): anti-spike, anti-RBD, anti-N, anti-RBD-Omicron

# Endpoints
```{r echo=TRUE}
# time to event data
df$death_d <- as.numeric(df$deathdated1_DROP - df$randdate) 
df$discharge_d <- as.numeric(df$visitdated1_DISCH - df$randdate)
df$withdraw_d <- as.numeric(df$withdrawdated1_DROP - df$randdate)
df$withdrawi_d <- as.numeric(df$invdecisdated1_DROP - df$randdate)
# df$ltfu_d <- as.numeric(df$ltfudated1_DROP - df$randdate) ## there were no LTFU
df$readmission_d <- as.numeric(df$readmdate - df$randdate)
df$maxfup_d <- as.numeric(df$lastdate - df$randdate)

# transform all daily clinical scores
whoscore_transform <- function(df, clinstatus_var, whoscore_var) {
  df <- df %>%
    mutate({{ clinstatus_var }} :=
             case_when({{ whoscore_var }} %in% c(0, 1, 3) ~ 1,
                       {{ whoscore_var }} == 4 ~ 2,
                       {{ whoscore_var }} == 5 ~ 3,
                       {{ whoscore_var }} == 6 ~ 4,
                       {{ whoscore_var }} %in% c(7, 8, 9) ~ 5,
                       {{ whoscore_var }} == 10 ~ 6)) %>%
    mutate({{ clinstatus_var }} := factor({{ clinstatus_var }}, levels = 1:6))
}
df <- whoscore_transform(df, clinstatus_2, whoscore_D3) ## D1 was baseline => D3 is +2 days after baseline.
df <- whoscore_transform(df, clinstatus_4, whoscore_D5)
df <- whoscore_transform(df, clinstatus_7, whoscore_D8)
df <- whoscore_transform(df, clinstatus_14, whoscore_D15)
df <- whoscore_transform(df, clinstatus_21, whoscore_D22)
df <- whoscore_transform(df, clinstatus_28, whoscore_D29) 
df <- whoscore_transform(df, clinstatus_35, whoscore_D36)
df <- whoscore_transform(df, clinstatus_discharge, whoscore_DISCH)
df <- whoscore_transform(df, clinstatus_dropout, whoscore_DROP)

# (i) Primary outcome: Mortality at day 28
df <- df %>% # 5 have no outcome data and withdrew or were withdrawn -> multiple imputation
  mutate(mort_28 = case_when(death_d <29 ~ 1,
                             discharge_d <29 ~ 0, # all discharged were discharged alive and not to hospiz
                             clinstatus_28 %in% c(2,3,4,5) ~ 0, # still at hospital but alive
                             discharge_d >28 ~ 0)) # discharged later, proof of still alive
        
# (ii) Mortality at day 60
df <- df %>% # same 5 that have no outcome data and withdrew or were withdrawn -> multiple imputation
  mutate(mort_60 = case_when(death_d <61 ~ 1,
                             discharge_d <61 ~ 0, # all discharged were discharged alive and not to hospiz
                             clinstatus_35 %in% c(2,3,4,5) ~ 0, # still at hospital but alive
                             discharge_d >60 ~ 0)) # discharged later, proof of still alive
        
# (iii) Time to death within max. follow-up time
df$death_reached <- df$death_yn
df <- df %>% # 2 are left without any time to event data => impute max. follow-up time
  mutate(death_time = case_when(death_d >=0 ~ c(death_d), # time to death, if no time to death, then...
                                discharge_d >=0 ~ c(discharge_d), # time to discharge, then...
                                withdraw_d >=0 ~ c(withdraw_d), # time to withdrawal, then...
                                withdrawi_d >=0 ~ c(withdrawi_d), # time to i_withdrawal, then...
                                readmission_d >=0 ~ c(readmission_d), # time to readmission, then...
                                maxfup_d >=0 ~ c(maxfup_d))) # time to max fup

# (iv) New mechanical ventilation among survivors within 28 days. Bari-Solidact only included clinstatus 4 and 5.
df <- df %>% # 4 NA are due to missing mortality data -> multiple imputation. The other NA are not eligible (died or clinstatus_baseline == 5) and thus are excluded from denominator -> no multiple imputation
  mutate(new_mv_28 = case_when(clinstatus_baseline == 4 & (mort_28 == 0 | is.na(mort_28)) 
                               & (clinstatus_2 == 5 | clinstatus_4 == 5 | clinstatus_7 == 5 | clinstatus_14 == 5 
                                  | clinstatus_21 == 5 | clinstatus_28 == 5)
                               ~ 1,
                               clinstatus_baseline == 4 & mort_28 == 0
                               ~ 0))

# (iv) Sens-analysis: Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator.
df <- df %>% # 4 NA are due to missing mortality data -> multiple imputation.
  mutate(new_mvd_28 = case_when(new_mv_28 == 1 | mort_28 == 1 ~ 1,
                                new_mv_28 == 0 | mort_28 == 0 ~ 0))

# (v) Clinical status at day 28
df <- df %>% # Adapt clinstatus_28, since currently excluding those discharged or died or missing data.
  mutate(clinstatus_28 = case_when(clinstatus_28 == 5 ~ 5,
                                   clinstatus_28 == 4 ~ 4,
                                   clinstatus_28 == 3 ~ 3,
                                   clinstatus_28 == 2 ~ 2,
                                   mort_28 == 1 ~ 6, # died within 28d
                                   mort_28 == 0 ~ 1)) # discharged alive / reached discharge criteria within 28d
df$clinstatus_28 <- factor(df$clinstatus_28, levels = 1:6) ## no missing data
## Imputation according to protocol: If there was daily data for the ordinal score available but with missing data for single days, then we carried last observed value forward unless for day 28, whereby we first considered data from the window (+/-3 days). -> no window data in Bari-Solidact => LVCF
dfcs <- df %>% 
    select(id_pat, clinstatus_baseline, clinstatus_2, clinstatus_4, clinstatus_7, clinstatus_14, clinstatus_21, clinstatus_28)
impute_last_forw = function(df){
  first = which(names(df)%in%c("clinstatus_baseline"))
  last = which(names(df)%in%c("clinstatus_28"))
  for (i in 1:dim(df)[[1]]){
    for (j in first[1]:last[1]){
      p = df[i, j]
      df[i,j] <- 
        ifelse(!is.na(df[i, j]), p, df[i, j-1])
    }
  }
  df
}
dfcs <- impute_last_forw(dfcs)
dfcs <- dfcs %>% # To control, don't overwrite
  rename(clinstatus_28_imp = clinstatus_28)
df <- left_join(df, dfcs[, c("clinstatus_28_imp", "id_pat")], by = join_by(id_pat == id_pat)) # # Merge imputed variable back // re-admissions already correctly incorporated

# (vi) Time to discharge or reaching discharge criteria up to day 28 // Patients who died prior to day 28 are assumed not having reached discharge, i.e. counted as 28 days. 
df <- df %>% 
  mutate(discharge_reached = case_when(discharge_d <29 ~ 1,
                                       TRUE ~ 0))
df <- df %>% # 2 are left without any time to event data => impute max. follow-up time
  mutate(discharge_time = case_when(discharge_d >=0 ~ c(discharge_d), # time to discharge, if no time to discharge, then...
                                    death_d >=0 ~ c(death_d), # time to death, then...
                                    withdraw_d >=0 ~ c(withdraw_d), # time to withdrawal, then...
                                    withdrawi_d >=0 ~ c(withdrawi_d), # time to i_withdrawal, then...
                                    # readmission_d >=0 ~ c(readmission_d), # time to readmission, then...
                                    maxfup_d >=0 ~ c(maxfup_d))) # time to max fup
df <- df %>% # add 28d for those that died
  mutate(discharge_time = case_when(mort_28 == 1 ~ 28,
                                    TRUE ~ discharge_time))
df <- df %>% # restrict to max fup time 28d
  mutate(discharge_time = case_when(discharge_time >28 ~ 28,
                                    TRUE ~ discharge_time))

# (vi) Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days
df <- df %>% # there are 5 re-admissions. the one readmitted at day 16 was re-discharged at day ?28?. The ones at d21 & d27 not. And then there are 2 that were readmitted later than d28. => Reclassify d21 and d27, add d28 to the one re-admitted at d16.
  mutate(discharge_reached_sus = case_when(readmission_d == 21 | readmission_d == 27 ~ 0,
                                           TRUE ~ discharge_reached))
df <- df %>%
  mutate(discharge_time_sus = case_when(readmission_d == 16 ~ 16,
                                        readmission_d == 21 | readmission_d == 27 ~ 28,
                                        TRUE ~ discharge_time))

# (vii) Viral clearance up to day 5, day 10, and day 15 (Viral load value <LOQ and/or undectectable)
df$vir_clear_5 <- df$vloqundet_yn_D3
df$vir_clear_10 <- df$vloqundet_yn_D8 # point prevalence, not cumulative
df$vir_clear_15 <- df$vloqundet_yn_D15 # point prevalence, not cumulative
df <- df %>% 
  mutate(vir_clear_15_cum = case_when(vir_clear_15 == 1 | vir_clear_10 == 1 | vir_clear_5 == 1 ~ 1,
                                      vir_clear_15 == 0 ~ 0,
                                      vir_clear_10 == 0 ~ 0,
                                      vir_clear_5 == 0 ~ 0))

# (viii) Quality of life at day 28 // detailed measures available - wait for other trials first

# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28

# (ix) Sens-analysis: Alternative definition/analysis of outcome: time to first (of these) adverse event, within 28 days, considering death as a competing risk (=> censor and set to 28 days)

# (x) Adverse events of special interest within 28 days: a) thromboembolic events (venous thromboembolism, pulmonary embolism, arterial thrombosis), b) secondary infections (bacterial pneumonia including ventilator-associated pneumonia, meningitis and encephalitis, endocarditis and bacteremia, invasive fungal infection including pulmonary aspergillosis), c) Reactivation of chronic infection including tuberculosis, herpes simplex, cytomegalovirus, herpes zoster and hepatitis B, d) serious cardiovascular and cardiac events (including stroke and myocardial infarction), e) events related to signs of bone marrow suppression (anemia, lymphocytopenia, thrombocytopenia, pancytopenia), f) malignancy, g) gastrointestinal perforation (incl. gastrointestinal bleeding/diverticulitis), h) liver dysfunction/hepatotoxicity (grade 3 and 4)

# (xi) Adverse events, any grade and serious adverse event, excluding death, within 28 days, grouped by organ classes


```
Discussion points OUTCOME data:
1) VL: Add a cumulative outcome definition (as sens-analysis)?
2) Rediscuss and double-check the 5 re-admissions.
3) Re QoL: Wait for other trials first. Find out more about the QoL measure used.
4) Get the (S)AE data

# Multiple imputation using chained equation
```{r}

```

