---
title: "Ghazaeian"
author: "A.Amstutz"
date: "2023-09-09"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)
library(here)
library(kableExtra)

library(jtools) # for summ() and plot_summs
library(sjPlot) # for tab_model
library(ggplot2) # survival/TTE analyses and other graphs
library(ggsurvfit) # survival/TTE analyses
library(survival) # survival/TTE analyses
library(gtsummary) # survival/TTE analyses
library(ggfortify) # autoplot
library(tidycmprsk) # competing risk analysis
library(ordinal) # clinstatus ordinal regression
library(mosaic) # OR for 0.5-corrected 2x2 table in case of rare events

```

# Load Data
```{r, include=FALSE}
df <- read_xlsx("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/Ghazaeian - Irct20190804044429N 2021/COVID-19 JAKi.xlsx")

```

# Baseline Characteristics
```{r echo=TRUE}
df$trial <- c("Ghazaeian")
df$country <- c("Iran")
df$ethn <- c("Persian/Mazani")
df <- df %>% # no missing in id_pat, trt, age, sex
  rename(id_pat = "patient ID",
         trt = "Treatment group",
         age = Age,
         sex = Sex)
df$id_pat = as.character(paste0('ghazaeian-', df$id_pat))
df %>% 
  drop_na(age) %>% 
  ggplot(aes(x = age)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Age",
       x = "Age",
       y = "Density")

df$icu <- 0 # no icu patients at enrolment

# Days with symptoms prior to randomization
df %>% # no missings
  drop_na(sympdur) %>% 
  ggplot(aes(x = sympdur)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "Symptom Duration",
       y = "Density")

# Severity of COVID-19 with respect to respiratory support at randomisation / from publication: "Scores on the ordinal scale are as follows: 1, not hospitalized, no limitations of activities; 2, not hospitalized, limitation of activities, home oxygen requirement, or both; 3, hospitalized, not requiring supplemental oxygen and no longer requiring ongoing medical care (used if hospitalization was extended for infection-control reasons); 4, hospitalized, not requiring supplemental oxygen but requiring ongoing medical care (Covid-19–related or other medical conditions); 5, hospitalized, requiring any supplemental oxygen; 6, hospitalized, receiving noninvasive ventilation or use of high-flow oxygen devices; 7, hospitalized, receiving invasive mechanical ventilation or ECMO; and 8, death. Same as ACTT2.

# transform clinical score
score_transform <- function(df, clinstatus_var, score_var) {
  df <- df %>%
    mutate({{ clinstatus_var }} :=
             case_when({{ score_var }} %in% c(1, 2, 3) ~ 1,
                       {{ score_var }} == 4 ~ 2,
                       {{ score_var }} == 5 ~ 3,
                       {{ score_var }} == 6 ~ 4,
                       {{ score_var }} == 7 ~ 5,
                       {{ score_var }} == 8 ~ 6)) %>%
    mutate({{ clinstatus_var }} := factor({{ clinstatus_var }}, levels = 1:6))
}
df <- score_transform(df, clinstatus_baseline, clinstatus_baseline) 
addmargins(table(df$clinstatus_baseline, df$trt, useNA = "always")) # no missing
df <- score_transform(df, clinstatus_7, df$clinstatus_7)
df <- score_transform(df, clinstatus_14, df$clinstatus_14)
df <- score_transform(df, clinstatus_28, df$`clinstatus-28`)

# Co-medication at baseline
class(df$"Comed-Dexa")
df <- df %>% 
  mutate(comed_dexa = case_when(`Comed-Dexa` == "yes" ~ 1,
                              TRUE ~ 0))
df <- df %>% 
  mutate(comed_rdv = case_when(`COMed-RDV` == "yes" ~ 1,
                              TRUE ~ 0))
df <- df %>% 
  mutate(comed_ab = case_when(`Comed-ab` == "yes" ~ 1,
                              TRUE ~ 0))
df <- df %>% 
  mutate(comed_acoa = case_when(`Comed-anticoagulant` == "yes" ~ 1,
                              TRUE ~ 0))
df <- df %>% 
  mutate(comed_interferon = case_when(`Comed-interferon` == "no" ~ 0,
                              TRUE ~ 1))
df$comed_toci <- 0 # but 4 received Actemra (Tocilizumab) 2-4 days after randomization

df <- df %>% # comed_other notes all as no - but there is a separate statin variable..
  mutate(comed_other = case_when(Statin == "no" ~ 0,
                                 Statin == "yes" ~ 1))

## group them for the subgroup analysis, according to protocol
df <- df %>% 
  mutate(comed_cat = case_when(comed_dexa == 0 & comed_toci == 0 ~ 1, # patients without Dexamethasone nor Tocilizumab
                               comed_dexa == 1 & comed_toci == 1 ~ 2, # patients with Dexamethasone and Tocilizumab
                               comed_dexa == 1 & comed_toci == 0 ~ 3, # patients with Dexamethasone but no Tocilizumab
                               comed_dexa == 0 & comed_toci == 1 ~ 4)) # patients with Tocilizumab but no Dexamethasone (if exist)

# Comorbidity at baseline, including immunocompromised
df <- df %>% 
  mutate(comorb_lung = case_when(Asthma == "yes" ~ 1,
                                 Asthma == "no" ~ 0))
df$comorb_liver <- NA
df <- df %>% 
  mutate(comorb_cvd = case_when(`IHD (ischemic heart disease)` == "yes" | Stroke == "yes" ~ 1,
                                 `IHD (ischemic heart disease)` == "no" & Stroke == "no" ~ 0))
df <- df %>% 
  mutate(comorb_aht = case_when(HTN == "yes" ~ 1,
                                 HTN == "no" ~ 0))
df <- df %>% 
  mutate(comorb_dm = case_when(`DM (diabetes melitus)` == "yes" ~ 1,
                               `DM (diabetes melitus)` == "no" ~ 0))
df <- df %>% 
  mutate(bmi = Weight/((Height/100)*(Height/100)))
df <- df %>% 
  mutate(comorb_obese = case_when(bmi > 30 ~ 1,
                                  bmi <= 30 ~ 0))
df <- df %>% 
  mutate(comorb_smoker = case_when(Smoker == "yes" ~ 1,
                                    Smoker == "no" ~ 0))
df <- df %>% 
  mutate(immunosupp = case_when(Malignancy == "yes" ~ 1,
                                    Malignancy == "no" ~ 0))
df <- df %>% 
  mutate(comorb_cancer = case_when(Malignancy == "yes" ~ 1,
                                    Malignancy == "no" ~ 0))
df <- df %>% 
  mutate(comorb_autoimm = case_when(`IBD (irritable bowel disease)` == "yes" | `Thyroid dis` == "yes" ~ 1,
                                    `IBD (irritable bowel disease)` == "no" & `Thyroid dis` == "no" ~ 0))
df$comorb_kidney <- NA

df <- df %>% # 18 missing - as in publication
  mutate(any_comorb = case_when(comorb_lung == 1 | comorb_liver == 1 | comorb_cvd == 1 |
                                  comorb_aht == 1 | comorb_dm == 1 | comorb_obese == 1 | comorb_smoker == 1
                                | immunosupp == 1 | comorb_cancer == 1 | comorb_autoimm == 1 | comorb_kidney == 1 
                                  ~ 1,
                                comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 & comorb_cancer == 0 & comorb_autoimm == 0 & comorb_kidney == 0
                                ~ 0))
# the remaining 41 missing have only NA in 1 comorb category => no evidence for comorbidity -> recode as 0
df <- df %>% 
  mutate(any_comorb = case_when(is.na(any_comorb) ~ 0,
                                TRUE ~ any_comorb))
# df %>%
#   select(any_comorb, comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp, comorb_kidney, comorb_autoimm, comorb_cancer) %>%
#   filter(is.na(any_comorb)) %>%
#   View()

## group them for the subgroup analysis, according to protocol // count all pre-defined comorbidities per patient first
comorb <- df %>% 
  select(id_pat, comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp, comorb_kidney, comorb_autoimm, comorb_cancer)
comorb$comorb_count <- NA
for (i in 1:dim(comorb)[[1]]) {
  comorb$comorb_count[i] <- ifelse(
    sum(comorb[i, ] %in% c(1)) > 0,
    sum(comorb[i, ] %in% c(1)),
    NA
  )
}
comorb <- comorb %>% 
  mutate(comorb_count = case_when(comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 & comorb_cancer == 0 & comorb_autoimm == 0 & comorb_kidney == 0 ~ 0,
                                TRUE ~ comorb_count))
# the remaining 11 missing have only NA in 1 comorb category => no evidence for comorbidity -> recode as 0
comorb <- comorb %>% 
  mutate(comorb_count = case_when(is.na(comorb_count) ~ 0,
                                TRUE ~ comorb_count))
df <- left_join(df, comorb[, c("comorb_count", "id_pat")], by = join_by(id_pat == id_pat)) ## merge imputed variable back
df <- df %>% # 
  mutate(comorb_cat = case_when(immunosupp == 1 ~ 4, # immunocompromised
                                comorb_count == 0 ~ 1, # no comorbidity
                                comorb_count == 1 ~ 2, # one comorbidity
                                comorb_count >1 & (immunosupp == 0 | is.na(immunosupp)) ~ 3)) # multiple comorbidities

# CRP
df$crp <- df$CRP ## 1 missing
df <- df %>% # 0 does not make sense
  mutate(crp = na_if(crp, 0))
df %>% 
  drop_na(crp) %>% 
  ggplot(aes(x = crp)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of CRP",
       x = "CRP",
       y = "Density")

# Vaccination
# Viremia
# Variant
# Serology
```
Clarifications and discussion points BASELINE data:
1) 4 received Actemra (Tocilizumab) 2-4 days after randomization


# Endpoints
```{r echo=TRUE}
# time to event data
df$randdate <- as_date(df$`Randomization  Date`)
df$death_date <- as_date(df$`death date`)
df$withdrawal_date <- as_date(df$`withdrawal_date`)

df$death_d <- as.numeric(df$death_date - df$randdate) 
df$withdraw_d <- as.numeric(df$withdrawal_date - df$randdate)
df$discharge_d <- as.numeric(df$`Hospital days`)

# (i) Primary outcome: Mortality at day 28 // CAVE: WE HAVE WITHDRAWALS!!!! at the same time variable "death"...
df <- df %>%
  mutate(mort_28 = case_when(death_d <29 ~ 1, 
                             TRUE ~ 0)) 

# (ii) Mortality at day 60
df$mort_60 <- df$mort_28 # max fup time was 28 days; thus mort_60 imputed from mort_28


# (iii) Time to death within max. follow-up time
df$death_reached <- df$mort_28
df <- df %>% # no missing and those that were discharged were discharged alive
  mutate(death_time = case_when(death_d >=0 ~ c(death_d), # time to death, if no time to death, then...
                                withdraw_d >=0 ~ c(withdraw_d), # time to withdrawal, then...
                                TRUE ~ 28)) # time to death censoring data // All happened within 14 days. And without change until 28d follow-up (those discharged were still alive and not hospitalized)


# (iv) New mechanical ventilation among survivors within 28 days.
df <- df %>% # all those intubated also then died
  mutate(new_mv_28 = case_when(mort_28 == 0 & (clinstatus_7 == 5 | clinstatus_14 == 5 | clinstatus_28 == 5) ~ 1,
                               mort_28 == 0 ~ 0))

# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
df <- df %>% # no missing anymore => correct
  mutate(new_mvd_28 = case_when(new_mv_28 == 1 | mort_28 == 1 ~ 1,
                                new_mv_28 == 0 | mort_28 == 0 ~ 0))


# (v) Clinical status at day 28
df <- df %>%
  mutate(clinstatus_28 = case_when(mort_28 == 1 ~ 6,
                                   discharge_d <29 ~ 1)) 
df$clinstatus_28 <- factor(df$clinstatus_28, levels = 1:6) # no missing
df$clinstatus_28_imp <- df$clinstatus_28 # equal to clinstatus_28_imp


# (vi) Time to discharge or reaching discharge criteria up to day 28
df <- df %>%
  mutate(discharge_reached = case_when(discharge_d <29 & mort_28 == 0 ~ 1,
                                       TRUE ~ 0))
df <- df %>% 
  mutate(discharge_time = case_when(discharge_d >=0 ~ c(discharge_d))) # withdraw_d was overruled for both (death and discharge)
df <- df %>% # restrict to max fup time 28d
  mutate(discharge_time = case_when(discharge_time >28 ~ 28,
                                    TRUE ~ discharge_time))
df <- df %>% # add 28d for those that died for sens-analysis
  mutate(discharge_time_sens = case_when(mort_28 == 1 ~ 28,
                                    TRUE ~ discharge_time))

# (vi) Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days. There were no re-admissions within 28d
df$discharge_reached_sus <- df$discharge_reached
df$discharge_time_sus <- df$discharge_time


# (vii) Viral clearance up to day 5, day 10, and day 15 (Viral load value <LOQ and/or undectectable): Not available
# (viii) Quality of life at day 28: Not available


# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28
df <- df %>% # ae_28 is empty, but under notes there is 1 pulmonary emboly. deaths are in the denominator but counted as 0, since they could have had an ae before death.
  mutate(ae_28 = case_when(note == "pulmonary emboli" ~ 1,
                           TRUE ~ 0))

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
df$ae_28_sev <- df$ae_28 # there was only 1 AE grade 3/4 per person

# (ix) Sens-analysis: Alternative definition/analysis of outcome: time to first (of these) adverse event, within 28 days, considering death as a competing risk (=> censor and set to 28 days)
# time to first ae not available

# (x) Adverse events of special interest within 28 days: a) thromboembolic events (venous thromboembolism, pulmonary embolism, arterial thrombosis), b) secondary infections (bacterial pneumonia including ventilator-associated pneumonia, meningitis and encephalitis, endocarditis and bacteremia, invasive fungal infection including pulmonary aspergillosis), c) Reactivation of chronic infection including tuberculosis, herpes simplex, cytomegalovirus, herpes zoster and hepatitis B, d) serious cardiovascular and cardiac events (including stroke and myocardial infarction), e) events related to signs of bone marrow suppression (anemia, lymphocytopenia, thrombocytopenia, pancytopenia), f) malignancy, g) gastrointestinal perforation (incl. gastrointestinal bleeding/diverticulitis), h) liver dysfunction/hepatotoxicity (grade 3 and 4)
df <- df %>% 
  mutate(aesi_28 = case_when(note == "pulmonary emboli" ~ "thromboembolic events"))

# (xi) Adverse events, any grade and serious adverse event, excluding death, within 28 days, grouped by organ classes
df$ae_28_list <- df$aesi_28 

df_ae <- df %>% 
  select(id_pat, trt, note, ae_28_list, aesi_28)
# Save
save(df_ae, file = "df_ae_ghazaeian.RData")

```
Discussion points OUTCOME data:
1) Get time to first adverse event?


# Define final datasets, set references, summarize missing data and variables
```{r echo=TRUE}
# keep the overall set
df_all <- df
# reduce the df set to our standardized set across all trials
df <- df %>% 
  select(id_pat, trt, sex, age, ethn, trial,
         country, icu, sympdur, 
         #vacc, 
         clinstatus_baseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney,
         any_comorb, comorb_cat, comorb_count,
         crp, 
         # sero, vl_baseline, variant,
         mort_28, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev
         # vir_clear_5, vir_clear_10, vir_clear_15
         )
# export for one-stage model, i.e., add missing variables 
df_os <- df
df_os$vacc <- NA
df_os$sero <- NA
df_os$vl_baseline <- NA
df_os$variant <- NA
df_os$vir_clear_5 <- NA
df_os$vir_clear_10 <- NA
df_os$vir_clear_15 <- NA
# Save
save(df_os, file = "df_os_ghazaeian.RData")

## set references, re-level
# df <- df %>% 
#   mutate(Treatment = relevel(Treatment, "no JAK inhibitor"))

# Create a bar plot to visualize missing values in each column
original_order <- colnames(df_os)
missing_plot <- df_os %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data Visualization") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
```
Discussion points
1) Missing variables:
  Baseline:
  - Vaccination
  - Viremia
  - Variant
  - Serology
  Outcomes:
  - viral load
2) Missing data:
- crp: 2 NAs
- NAs in new_mv_28, aesi_28, ae_28_list: Not part of denominator


# (i) Primary outcome: Mortality at day 28
```{r}
# adjusted for baseline patient characteristics (age, respiratory support at baseline (ordinal scale 1-3 vs 4-5), dexamethasone use at baseline (y/n), remdesivir use at baseline (y/n), anti-IL-6 use at baseline (y/n)).
table(df$mort_28, df$trt, useNA = "always")
mort.28 <- df %>% 
  glm(mort_28 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) all (intervention + control) also had dexa and rdv. No-one had toci.


# (ii) Mortality at day 60
```{r}
table(df$mort_60, df$trt, useNA = "always")
mort.60 <- df %>% 
  glm(mort_60 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.60, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) max fup time was 28 days; thus mort_60 imputed from mort_28


# (iii) Time to death within max. follow-up time
```{r}
# table(df$death_reached, df$death_time, useNA = "always")
# table(df$death_reached, df$mort_60, useNA = "always")

# df %>%
#   drop_na(death_time) %>%
#   filter(death_reached == 1) %>%
#   group_by(trt) %>%
#   summarise(median = median(death_time),
#             IQR = IQR(death_time),
#             Q1 = quantile(death_time, probs = 0.25),
#             Q3 = quantile(death_time, probs = 0.75)) 

# time to death, by group. Kaplan-Meier estimate of conditional survival probability.
km.ttdeath.check <- with(df, Surv(death_time, death_reached))
head(km.ttdeath.check, 100)
km.ttdeath_trt <- survfit(Surv(death_time, death_reached) ~ trt, data=df)
# summary(km.ttdeath_trt, times = 28)
ttdeath_28d_tbl <- km.ttdeath_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d survival (95% CI)**"
  )
# Nicely formatted table
kable(ttdeath_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
# KM curves
survfit2(Surv(death_time, death_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdeath <- df %>% 
  coxph(Surv(death_time, death_reached) ~ trt 
        + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
        , data =.)
ttdeath_reg_tbl <- tbl_regression(ttdeath, exp = TRUE)
# Nicely formatted table
kable(ttdeath_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
# Assessing proportional hazards // also just check KM curve
ph.check <- coxph(Surv(death_time, death_reached) ~ trt
                , data = df)
cz <- cox.zph(ph.check)
print(cz)
plot(cz)
```
Discussion points
1) 


# (iv) New mechanical ventilation among survivors within 28 days
```{r}
table(df$new_mv_28, df$trt, useNA = "always")
new.mv.28 <- df %>% 
  glm(new_mv_28 ~ trt 
       #+ age + clinstatus_baseline 
       #+ comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(new.mv.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
table(df$new_mvd_28, df$trt, useNA = "always")
new.mvd.28 <- df %>% 
  glm(new_mvd_28 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(new.mvd.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```
Discussion points
1) CAVE new_mv_28: Besides the deaths no-one was intubated, and the deaths are excluded => no further events than death => not a single event in either arm!


# (v) Clinical status at day 28
```{r warning=FALSE}
table(df$clinstatus_28_imp, df$trt, useNA = "always")
clin.28 <- df %>% 
  clm(clinstatus_28_imp ~ trt 
      + age 
      #+ clinstatus_baseline 
      #+ comed_dexa + comed_rdv + comed_toci
      , link= c("logit"), data=.)
# Summary and extract coefficients
coefficients_table <- summary(clin.28)$coefficients
# Calculate Odds Ratios and Confidence Intervals
odds_ratios <- exp(coefficients_table[, "Estimate"])
ci_lower <- exp(coefficients_table[, "Estimate"] - 1.96 * coefficients_table[, "Std. Error"])
ci_upper <- exp(coefficients_table[, "Estimate"] + 1.96 * coefficients_table[, "Std. Error"])
# Create a data frame to store Odds Ratios and CIs
clin.28_tbl <- data.frame(
  "Variable" = rownames(coefficients_table),
  "Odds Ratio" = odds_ratios,
  "CI Lower" = ci_lower,
  "CI Upper" = ci_upper
)
# Nicely formatted table
kable(clin.28_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```
Discussion points
1)


# (vi) Time to discharge or reaching discharge criteria up to day 28
```{r warning=FALSE}
# Kaplan-Meier estimate of conditional discharge probability
# Just censoring => Cause-specific hazards
km.ttdischarge.check <- with(df, Surv(discharge_time, discharge_reached))
head(km.ttdischarge.check, 100)
km.ttdischarge_trt <- survfit(Surv(discharge_time, discharge_reached) ~ trt, data=df)
# KM curve
survfit2(Surv(discharge_time, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# table(df$discharge_reached, df$discharge_time)
# df %>% 
#   select(trt, discharge_reached, discharge_time, death_time, mort_28, clinstatus_28_imp) %>% 
#   View()

# testing: cox ph
ttdischarge <- df %>% 
  coxph(Surv(discharge_time, discharge_reached) ~ trt 
        + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
        , data =.)
ttdischarge_reg_tbl <- tbl_regression(ttdischarge, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Sub-distribution hazards
# df <- df %>% # cuminc needs a factor variable with censored patients coded as 0, the event as 1 (discharge) and the competing event as 2 (death).
#   mutate(discharge_reached_comp = case_when (discharge_reached == 0 & (mort_28 == 0 | is.na(mort_28)) ~ 0,
#                                              discharge_reached == 1 & (mort_28 == 0 | is.na(mort_28)) ~ 1,
#                                              mort_28 == 1 ~ 2))
# table(df$discharge_reached_comp) # but in Ghazaeian, we only either have discharged (1) or died (2) in hospital, no-one was still hospitalized at day 28 and thus noone is coded as 0 => Fine-Gray not possible => stick to cause-specific hazards, just censoring
# df$discharge_reached_comp <- as.factor(df$discharge_reached_comp)
# # Cumulative incidence for competing risks
# cuminc(Surv(discharge_time, discharge_reached_comp) ~ 1, data = df)
# cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df) %>%
#   ggcuminc(outcome = c("1", "2")) +
#   ylim(c(0, 1)) +
#   labs(
#     x = "Days"
#   ) +
#   add_confidence_interval() +
#   add_risktable()
# # testing: Fine-Gray regression
# ttdischarge.comp <- crr(Surv(discharge_time, discharge_reached_comp) ~ trt
#     + age
#     # + clinstatus_baseline
#     + comed_dexa + comed_rdv + comed_toci,
#     data = df)
# ttdischarge_comp_reg_tbl <- tbl_regression(ttdischarge.comp, exp = TRUE)
# # Nicely formatted table
# kable(ttdischarge_comp_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
#   kable_styling(bootstrap_options = "striped", full_width = FALSE)
ttdischarge_comp_reg_tbl <- ttdischarge_reg_tbl
ttdischarge.comp <- ttdischarge
# Censoring and assigned worst outcome (28d) to competing event (death) // hypothetical estimand
survfit2(Surv(discharge_time_sens, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sens <- df %>% 
  coxph(Surv(discharge_time_sens, discharge_reached) ~ trt 
        + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
        , data =.)
ttdischarge_sens_reg_tbl <- tbl_regression(ttdischarge.sens, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sens_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Assessing proportional hazards using default discharge_time and discharge_reached
ph.check <- coxph(Surv(discharge_time, discharge_reached) ~ trt 
                , data = df)
cz <- cox.zph(ph.check)
print(cz)
plot(cz)

# Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days
# Use cause-specific hazards
survfit2(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall sustained hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sus <- df %>% 
  coxph(Surv(discharge_time_sus, discharge_reached_sus) ~ trt 
        + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
        , data =.)
ttdischarge_sus_reg_tbl <- tbl_regression(ttdischarge.sus, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sus_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
Discussion points
1) Use F&G for sens-analysis (sustained discharge)?
2) "subdistribution / Fine & Gray" -> but here, the same as cause-specific, since noone was still hospitalized at day 28, all either dead (2) or discharged (1).
3) Those in cont were faster discharged (because the sicker ones were all kept longer at hospital and died?) -> point estimate inconsistent with rest of outcomes.


# (vii) Viral clearance up to day 5, day 10, and day 15
```{r warning=FALSE}

```
Discussion points
1) Not available


# (viii) Quality of life at day 28 
```{r warning=FALSE}

```
Discussion points
1) Not available


# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28
```{r warning=FALSE}
table(df$ae_28, df$trt, useNA = "always") # only 1 event, in int
# Create a 2x2 contingency table
# The oddsRatio function from the mosaic package calculates the odds ratio for a 2 x 2 contingency table and a confidence interval for the each estimate. x should be a matrix, data frame or table. "Successes" should be located in column 1 of x, and the treatment of interest should be located in row 2. The odds ratio is calculated as (Odds row 2) / (Odds row 1). The confidence interval is calculated from the log(OR) and back-transformed.
tbl <- tibble(
  Event = c(0.5, 1.5), # add the 0.5 correction, intervention second
  NoEvent = c(51.5, 45.5) # add the 0.5 correction, intervention second
)
oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
ae.28.corr <- oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28

```
Discussion points
1) only 1 event, in int
2) what to do with ae.28.sev



# Subgroup analysis: Ventilation requirement (proxy for disease severity) on primary endpoint
```{r warning=FALSE}
# table(df$clinstatus_baseline, df$mort_28, useNA = "always") # 2 - 3 included
table(df$clinstatus_baseline, df$mort_28, df$trt, useNA = "always") # 0 in 2x2 table
# class(df$clinstatus_baseline)
df$clinstatus_baseline_f <- df$clinstatus_baseline
df$clinstatus_baseline <- as.numeric(df$clinstatus_baseline)

mort.28.vent <- df %>% 
  glm(mort_28 ~ trt*clinstatus_baseline
      #+ age 
      #+ clinstatus_baseline 
      #+ comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.vent, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```
Discussion points
1) How to apply rare event correction to interaction estimation?


# Subgroup analysis: Age on primary endpoint
```{r warning=FALSE}
mort.28.age <- df %>% 
  glm(mort_28 ~ trt*age
      # + age 
      #+ clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.age, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) 


# Subgroup analysis: Comorbidities on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction
table(df$comorb_cat, df$mort_28, useNA = "always") 
# class(df$comorb_cat)
mort.28.comorb <- df %>%
  glm(mort_28 ~ trt*comorb_cat 
      #+ age 
      #+ clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.comorb, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# 4 comorbidity categories as factor
df$comorb_cat_f <- as.factor(df$comorb_cat)
# table(df$comorb_cat_f, df$mort_28, useNA = "always") 
mort.28.comorb.f <- df %>% 
  glm(mort_28 ~ trt*comorb_cat_f 
      + age 
      #+ clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.comorb.f, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# full comorbidity count
# table(df$comorb_count, df$mort_28, useNA = "always") 
mort.28.comorb.count <- df %>%
  glm(mort_28 ~ trt*comorb_count 
      + age 
      + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.comorb.count, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) Numeric or factor or count?


# Subgroup analysis: Concomitant COVID-19 treatment on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction

# table(df$comed_cat, df$trt, useNA = "always")
# 1: patients without Dexamethasone nor Tocilizumab => JAKi effect alone
# 2: patients with Dexamethasone and Tocilizumab => JAKi effect with Dexa + Toci
# 3: patients with Dexamethasone but no Tocilizumab => JAKi effect with Dexa only
# 4: patients with Tocilizumab but no Dexamethasone (if exist) => JAKi effect with Toci only 
mort.28.comed <- df %>%
  glm(mort_28 ~ trt*comed_cat 
      + age 
      + clinstatus_baseline 
      #+ comed_dexa 
      + comed_rdv 
      #+ comed_toci
      , family = "binomial", data=.)
summ(mort.28.comed, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# comedication as ordinal factor
# df$comed_cat_f <- factor(df$comed_cat, levels = 1:4)
# table(df$comed_cat_f, df$mort_28, useNA = "always") 
# mort.28.comed.f <- df %>%
#   glm(mort_28 ~ trt*comed_cat_f 
#       + age 
#       + clinstatus_baseline 
#       # + comed_dexa 
#       + comed_rdv 
#       # + comed_toci
#       , family = "binomial", data=.)
# summ(mort.28.comed.f, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) Only Category 3 available => not possible / all (intervention + control) also had dexa and rdv. No-one had toci.


# Subgroup analysis: Vaccination on adverse events
```{r warning=FALSE}

```
Discussion points
1) Vaccination data not available


# SENS Subgroup analysis: Duration since symptom onset on primary endpoint
```{r warning=FALSE}
# table(df$sympdur, df$mort_28, useNA = "always")
mort.28.symp <- df %>% 
  glm(mort_28 ~ trt*sympdur
      + age 
      #+ clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.symp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) 


# SENS Subgroup analysis: CRP on primary endpoint
```{r warning=FALSE}
# table(df$crp, df$mort_28, useNA = "always")
mort.28.crp <- df %>% 
  glm(mort_28 ~ trt*crp
      + age 
      #+ clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28.crp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1) Truncated or not? How to standardize across studies (see Barisolidact)


# SENS Subgroup analysis: variant on primary endpoint
```{r warning=FALSE}

```
Discussion points
1) No data on variant available


# Collect all treatment effect estimates across endpoints (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
result_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric()
)
# Function to extract treatment results from different model types (glm, clm, coxph and crr)
extract_trt_results <- function(model, variable_name) {
  if (inherits(model, "glm") || inherits(model, "clm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "coxph")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "se(coef)"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "tidycrr")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- c(exp(model$tidy$conf.low[1]), exp(model$tidy$conf.high[1]))
    se <- model$tidy$std.error[1]
    p_value <- model$tidy$p.value[1]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_trt_results(mort.28, "death at day 28")
result_list[[2]] <- extract_trt_results(mort.60, "death at day 60")
result_list[[3]] <- extract_trt_results(ttdeath, "death within fup")
# result_list[[x]] <- extract_trt_results(new.mv.28, "new MV within 28d") // not possible
result_list[[4]] <- extract_trt_results(new.mvd.28, "new MV or death within 28d")
result_list[[5]] <- extract_trt_results(clin.28, "clinical status at day 28")
# result_list[[x]] <- extract_trt_results(ttdischarge, "discharge within 28 days")
result_list[[6]] <- extract_trt_results(ttdischarge.comp, "discharge within 28 days, death=comp.event")
# result_list[[7]] <- extract_trt_results(ttdischarge.sens, "discharge within 28 days, death=hypo.event")
# result_list[[8]] <- extract_trt_results(ttdischarge.sus, "sustained discharge within 28 days")
# result_list[[9]] <- extract_trt_results(vir.clear.5, "viral clearance until day 5")
# result_list[[10]] <- extract_trt_results(vir.clear.10, "viral clearance until day 10")
# result_list[[11]] <- extract_trt_results(vir.clear.15, "viral clearance until day 15")
# result_list[[12]] <- extract_trt_results(ae.28, "any AE grade 3,4 within 28 days")

# Filter out NULL results and bind the results into a single data frame
result_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name
result_df$trial <- "Ghazaeian"


### Add the corrected models
# ae.28.corr
# summary(ae.28.corr)
hazard_odds_ratio <- ae.28.corr[["OR"]]
ci_lower <- c(0.1349) # how to extract directly from the object?
ci_upper <- c(85.44) # how to extract directly from the object?
standard_error <- (ci_upper - ci_lower) / (2 * 1.96)
z_score <- (log(hazard_odds_ratio) - log(1)) / standard_error
p_value <- 2 * (1 - pnorm(abs(z_score)))
new_row <- data.frame(
    variable = "any AE grade 3,4 within 28 days",
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci_lower,
    ci_upper = ci_upper,
    standard_error = standard_error,
    p_value = p_value,
    trial = "Ghazaeian")
result_df <- rbind(result_df, new_row)

# Nicely formatted table
kable(result_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
save(result_df, file = "trt_effects_ghazaeian.RData")
```
Discussion points
1) 


# Collect all interaction estimates (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
interaction_df <- data.frame(
  variable = character(),
  log_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric()
)
# Extract and format results for the interaction term
extract_interaction <- function(model, variable_name) {
      trt_coef <- coef(model)[grep("^trt:", names(coef(model)))]
      log_odds_ratio <- exp(trt_coef)
      ci <- exp(confint(model)[grep("^trt:", names(coef(model))), ])
      se <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Std. Error"]
      p_value <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Pr(>|z|)"]
      # capture the results
      result <- data.frame(
        variable = variable_name,
        log_odds_ratio = log_odds_ratio,
        ci_lower = ci[1],
        ci_upper = ci[2],
        standard_error = se,
        p_value = p_value
      )
    return(result)
}

# Loop through
result_list <- list()

# result_list[[1]] <- extract_interaction(mort.28.vent, "respiratory support")
result_list[[3]] <- extract_interaction(mort.28.age, "age")
result_list[[4]] <- extract_interaction(mort.28.comorb, "comorbidity")
# result_list[[x]] <- extract_interaction(mort.28.comed, "comedication") // not possible
result_list[[5]] <- extract_interaction(mort.28.symp, "symptom duration")
result_list[[6]] <- extract_interaction(mort.28.crp, "crp")
# result_list[[x]] <- extract_interaction(mort.28.var, "variant") // not available

# Filter out NULL results and bind the results into a single data frame
interaction_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name
interaction_df$trial <- "Ghazaeian"

# Nicely formatted table
kable(interaction_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
save(interaction_df, file = "int_effects_ghazaeian.RData")
```
Discussion points
1) decide what to do with respiratory support!

