---
title: "one-stage"
author: "A.Amstutz"
date: "2023-10-18"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)
library(here)
library(kableExtra)

library(jtools) # for summ() and plot_summs
library(sjPlot) # for tab_model
library(ggplot2) # survival/TTE analyses and other graphs
library(ggsurvfit) # survival/TTE analyses
library(survival) # survival/TTE analyses
library(gtsummary) # survival/TTE analyses
library(ggfortify) # autoplot
library(lme4) # glmer / clmm
library(sjPlot) # for tab_model
library(glmmTMB) # to specify estimation method explicitly -> i.e. ML
library(coxme) # for mixed-effects cox model
library(ordinal) # for ordinal regression, clm (fixed effects) & clmm (mixed effects)
library(tidycmprsk) # competing risk analysis
library(logistf) # Firth regression in case of rare events

library(meta)
library(forestplot)
library(metafor) #forest()

```

# Load standardized dataset of all trials
```{r echo=TRUE, message=FALSE, warning=FALSE}
## barisolidact
df_barisolidact <- readRDS("df_os_barisolidact.RData")
df_barisolidact <- df_barisolidact %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## actt2
df_actt2 <- readRDS("df_os_actt2.RData")
df_actt2 <- df_actt2 %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## ghazaeian
df_ghazaeian <- readRDS("df_os_ghazaeian.RData")
df_ghazaeian <- df_ghazaeian %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## tofacov
df_tofacov <- readRDS("df_os_tofacov.RData")
df_tofacov <- df_tofacov %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## covinib
df_covinib <- readRDS("df_os_covinib.RData")
df_covinib <- df_covinib %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## COV-BARRIER
df_covbarrier <- readRDS("df_os_cov-barrier.RData")
df_covbarrier <- df_covbarrier %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## Murugesan
df_murugesan <- readRDS("df_os_murugesan.RData")
df_murugesan <- df_murugesan %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)
## RECOVERY
df_recovery <- readRDS("df_os_recovery.RData")
df_recovery <- df_recovery %>% 
    select(id_pat, trial, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney, any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15)

# append
df_tot <- rbind(df_barisolidact, df_actt2, df_ghazaeian, df_tofacov, df_covinib, df_covbarrier, df_recovery)
df_tot_Muru <- rbind(df_barisolidact, df_actt2, df_ghazaeian, df_tofacov, df_covinib, df_covbarrier, df_murugesan, df_recovery)

# Save
saveRDS(df_tot, file = "df_tot.RData")
write_xlsx(df_tot, path = "df_tot.xlsx")
saveRDS(df_tot_Muru, file = "df_tot_Muru.RData")

```


# (i) Primary outcome: Mortality at day 28
```{r warning=FALSE}
addmargins(table(df_tot$trial, df_tot$mort_28, useNA = "always"))
addmargins(table(df_tot$mort_28, df_tot$trt, useNA = "always"))
# table(df_tot$age, df_tot$trt, useNA = "always")
table(df_tot$clinstatus_baseline, df_tot$trt, useNA = "always")
table(df_tot$clinstatus_baseline, df_tot$trial, useNA = "always")
table(df_tot$vbaseline, df_tot$trial, useNA = "always")

# reformatting
df_tot$trt_f <- as.factor(df_tot$trt)
df_tot$trial_f <- as.factor(df_tot$trial)
df_tot$clinstatus_baseline_n <- as.numeric(df_tot$clinstatus_baseline)
df_tot <- df_tot %>%
  mutate(trial_n = case_when(trial == "Bari-Solidact" ~ 1,
                             trial == "ACTT2" ~ 2,
                             trial == "Ghazaeian" ~ 3,
                             trial == "TOFACOV" ~ 4,
                             trial == "COVINIB" ~ 5,
                             trial == "COV-BARRIER" ~ 6,
                             trial == "RECOVERY" ~ 7,
                             # trial == "Murugesan" ~ 8
                             ))
table(df_tot$trial_n, useNA = "always")
```

GOAL: random treatment effect, stratified trial intercept, stratified prognostic factors, AND centering the treatment variable by the proportion treated in the trial (to improve the estimation of between-study variance) AND maximum likelihood (ML) estimator (due to small trials with rare events). REML is default in glmer, for ML use glmmTMB. See notes.

### (1) common treatment effect, random trial intercept, common prognostic factors
```{r}
# (1) common treatment effect, random trial intercept, common prognostic factors, ML
mort28.ctreat.rtrial.ml <- glmmTMB(mort_28 ~ trt + (1|trial)
                              + age + clinstatus_baseline
                              , data = df_tot, family = binomial)
tab_model(mort28.ctreat.rtrial.ml)

mort28.ctreat.rtrial.ml.vb <- glmmTMB(mort_28 ~ trt + (1|trial)
                              + age + vbaseline
                              , data = df_tot, family = binomial)
# tab_model(mort28.ctreat.rtrial.ml.vb)
```

### (2) random treatment effect, random trial intercept, common prognostic factors and residual variances
```{r}
# (3) random treatment effect, random trial intercept, common prognostic factors and residual variances, ML
mort28.rtreat.rtrial.ml <- glmmTMB(mort_28 ~ trt + (1 + trt|trial_f)
                              + age + clinstatus_baseline
                              , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.rtrial.ml)

mort28.rtreat.rtrial.ml.vb <- glmmTMB(mort_28 ~ trt + (1 + trt|trial_f)
                              + age + vbaseline
                              , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.rtrial.ml.vb)
```

### (3) random treatment effect, stratified trial intercept, common prognostic factors and residual variances
```{r}
# (2) random treatment effect, stratified trial intercept, common prognostic factors and residual variances, and WITHOUT centering the treatment variable, ML
mort28.rtreat.strial.ml <- glmmTMB(mort_28 ~ trt_f + trial_f + (trt - 1 | trial_f)
                              + age + clinstatus_baseline
                              , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.ml)

mort28.rtreat.strial.ml.vb <- glmmTMB(mort_28 ~ trt_f + trial_f + (trt - 1 | trial_f)
                              + age + vbaseline 
                              , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.ml.vb)

# The - 1 within (trt - 1 | trial_n) specifies that there is no random intercept for the grouping factor trial_n. We're specifying random slopes for the variable trt within each level of trial_n, i.e., the effect may vary from one trial to another. Together with trial_f, this gives the stratified intercept model. This is a more flexible model compared to a model with a random intercept, which assumes a common baseline for all groups.

# dummy variable for each trial (trial_1, trial_2, trial_3, etc - e.g. where trial_1 = 1 if in trial 1 and 0 otherwise)
# df_tot <- df_tot %>%
#   mutate(trial_1 = case_when(trial == "Bari-Solidact" ~ 1, TRUE ~ 0),
#          trial_2 = case_when(trial == "ACTT2" ~ 1, TRUE ~ 0),
#          trial_3 = case_when(trial == "Ghazaeian" ~ 1, TRUE ~ 0),
#          trial_4 = case_when(trial == "TOFACOV" ~ 1, TRUE ~ 0),
#          trial_5 = case_when(trial == "COVINIB" ~ 1, TRUE ~ 0),
#          trial_6 = case_when(trial == "COV-BARRIER" ~ 1, TRUE ~ 0),
#          trial_7 = case_when(trial == "RECOVERY" ~ 1, TRUE ~ 0))

## Use "Stata syntax":
# mort28.rtreat.strial.2 <- glmer(mort_28 ~ trt_f + trial* + (trt -1 | trial_n)
#                               + age + clinstatus_baseline_n
#                               # + comed_dexa + comed_rdv + comed_toci
#                               , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.2)
```

### (4) random treatment effect, stratified trial intercept, common prognostic factors and residual variances, AND centering the treatment variable
```{r}
# (4) random treatment effect, stratified trial intercept, common prognostic factors and residual variances, ML, but WITH centering the treatment variable

# calculate the proportion treated by trial
proportions <- df_tot %>%
  group_by(trial) %>%
  summarize(proportion_treated = sum(trt) / n())
df_tot <- left_join(df_tot, proportions[, c("proportion_treated", "trial")], by = join_by(trial == trial))
# table(df_tot$trial, df_tot$proportion_treated)

# create the centered treatment variable
df_tot$trt_centered_n <- df_tot$trt - df_tot$proportion_treated
df_tot$trt_centered_f <- as.factor(df_tot$trt_centered_n)
# table(df_tot$trial, df_tot$trt_centered_f) ## keep it minus?

# build the model. '-1' indicating there is no overall random intercept.
mort28.rtreat.strial.cent.ml <- glmmTMB(mort_28 ~ trt_centered_n + trial_f + (trt_centered_n -1 | trial_f) -1
                   + age + clinstatus_baseline
                   , data = df_tot, family = binomial
                   )
# tab_model(mort28.rtreat.strial.cent.ml)

mort28.rtreat.strial.cent.ml.vb <- glmmTMB(mort_28 ~ trt_centered_n + trial_f + (trt_centered_n -1 | trial_f) -1 
                   + age + vbaseline 
                   , data = df_tot, family = binomial
                   )
# tab_model(mort28.rtreat.strial.cent.ml.vb)
```

### (5) random treatment effect, stratified trial intercept, stratified prognostic factors and residual variances, with centering the treatment variable AND the prognostic variables
```{r}
# "Our default recommendation is to use stratified prognostic effects (i.e. estimate a separate effect of each included prognostic factor for each trial), with trial-specific centering of each prognostic factor to improve ML estimation (for the reasons explained in Section 6.2.8.3). However, if outcome data or prognostic factor categories are sparse, the stratification approach may lead to estimation difficulties, and then allowing prognostic factor effects to be random is a sensible alternative." (p. 145) Centering disentangles (i.e. makes uncorrelated) the estimation of main parameters of interest from other nuisance parameters, which leads to less downward bias in estimates of variance parameters (Figure 6.2) and thus improves coverage of 95% confidence intervals. This can be achieved by centering the covariates by their mean values within trials, such that the interaction estimate is then only based on within-trial information.

# Calculate the mean values of (continuous) prognostic factors within each trial
trial_means <- df_tot %>%
  group_by(trial) %>%
  summarize(mean_age = mean(age, na.rm = TRUE), mean_clinstatus = mean(clinstatus_baseline_n, na.rm = TRUE))
# Merge back
df_tot <- df_tot %>% left_join(trial_means, by = "trial")
# Center the age and clinstatus_baseline variables
df_tot <- df_tot %>%
  mutate(age_centered = age - mean_age, 
         clinstatus_baseline_centered = clinstatus_baseline_n - mean_clinstatus)

# df_tot %>%
#   select(trial_f, trial_n, trt, trt_centered_n, age, mean_age, age_centered, clinstatus_baseline_n, mean_clinstatus, clinstatus_baseline_centered, mort_28) %>%
#   View()


### Add the prognostic factors stratified, i.e. create stratified variables for each prognostic factor
## uncentered
df_tot <- df_tot %>%
  mutate(age_trial_1 = case_when(trial == "Bari-Solidact" ~ age, TRUE ~ 0),
         age_trial_2 = case_when(trial == "ACTT2" ~ age, TRUE ~ 0),
         age_trial_3 = case_when(trial == "Ghazaeian" ~ age, TRUE ~ 0),
         age_trial_4 = case_when(trial == "TOFACOV" ~ age, TRUE ~ 0),
         age_trial_5 = case_when(trial == "COVINIB" ~ age, TRUE ~ 0),
         age_trial_6 = case_when(trial == "COV-BARRIER" ~ age, TRUE ~ 0),
         age_trial_7 = case_when(trial == "RECOVERY" ~ age, TRUE ~ 0))
df_tot <- df_tot %>%
  mutate(clinstat_trial_1 = case_when(trial == "Bari-Solidact" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_2 = case_when(trial == "ACTT2" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_3 = case_when(trial == "Ghazaeian" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_4 = case_when(trial == "TOFACOV" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_5 = case_when(trial == "COVINIB" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_6 = case_when(trial == "COV-BARRIER" ~ clinstatus_baseline, TRUE ~ "0"),
         clinstat_trial_7 = case_when(trial == "RECOVERY" ~ clinstatus_baseline, TRUE ~ "0"))

## centered
df_tot <- df_tot %>%
  mutate(age_cent_trial_1 = case_when(trial == "Bari-Solidact" ~ age_centered, TRUE ~ 0),
         age_cent_trial_2 = case_when(trial == "ACTT2" ~ age_centered, TRUE ~ 0),
         age_cent_trial_3 = case_when(trial == "Ghazaeian" ~ age_centered, TRUE ~ 0),
         age_cent_trial_4 = case_when(trial == "TOFACOV" ~ age_centered, TRUE ~ 0),
         age_cent_trial_5 = case_when(trial == "COVINIB" ~ age_centered, TRUE ~ 0),
         age_cent_trial_6 = case_when(trial == "COV-BARRIER" ~ age_centered, TRUE ~ 0),
         age_cent_trial_7 = case_when(trial == "RECOVERY" ~ age_centered, TRUE ~ 0))
df_tot <- df_tot %>%
  mutate(clinstat_cent_trial_1 = case_when(trial == "Bari-Solidact" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_2 = case_when(trial == "ACTT2" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_3 = case_when(trial == "Ghazaeian" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_4 = case_when(trial == "TOFACOV" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_5 = case_when(trial == "COVINIB" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_6 = case_when(trial == "COV-BARRIER" ~ clinstatus_baseline_centered, TRUE ~ 0),
         clinstat_cent_trial_7 = case_when(trial == "RECOVERY" ~ clinstatus_baseline_centered, TRUE ~ 0))

# df_tot %>%
#   select(trial_f, trial_n, trt, trt_centered_n, age, mean_age, age_centered, clinstatus_baseline_n, mean_clinstatus, clinstatus_baseline_centered, mort_28,
#          age_trial_1, age_cent_trial_1, age_trial_5, age_cent_trial_5, age_trial_6, age_cent_trial_6, clinstat_trial_1, clinstat_cent_trial_1, clinstat_trial_5, clinstat_cent_trial_5, clinstat_trial_6, clinstat_cent_trial_6,
#          ) %>%
#   View()


# (5) random treatment effect, stratified trial intercept, stratified prognostic factors and residual variances, with centering of the treatment variable AND the prognostic factors

mort28.rtreat.strial.cent.ml.spf.cent <- glmmTMB(mort_28 ~ trt_centered_n
                                                 + trial_f
                                                 + (trt_centered_n -1 | trial_f) -1
                                                 + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 + age_cent_trial_4
                                                 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                                                 + clinstat_cent_trial_1 + clinstat_cent_trial_2 + clinstat_cent_trial_3
                                                 + clinstat_cent_trial_4 + clinstat_cent_trial_5 + clinstat_cent_trial_6 + clinstat_cent_trial_7
                                                 , data = df_tot, family = binomial)
tab_model(mort28.rtreat.strial.cent.ml.spf.cent)

# mort28.rtreat.strial.cent.ml.spf.cent <- glmmTMB(mort_28 ~ trt_centered_n
#                                                  + trial_f
#                                                  + (trt_centered_n -1 | trial_f) -1
#                                                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 + age_cent_trial_4
#                                                  + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
#                                                  + clinstat_trial_1 + clinstat_trial_2 + clinstat_trial_3
#                                                  + clinstat_trial_4 + clinstat_trial_5 + clinstat_trial_6 + clinstat_trial_7
#                                                  , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.cent.ml.spf.cent)

# Not all trials have all levels, cannot estimate a stratified clinstatus_baseline by trial. Moreover, not sure the centering of the factor variable clinstatus_baseline is ok like this. => Second best option: Add clinstatus_baseline as a random parameter

mort28.rtreat.strial.cent.ml.sage.cent.rclinstat <- glmmTMB(mort_28 ~ trt_centered_n 
                                                 + trial_f 
                                                 + (trt_centered_n -1 | trial_f) -1 
                                                 + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 + age_cent_trial_4 
                                                 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                                                 + (clinstatus_baseline_n -1 | trial_f)
                                                 , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.cent.ml.sage.cent.rclinstat) 

mort28.rtreat.strial.cent.ml.sage.cent.rvb <- glmmTMB(mort_28 ~ trt_centered_n 
                                                 + trial_f 
                                                 + (trt_centered_n -1 | trial_f) -1 
                                                 + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 + age_cent_trial_4 
                                                 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                                                 + (vbaseline -1 | trial_f)
                                                 , data = df_tot, family = binomial)
# tab_model(mort28.rtreat.strial.cent.ml.sage.cent.rvb) 

```
Discussion:
1. Clinstatus_baseline is a factor -> How to center a factor? By the proportion in each level? More important: Not all trials have all levels, hard to estimate a stratified clinstatus_baseline by trial => Second best option: Add clinstatus_baseline as a random parameter
2. Remember to add all TRIAL centering variables if adding new trials!
3. Investigate the point that a "one-stage random effects model" equals a "two-stage fixed effects model" (only use trials without rare event problems // double-check adjustments in two-stage)


## Collect the effect estimates from all models
```{r}
# Empty data frame to store the results
result_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  p_value = numeric()
)
# Function to extract treatment results from two different model types (glmer -> REML, glmmTMB -> ML)
extract_trt_results <- function(model, variable_name, n_int, n_cont) {
  if (inherits(model, "glmmTMB")) {
    coefficients_table <- summary(model)$coefficients$cond
    trt_coef <- coefficients_table[grep("^trt", rownames(coefficients_table)), "Estimate"]
    hazard_odds_ratio <- exp(trt_coef)
    ci_table <- exp(confint(model))
    ci <- ci_table[grep("^trt", rownames(ci_table)), c("2.5 %", "97.5 %")]
    p_value <- coefficients_table[grep("^trt", rownames(coefficients_table)), "Pr(>|z|)"]
  } else if (inherits(model, "glmerMod")) {
    coefficients_table <- summary(model)$coefficients
    trt_coef <- coefficients_table[grep("^trt", rownames(coefficients_table)), "Estimate"]
    hazard_odds_ratio <- exp(trt_coef)
    ci_table <- exp(confint(model))
    ci <- ci_table[grep("^trt", rownames(ci_table)), c("2.5 %", "97.5 %")]
    p_value <- coefficients_table[grep("^trt", rownames(coefficients_table)), "Pr(>|z|)"]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    p_value = p_value,
    n_intervention = n_int,
    n_control = n_cont
  )
  return(result)
}
# Loop through
result_list <- list()

result_list[[1]] <- extract_trt_results(mort28.ctreat.rtrial.ml, "c trt, r intercept, c age/clinstatus", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
# result_list[[2]] <- extract_trt_results(mort28.ctreat.rtrial.ml.vb, "c trt, r intercept, c age/clinstatus, no cent, vb", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])

result_list[[3]] <- extract_trt_results(mort28.rtreat.rtrial.ml, "r trt, r intercept, c age/clinstatus", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
# result_list[[4]] <- extract_trt_results(mort28.rtreat.rtrial.ml.vb, "r trt, r intercept, c age/clinstatus, no cent, vb", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])

result_list[[5]] <- extract_trt_results(mort28.rtreat.strial.ml, "r trt, s intercept, c age/clinstatus", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
# result_list[[6]] <- extract_trt_results(mort28.rtreat.strial.ml.vb, "r trt, s intercept, c age/clinstatus, no cent, vb", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])

result_list[[7]] <- extract_trt_results(mort28.rtreat.strial.cent.ml, "r cent trt, s intercept, c age/clinstatus", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
# result_list[[8]] <- extract_trt_results(mort28.rtreat.strial.cent.ml.vb, "r trt, s intercept, c age/clinstatus, cent trt, vb", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])

result_list[[9]] <- extract_trt_results(mort28.rtreat.strial.cent.ml.sage.cent.rclinstat, "r cent trt, s intercept, s and cent age, r clinstatus", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
# result_list[[10]] <- extract_trt_results(mort28.rtreat.strial.cent.ml.sage.cent.rvb, "r trt, s intercept, s and cent age, r clinstatus, cent trt, vb", addmargins(table(df_tot$mort_28, df_tot$trt))[3,2], addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])

# Filter out NULL results and bind the results into a single data frame
result_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Nicely formatted table
kable(result_df, format = "html", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

### Plot the effect estimates and 95% CI for all endpoints
```{r}
# Convert necessary columns to numeric
result_df[, c("hazard_odds_ratio", "ci_lower", "ci_upper", "p_value", "n_intervention", "n_control")] <- lapply(result_df[, c("hazard_odds_ratio", "ci_lower", "ci_upper", "p_value", "n_intervention", "n_control")], as.numeric)

result_df$variable <- factor(result_df$variable, 
                             levels = c("c trt, r intercept, c age/clinstatus", 
                                        "r trt, r intercept, c age/clinstatus", 
                                        "r trt, s intercept, c age/clinstatus",
                                        "r cent trt, s intercept, c age/clinstatus",
                                        "r cent trt, s intercept, s and cent age, r clinstatus"))

# Plotting the reordered data


# Plotting
ggplot(result_df, aes(x = variable, y = hazard_odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.3) +
  labs(title = "Mortality at Day 28 - All models",
       x = "Model",
       y = "Odds Ratio") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = margin(l = 50, r = 10, b = 20, t = 20, unit = "pt")) +  # Adjust left margin
  scale_y_continuous(limits = c(0.5, 1.0), breaks = c(0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1.0), trans = "log10") +
  coord_flip()  # Flip axes to show longer variable names

```

### The main recommendations for one-stage IPD meta-analysis models using GLMMs (IPDMA handbook R.Riley)
1. *__Use a random treatment effect.__*
* Justification: Typically the included trials are conducted in different settings, populations and time periods. Therefore, some heterogeneity of treatment effect is expected. Heterogeneity might be reduced by inclusion of prognostic factors or trial-level covariates, but usually unexplained heterogeneity remains and so should be acknowledged. Homogeneity of treatment effect is a strong assumption, and often will be inappropriate due to unexplained between-trial heterogeneity. To address this, the treatment effect parameter can be made random, such that the true treatment effect in each trial is assumed drawn from a particular distribution, typically a normal distribution.

2. *__Stratify by trial the intercept and parameters for other non-treatment variables (such as prognostic factors and residual variances).__* If convergence issues arise, then consider making the intercept (and other non-treatment variables) random.
* Justification: Although a random intercept will usually give similar results to a stratified intercept, in some situations it may compromise randomisation (as it allows baseline risk information to be shared across trials). Many researchers assume nuisance parameters are common (often because this is the default in software packages), but this is not recommended as it may lead to inappropriate conclusions, as now described. (p. 138) The advantage of the stratified intercept approach is that it makes no assumptions about the distribution of intercepts across trials - and mirrors exactly the two-stage approach. The advantage of the random intercepts approach is that it requires fewer parameters to be estimated and so may reduce model convergence issues. But usually give very similar results. (p. 141 & 143)

3. *__Use trial-specific centering of the treatment variable (and any other included variables, such as prognostic factors) when using ML estimation of a one-stage model with a stratified intercept.__* 
* Justification: Simulation studies and mathematical reasoning show that this improves ML estimation of between-trial variances and the coverage of confidence intervals for the summary treatment effect. Centering disentangles (i.e. makes uncorrelated) the estimation of main parameters of interest from other nuisance parameters, which leads to less downward bias in estimates of variance parameters (Figure 6.2) and thus improves coverage of 95% confidence intervals. This can be achieved by centering the covariates by their mean values within trials, such that the interaction estimate is then only based on within-trial information. Our default recommendation is to use stratified prognostic effects (i.e. estimate a separate effect of each included prognostic factor for each trial), with trial-specific centering of each prognostic factor to improve ML estimation (for the reasons explained in Section 6.2.8.3). However, if outcome data or prognostic factor categories are sparse, the stratification approach may lead to estimation difficulties, and then allowing prognostic factor effects to be random is a sensible alternative. (p. 145) As previously discussed (Section 6.2.4.1), Jackson et al. and Riley et al. show that for one-stage models with a stratified intercept, ML estimation is improved when using trial-specific centering of treatment and other included variables.181,185 Centering disentangles (i.e. makes uncorrelated) the estimation of main parameters of interest from other nuisance parameters, which leads to less downward bias in estimates of variance parameters (Figure 6.2) and thus improves coverage of 95% confidence intervals. (p. 147)

4. For frequentist estimation of one-stage models for binary, ordinal or count outcomes, use REML estimation of the pseudo-likelihood approach unless most trials in the meta-analysis are small (in terms of participants or outcome events), which then warrants ML estimation of the exact likelihood.
* Justification: Although estimation of the exact likelihood is preferred, ML estimation is known to produce downwardly biased estimates of between-trial variances. Therefore, unless most included trials are small, REML estimation of an approximate pseudo-likelihood specification may improve estimation of between-trial variances and coverage of confidence intervals.

5. A one-stage IPD meta-analysis utilises a more exact statistical likelihood than a two-stage meta-analysis approach, which is advantageous when included trials have few participants or outcome events. (p.127)

6. The word ‘fixed’ is ambiguous; it could refer to either a common or stratified parameter, even though they imply different model specifications and assumptions. Therefore, we recommend that the word ‘fixed’ be avoided in one-stage IPD models, and encourage researchers to use common or stratified instead.

7. See sample R code here: https://www.ipdma.co.uk/one-stage-ipd-ma 

## Present the main model: "r cent trt, s intercept, s and cent age, r clinstatus"
1. Treatment parameter: Random treatment effect
2. Trial parameter: Stratified intercept by trial
3. Prognostic factor 'age': Stratified by trial
4. Prognostic factor 'clinical status': As a random parameter by trial. Not all trials have all levels, hard to estimate a stratified clinstatus_baseline by trial => Random effect as sensible alternative (see guidance above)
5. Trial-specific centering of the treatment variable and age (to improve estimation of the between-study variance). 

# (i) Mortality at day 28
```{r}
mort28 <- glmmTMB(mort_28 ~ trt_centered_n 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7 
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline
                  , data = df_tot, family = binomial)
tab_model(mort28)

```

# (ii) Mortality at day 60
```{r warning=FALSE}
addmargins(table(df_tot$mort_60, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$mort_60, df_tot$trial, useNA = "always"))

mort60 <- glmmTMB(mort_60 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7 
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(mort60)
```

# (iii) Time to death within max. follow-up time
```{r warning=FALSE}
# table(df_tot$death_reached, df_tot$death_time, useNA = "always")
# table(df_tot$death_reached, df_tot$mort_60, useNA = "always") # 1 death after day 60 in Bari-Solidact

# df_tot %>%
#   drop_na(death_time) %>%
#   filter(death_reached == 1) %>%
#   group_by(trt) %>%
#   summarise(median = median(death_time),
#             IQR = IQR(death_time),
#             Q1 = quantile(death_time, probs = 0.25),
#             Q3 = quantile(death_time, probs = 0.75))

# cap at 60 days // always censor first the censoring variable..
df_tot <- df_tot %>%
  mutate(death_reached_60 = case_when(death_time >60 ~ 0,
                                   TRUE ~ c(death_reached)))
df_tot <- df_tot %>%
  mutate(death_time_60 = case_when(death_time >60 ~ 60,
                                TRUE ~ c(death_time)))

## time to death, by group. Kaplan-Meier estimate of conditional survival probability.
# km.ttdeath.check <- with(df_tot, Surv(death_time, death_reached))
# head(km.ttdeath.check, 100)
km.ttdeath_trt <- survfit(Surv(death_time, death_reached) ~ trt, data=df_tot)
# summary(km.ttdeath_trt, times = 28)
ttdeath_28d_tbl <- km.ttdeath_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d survival (95% CI)**"
  )
# Nicely formatted table
kable(ttdeath_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

autoplot(km.ttdeath_trt)
survfit2(Surv(death_time, death_reached) ~ trt, data=df_tot) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable()

# autoplot by trial with max. 60d fup
km.ttdeath_trial <- survfit(Surv(death_time_60, death_reached_60) ~ trial, data=df_tot)
autoplot(km.ttdeath_trial)

# Assessing proportional hazards // check the KM curve: OK

# Cox proportional hazards model adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
ttdeath <- coxme(Surv(death_time, death_reached) ~ trt_centered_n
                  + trial_f
                  + (trt_centered_n -1 | trial_f) -1
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot)
tab_model(ttdeath)
```

# (iv) New mechanical ventilation or death within 28 days
```{r warning=FALSE}
addmargins(table(df_tot$new_mvd_28, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$new_mvd_28, df_tot$trial, useNA = "always"))

new.mvd28 <- glmmTMB(new_mvd_28 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7 
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(new.mvd28)
```

# (iv.i) New mechanical ventilation among survivors within 28 days
```{r warning=FALSE}
addmargins(table(df_tot$new_mv_28, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$new_mv_28, df_tot$trial, useNA = "always"))

new.mv28 <- glmmTMB(new_mv_28 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7 
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(new.mv28)
```

# (v) Clinical status at day 28
```{r warning=FALSE}
addmargins(table(df_tot$clinstatus_28_imp, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$clinstatus_28_imp, df_tot$trial, useNA = "always"))

# Ordinal regression model adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
clin28 <- df_tot %>% 
    clmm(clinstatus_28_imp ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6
                  + age_cent_trial_7
                  + clinstatus_baseline_n
                  # + (clinstatus_baseline_n -1 | trial_f)
      , link= c("logit"), data=.)
tab_model(clin28)
```

# (vi) Time to discharge or reaching discharge criteria up to day 28. Death = Competing event
```{r warning=FALSE}
# table(df_tot$discharge_reached, df_tot$discharge_time, useNA = "always")
# table(df_tot$discharge_reached, df_tot$trial, useNA = "always")
# table(df_tot$discharge_time, df_tot$trial, useNA = "always")
# 
# df_tot %>%
#   drop_na(discharge_time) %>%
#   filter(discharge_reached == 1) %>%
#   group_by(trt) %>%
#   summarise(median = median(discharge_time),
#             IQR = IQR(discharge_time),
#             Q1 = quantile(discharge_time, probs = 0.25),
#             Q3 = quantile(discharge_time, probs = 0.75))

# Kaplan-Meier estimate of conditional discharge probability
km.ttdischarge.check <- with(df_tot, Surv(discharge_time, discharge_reached))
# head(km.ttdischarge.check, 100)
km.ttdischarge_trt <- survfit(Surv(discharge_time, discharge_reached) ~ trt, data=df_tot)
# summary(km.ttdischarge_trt, times = 28)
ttdischarge_28d_tbl <- km.ttdischarge_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d hospitalization (95% CI)**"
  )
# Nicely formatted table
kable(ttdischarge_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
# KM graph
survfit2(Surv(discharge_time, discharge_reached) ~ trt, data=df_tot) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()

# Assessing proportional hazards (using default discharge_time and discharge_reached) -> see KM plots, after the point when the curves really stqrt diverging, it does not cross over again: OK
# ph.check <- coxph(Surv(discharge_time, discharge_reached) ~ trt
#                 , data = df_tot)
# cz <- cox.zph(ph.check)
# print(cz)
# plot(cz)

## Sub-distribution hazards, i.e., represents the rate per unit of time of the event as well as the influence of competing events. Instantaneous rate of occurrence of the given type of event in subjects who have not yet experienced an event of that type.
df_tot <- df_tot %>% # cuminc needs a factor variable with censored patients coded as 0, the event as 1 and the competing event as 2.
  mutate(discharge_reached_comp = case_when (discharge_reached == 0 & (mort_28 == 0 | is.na(mort_28)) ~ 0,
                                             discharge_reached == 1 & (mort_28 == 0 | is.na(mort_28)) ~ 1,
                                             mort_28 == 1 ~ 2))
df_tot$discharge_reached_comp <- as.factor(df_tot$discharge_reached_comp) 
# Cumulative incidence for the event=discharge (1) and the competing event=death (2)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ 1, data = df_tot)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_tot) %>% 
  ggcuminc(outcome = c("1", "2")) +
  ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()
# in int only
df_int <- df_tot %>% 
  filter(trt == 1)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_int) %>% 
  ggcuminc(outcome = c("1", "2")) +
  #ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()
# in cont only
df_cont <- df_tot %>% 
  filter(trt == 0)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_cont) %>% 
  ggcuminc(outcome = c("1", "2")) +
  #ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()

# Fine-Gray regression model (Competing Risk regression), adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
## problem: How to incorporate clustering and random effects?
### crr has a clustering function - however, cannot add random effects. Probably would need multistate models: https://cran.r-project.org/web/packages/survival/vignettes/compete.pdf
ttdischarge.comp <- crr(Surv(discharge_time, discharge_reached_comp) ~ trt_centered_n
                  + cluster(trial_f)
                  # + (trt_centered_n -1 | trial_f) -1
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + clinstatus_baseline_n
                  # + (clinstatus_baseline_n -1 | trial_f)
                  ,data = df_tot)
ttdischarge_comp_reg_tbl <- tbl_regression(ttdischarge.comp, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_comp_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```
Discussion points:
1. crr has a clustering function - however, cannot add random effects. coxme cannot deal with competing risks. Needs a multistate model? https://cran.r-project.org/web/packages/survival/vignettes/compete.pdf

# (vi.i) Time to discharge or reaching discharge criteria up to day 28. Death = Hypothetical
```{r warning=FALSE}
# Censoring and assigned worst outcome (28d) to competing event (death) // hypothetical estimand where no-one died. (Another option could be, but we don't do it, is to exclude the deaths entirely, i.e. discharge among those that survived (but that might bias in favour of those in control that died more, i.e. healthier comparator))

# Cox proportional hazards model adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
ttdischarge.hypo <- coxme(Surv(discharge_time_sens, discharge_reached) ~ trt_centered_n
                  + trial_f
                  + (trt_centered_n -1 | trial_f) -1
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot)
tab_model(ttdischarge.hypo)
```

# (vi.ii) Time to discharge or reaching discharge criteria up to day 28. Death = Censored
```{r warning=FALSE}
## Censoring the deaths => Cause-specific hazards, i.e., represents the rate per unit of time of the event among those not having failed from other events. Instantaneous rate of occurrence of the given type of event in individuals who are currently event‐free. But by simply censoring the competing event, we bias in favour of comparator (if treatment leads to less deaths)

# Cox proportional hazards model adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
ttdischarge.cens <- coxme(Surv(discharge_time, discharge_reached) ~ trt_centered_n
                  + trial_f
                  + (trt_centered_n -1 | trial_f) -1
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot)
tab_model(ttdischarge.cens)
```

# (vi.iii) Time to sustained discharge or reaching discharge criteria up to day 28. Death = Censored
```{r warning=FALSE}
# Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days
# Use cause-specific hazards
# table(df_tot$discharge_reached_sus, df_tot$discharge_time_sus, useNA = "always")
# table(df_tot$discharge_reached_sus, df_tot$trial, useNA = "always")
# table(df_tot$discharge_time_sus, df_tot$trial, useNA = "always")

# Kaplan-Meier estimate of conditional discharge probability
km.ttdischarge.sus.check <- with(df_tot, Surv(discharge_time_sus, discharge_reached_sus))
# head(km.ttdischarge.sus.check, 100)
km.ttdischarge_sus_trt <- survfit(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df_tot)
# summary(km.ttdischarge_trt, times = 28)
ttdischarge_sus_28d_tbl <- km.ttdischarge_sus_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d sustained hospitalization (95% CI)**"
  )
# Nicely formatted table
kable(ttdischarge_sus_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
# KM graph
survfit2(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df_tot) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall sustained hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()

# Cox proportional hazards model adhering to main model "r cent trt, s intercept, s and cent age, r clinstatus"
ttdischarge.sus <- coxme(Surv(discharge_time_sus, discharge_reached_sus) ~ trt_centered_n
                  + trial_f
                  + (trt_centered_n -1 | trial_f) -1
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot)
tab_model(ttdischarge.sus)
```

# (vii) Viral clearance up to day 5
```{r warning=FALSE}
addmargins(table(df_tot$vir_clear_5, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$vir_clear_5, df_tot$trial, useNA = "always"))

vir.clear5 <- glmmTMB(vir_clear_5 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(vir.clear5)
```

# (viii) Viral clearance up to day 10
```{r warning=FALSE}
addmargins(table(df_tot$vir_clear_10, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$vir_clear_10, df_tot$trial, useNA = "always"))

vir.clear10 <- glmmTMB(vir_clear_10 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(vir.clear10)
```

# (ix) Viral clearance up to day 15
```{r warning=FALSE}
addmargins(table(df_tot$vir_clear_15, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$vir_clear_15, df_tot$trial, useNA = "always"))

vir.clear15 <- glmmTMB(vir_clear_15 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(vir.clear15)
```

# (x) Adverse event(s) grade 3 or 4, or a serious adverse event(s), excluding death, by day 28. ANY
```{r warning=FALSE}
addmargins(table(df_tot$ae_28, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$ae_28, df_tot$trial, useNA = "always"))

ae28 <- glmmTMB(ae_28 ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = binomial)
tab_model(ae28)
```

# (x.i) Adverse event(s) grade 3 or 4, or a serious adverse event(s), excluding death, by day 28. SEVERAL
```{r warning=FALSE}
addmargins(table(df_tot$ae_28_sev, df_tot$trt, useNA = "always"))
addmargins(table(df_tot$ae_28_sev, df_tot$trial, useNA = "always"))

ae28sev <- glmmTMB(ae_28_sev ~ trt_centered_n 
                  + trial_f 
                  + (trt_centered_n -1 | trial_f) -1 
                  + age_cent_trial_1 
                  + age_cent_trial_2
                  + age_cent_trial_3
                  + age_cent_trial_4
                  + age_cent_trial_5
                  + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f)
                  , data = df_tot, family = poisson) # double-check !
tab_model(ae28sev)
```

# Collect all treatment effect estimates across endpoints
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
result_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric(),
  n_intervention = numeric(),
  n_intervention_tot = numeric(),
  n_control = numeric(),
  n_control_tot = numeric()
)

# Function to extract treatment results from different model types
extract_trt_results <- function(model, variable_name, n_int, n_int_tot, n_cont, n_cont_tot) {
  if (inherits(model, "glmmTMB")) {
    trt_coef <- fixef(model)$cond[grep("^trt", names(fixef(model)$cond))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", rownames(confint(model))), ])
    se <- summary(model)$coefficients$cond[grep("^trt", rownames(summary(model)$coefficients$cond)), "Std. Error"]
    p_value <- summary(model)$coefficients$cond[grep("^trt", rownames(summary(model)$coefficients$cond)), "Pr(>|z|)"]
  } else if (inherits(model, "coxme")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", rownames(confint(model))), ])
    se <- NA
    p_value <- NA
  } else if (inherits(model, "clmm")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", rownames(confint(model))), ])
    se <- summary(model)$coefficients[grep("^trt", rownames(summary(model)$coefficients)), "Std. Error"]
    p_value <- summary(model)$coefficients[grep("^trt", rownames(summary(model)$coefficients)), "Pr(>|z|)"]
  } else if (inherits(model, "tidycrr")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- c(exp(model$tidy$conf.low[1]), exp(model$tidy$conf.high[1]))
    se <- model$tidy$std.error[1]
    p_value <- model$tidy$p.value[1]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_intervention = n_int,
    n_intervention_tot = n_int_tot,
    n_control = n_cont,
    n_control_tot = n_cont_tot
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_trt_results(mort28, "death at day 28",
                                        addmargins(table(df_tot$mort_28, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$mort_28, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$mort_28, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$mort_28, df_tot$trt))[3,1])
result_list[[2]] <- extract_trt_results(mort60, "death at day 60",
                                        addmargins(table(df_tot$mort_60, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$mort_60, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$mort_60, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$mort_60, df_tot$trt))[3,1])
result_list[[3]] <- extract_trt_results(ttdeath, "death within fup",
                                        addmargins(table(df_tot$death_reached, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$death_reached, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$death_reached, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$death_reached, df_tot$trt))[3,1])
result_list[[4]] <- extract_trt_results(new.mvd28, "new MV or death within 28d",
                                        addmargins(table(df_tot$new_mvd_28, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$new_mvd_28, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$new_mvd_28, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$new_mvd_28, df_tot$trt))[3,1])
result_list[[5]] <- extract_trt_results(new.mv28, "new MV within 28d",
                                        addmargins(table(df_tot$new_mv_28, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$new_mv_28, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$new_mv_28, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$new_mv_28, df_tot$trt))[3,1])
result_list[[6]] <- extract_trt_results(clin28, "clinical status at day 28",
                                        addmargins(table(df_tot$clinstatus_28_imp, df_tot$trt))[7,2], 
                                        addmargins(table(df_tot$clinstatus_28_imp, df_tot$trt))[7,2],
                                        addmargins(table(df_tot$clinstatus_28_imp, df_tot$trt))[7,1],
                                        addmargins(table(df_tot$clinstatus_28_imp, df_tot$trt))[7,1])
result_list[[7]] <- extract_trt_results(ttdischarge.comp, "discharge within 28 days, death=comp.event",
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,1])
result_list[[8]] <- extract_trt_results(ttdischarge.hypo, "discharge within 28 days, death=hypo.event",
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,1])
result_list[[9]] <- extract_trt_results(ttdischarge.cens, "discharge within 28 days, death=censored",
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$discharge_reached, df_tot$trt))[3,1])
result_list[[10]] <- extract_trt_results(ttdischarge.sus, "sustained discharge within 28 days",
                                        addmargins(table(df_tot$discharge_reached_sus, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$discharge_reached_sus, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$discharge_reached_sus, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$discharge_reached_sus, df_tot$trt))[3,1])
result_list[[11]] <- extract_trt_results(vir.clear5, "viral clearance until day 5",
                                        addmargins(table(df_tot$vir_clear_5, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$vir_clear_5, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$vir_clear_5, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$vir_clear_5, df_tot$trt))[3,1])
result_list[[12]] <- extract_trt_results(vir.clear10, "viral clearance until day 10",
                                        addmargins(table(df_tot$vir_clear_10, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$vir_clear_10, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$vir_clear_10, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$vir_clear_10, df_tot$trt))[3,1])
result_list[[13]] <- extract_trt_results(vir.clear15, "viral clearance until day 15",
                                        addmargins(table(df_tot$vir_clear_15, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$vir_clear_15, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$vir_clear_15, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$vir_clear_15, df_tot$trt))[3,1])
result_list[[14]] <- extract_trt_results(ae28, "Any AE grade 3,4 within 28 days",
                                        addmargins(table(df_tot$ae_28, df_tot$trt))[2,2], 
                                        addmargins(table(df_tot$ae_28, df_tot$trt))[3,2],
                                        addmargins(table(df_tot$ae_28, df_tot$trt))[2,1],
                                        addmargins(table(df_tot$ae_28, df_tot$trt))[3,1])
result_list[[15]] <- extract_trt_results(ae28sev, "AEs grade 3,4 within 28 days",
                                        addmargins(table(df_tot$ae_28_sev, df_tot$trt))[9,2], 
                                        addmargins(table(df_tot$ae_28_sev, df_tot$trt))[9,2],
                                        addmargins(table(df_tot$ae_28_sev, df_tot$trt))[9,1],
                                        addmargins(table(df_tot$ae_28_sev, df_tot$trt))[9,1])

# Filter out NULL results and bind the results into a single data frame
result_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the analysis approach
result_df$approach <- "one-stage"

# Add the coxme p-values (and se) manually - for now...

# Nicely formatted table
kable(result_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(result_df, file = "overall_results_one-stage.RData")
```
Discussion points

# Plot all treatment effect estimates across endpoints
```{r warning=FALSE}
# Order
result_df$variable <- factor(result_df$variable, 
                             levels = c("AEs grade 3,4 within 28 days",
                                        "Any AE grade 3,4 within 28 days",
                                        "viral clearance until day 15",
                                        "viral clearance until day 10",
                                        "viral clearance until day 5",
                                        "sustained discharge within 28 days",
                                        "discharge within 28 days, death=censored",
                                        "discharge within 28 days, death=hypo.event",
                                        "discharge within 28 days, death=comp.event",
                                        "clinical status at day 28",
                                        "new MV within 28d",
                                        "new MV or death within 28d",
                                        "death within fup",
                                        "death at day 60",
                                        "death at day 28"))
# Plotting
ggplot(result_df, aes(x = variable, y = hazard_odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.5) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "red", size = 0.5) +
  labs(title = "All endpoint results - one-stage",
       x = "Endpoints",
       y = "aOR/aHR/aIRR") +
  theme_minimal() +
  scale_y_continuous(limits = c(0.4, 1.6), breaks = seq(0.4, 1.6, 0.1)) +
  coord_flip()  # Flip axes to show longer variable names
```


# TREATMENT-COVARIATE INTERACTIONS

# Interaction: Respiratory support (proxy for disease severity) on primary endpoint
```{r}
rs.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  # + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*clinstatus_baseline_n # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(rs.mort28)

# effect by subgroup
table(df_tot$clinstatus_baseline, df_tot$mort_28, useNA = "always")
# class(df_tot$clinstatus_baseline)
mort.28.vent.rs.2 <- df_tot %>% 
  filter(clinstatus_baseline == "2") %>% # no oxygen
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.rs.2, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.vent.rs.3 <- df_tot %>% 
  filter(clinstatus_baseline == "3") %>% # LF oxygen
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.vent.rs.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.vent.rs.4 <- df_tot %>% 
  filter(clinstatus_baseline == "4") %>% # HF oxygen/NIV
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.vent.rs.4, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.vent.rs.5 <- df_tot %>% 
  filter(clinstatus_baseline == "5") %>% # ECMO
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.vent.rs.5, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points:
1. See notes and https://www.ipdma.co.uk/treatment-covariate-interactions for further explanation

# Interaction: Ventilation requirement (proxy for disease severity) on primary endpoint
```{r}
# calculate the proportion ventilated by trial
table(df_tot$vbaseline, df_tot$trial)
proportions <- df_tot %>%
  group_by(trial) %>%
  summarize(proportion_ventilated = sum(vbaseline, na.rm = TRUE) / sum(!is.na(vbaseline)))
df_tot <- left_join(df_tot, proportions[, c("proportion_ventilated", "trial")], by = join_by(trial == trial))

# create the centered vbaseline variable
df_tot$vbaseline_centered_n <- df_tot$vbaseline - df_tot$proportion_ventilated

vb.mort28 <- glmmTMB(mort_28 ~
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*vbaseline_centered_n # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(vb.mort28)

# vb.mort28 <- glmmTMB(mort_28 ~
#                   + trial_f # stratified intercept
#                   + (trt_centered_n - 1 | trial_f) - 1 # random treatment effect (and centered)
#                   + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
#                   + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
#                   + (clinstatus_baseline_n - 1 | trial_f) # random prognostic factor clinstatus_baseline within trial
#                   + (trt_centered_n - 1 | trial_f) * (vbaseline_centered_n - 1 | trial_f) # random interaction term
#                   , data = df_tot, family = binomial)
# summary(vb.mort28) ## but where is my p-value interaction now?

# # Using the common effect, random intercept model
# vb.mort28 <- glmmTMB(mort_28 ~ (1|trial) + age + trt*vbaseline
#                   , data = df_tot, family = binomial)
# summary(vb.mort28)


table(df_tot$vbaseline, df_tot$mort_28)
# effect by subgroup
mort.28.vent.vb.yes <- df_tot %>% 
  filter(vbaseline == 1) %>% # ventilated
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.vent.vb.yes, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.vent.vb.no <- df_tot %>% 
  filter(vbaseline == 0) %>% # not ventilated
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.vent.vb.no, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points

# Interaction: Age on primary endpoint
```{r}
# primary model: random and centered treatment effect, stratified intercept, stratified and centered age, random clinstatus_baseline, and the interaction term
age.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*age_centered # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(age.mort28)

# If adjustments are not centered and stratified
# age.mort28 <- glmmTMB(mort_28 ~  
#                   + trial_f # stratified intercept
#                   + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
#                   + clinstatus_baseline_n
#                   + trt_centered_n*age_centered # interaction term (common effect), centered
#                   , data = df_tot, family = binomial)
# summary(age.mort28)
# 
# # Using the common effect, random intercept model
# age.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline + trt*age
#                   , data = df_tot, family = binomial)
# summary(age.mort28)
# # Using the common effect, random intercept model, but with treatment centering (does not make the difference)
# age.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline + trt_centered_n*age_centered
#                   , data = df_tot, family = binomial)
# summary(age.mort28)

# effect by subgroup
df_tot <- df_tot %>% 
  mutate(age_70 = case_when(age < 70 ~ 0,
                            age > 69 ~ 1))
table(df_tot$age_70, useNA = "always")

mort.28.age.a70 <- df_tot %>% 
  filter(age_70 == 1) %>% # 70 and above
  glm(mort_28 ~ trt
      # + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.age.a70, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.age.b70 <- df_tot %>% 
  filter(age_70 == 0) %>% # below 70
  glm(mort_28 ~ trt
      # + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.age.b70, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```

# Interaction: Comorbidity on primary endpoint
```{r}
# Calculate the mean values of (continuous) comorb variable
# class(df_tot$comorb_cat)
comorb_means <- df_tot %>%
  group_by(trial) %>%
  summarize(mean_comorb = mean(comorb_cat, na.rm = TRUE))
# Merge back
df_tot <- df_tot %>% left_join(comorb_means, by = "trial")
# Center comorb variables
# df_tot <- df_tot %>%
#   mutate(comorb_centered = comorb_cat - mean_comorb)
# table(df_tot$comorb_cat_n)
df_tot$comorb_cat_n <- as.numeric(df_tot$comorb_cat)

comorb.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*comorb_cat_n # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(comorb.mort28)

# # If adjustments are not centered and stratified (does not make the big difference)
# comorb.mort28 <- glmmTMB(mort_28 ~
#                   + trial_f # stratified intercept
#                   + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
#                   + clinstatus_baseline_n
#                   + age
#                   + trt_centered_n*comorb_cat_n # interaction term (common effect), centered
#                   , data = df_tot, family = binomial)
# summary(comorb.mort28)
# 
# # Using the common effect, random intercept model
# comorb.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline_n + age + trt*comorb_cat_n
#                   , data = df_tot, family = binomial)
# summary(comorb.mort28)

# Comorbidity Count
comorb.count.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*comorb_count # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(comorb.count.mort28)

# Any Comorbidity
comorb.any.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*comorb_any # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(comorb.any.mort28)


# effect by subgroup
table(df_tot$comorb_cat, df_tot$mort_28, useNA = "always") 
mort.28.comorb.1 <- df_tot %>% 
  filter(comorb_cat == 1) %>% # no comorbidity
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comorb.1, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.2 <- df_tot %>% 
  filter(comorb_cat == 2) %>% # 1 comorbidity
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comorb.2, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.3 <- df_tot %>% 
  filter(comorb_cat == 3) %>% # multiple comorbidities
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comorb.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.4 <- df_tot %>% 
  filter(comorb_cat == 4) %>% # immunocompromised
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comorb.4, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Interaction: Comedication on primary endpoint
```{r}
# Calculate the mean values of (continuous) comed variable
# class(df_tot$comed_cat)
comed_means <- df_tot %>%
  group_by(trial) %>%
  summarize(mean_comed = mean(comed_cat, na.rm = TRUE))
# Merge back
df_tot <- df_tot %>% left_join(comed_means, by = "trial")
# Center comed variables
# df_tot <- df_tot %>%
#   mutate(comed_centered = comed_cat - mean_comed)
df_tot$comed_cat_n <- as.numeric(df_tot$comed_cat)

comed.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*comed_cat_n # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(comed.mort28)

# # Using the common effect, random intercept model
# comed.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline_n + age + trt*comed_cat_n
#                   , data = df_tot, family = binomial)
# summary(comed.mort28)

# effect by subgroup
table(df_tot$comed_cat, df_tot$trt, useNA = "always")
# 1: patients without Dexamethasone nor Tocilizumab => JAKi effect alone
# 2: patients with Dexamethasone and Tocilizumab => JAKi effect with Dexa + Toci
# 3: patients with Dexamethasone but no Tocilizumab => JAKi effect with Dexa only
# 4: patients with Tocilizumab but no Dexamethasone (if exist) => JAKi effect with Toci only 

mort.28.comed.1 <- df_tot %>% 
  filter(comed_cat == 1) %>% # without Dexamethasone nor Tocilizumab
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comed.1, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comed.2 <- df_tot %>%
  filter(comed_cat == 2) %>% # without Dexamethasone nor Tocilizumab
  glm(mort_28 ~ trt
      + age
      # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed.2, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comed.3 <- df_tot %>% 
  filter(comed_cat == 3) %>% # Dexamethasone but no Tocilizumab
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comed.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comed.4 <- df_tot %>% 
  filter(comed_cat == 4) %>% # Dexamethasone but no Tocilizumab
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.comed.4, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```

# Interaction: Vaccination on AEs
```{r}
# calculate the proportion vaccinated by trial
table(df_tot$vacc, df_tot$trial)
proportions <- df_tot %>%
  group_by(trial) %>%
  summarize(proportion_vacc = sum(vacc, na.rm = TRUE) / sum(!is.na(vacc)))
df_tot <- left_join(df_tot, proportions[, c("proportion_vacc", "trial")], by = join_by(trial == trial))
# create the centered vacc variable
df_tot$vacc_centered_n <- df_tot$vacc - df_tot$proportion_vacc

vacc.ae28 <- glmmTMB(ae_28 ~  
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*vacc_centered_n # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(vacc.ae28)

# # Using the common effect, random intercept model
# comed.ae28 <- glmmTMB(ae_28 ~ (1|trial) + clinstatus_baseline_n + age + trt*vacc_centered_n
#                   , data = df_tot, family = binomial)
# summary(comed.ae28)

# effect by subgroup
table(df_tot$vacc, df_tot$trt, useNA = "always") 

ae.28.vacc.1 <- df_tot %>%
  filter(vacc == 1) %>% # vaccinated
  logistf(ae_28 ~ trt
      + age
      + clinstatus_baseline
    , family = "binomial", data=.)
summary(ae.28.vacc.1)

ae.28.vacc.0 <- df_tot %>% 
  filter(vacc == 0) %>% # not vaccinated
  glm(ae_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(ae.28.vacc.0, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Interaction: Symptom onset on primary endpoint
```{r}
# Calculate the mean values of (continuous) symp variable
# class(df_tot$sympdur)
sympdur_means <- df_tot %>%
  group_by(trial) %>%
  summarize(mean_sympdur = mean(sympdur, na.rm = TRUE))
# Merge back
df_tot <- df_tot %>% left_join(sympdur_means, by = "trial")
# Center comed variables
df_tot <- df_tot %>%
  mutate(sympdur_centered = sympdur - mean_sympdur)

symp.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*sympdur_centered # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(symp.mort28)

# # Using the common effect, random intercept model
# symp.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline_n + age + trt*sympdur_centered
#                   , data = df_tot, family = binomial)
# summary(symp.mort28)

# effect by subgroup
df_tot <- df_tot %>% 
  mutate(sympdur_cat = case_when(sympdur < 6 ~ 2,
                                 sympdur > 5 & sympdur < 11 ~ 1,
                                 sympdur > 10 ~ 0))
table(df_tot$sympdur_cat, useNA = "always")
# table(df_tot$sympdur, useNA = "always")
mort.28.sympdur.a10 <- df_tot %>% 
  filter(sympdur_cat == 0) %>% # more than 10 days
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.sympdur.a10, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.sympdur.510 <- df_tot %>% 
  filter(sympdur_cat == 1) %>% # 5-10 days
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.sympdur.510, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.sympdur.b5 <- df_tot %>% 
  filter(sympdur_cat == 2) %>% # 5d or less
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.sympdur.b5, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Interaction: CRP on primary endpoint
```{r}
# Calculate the mean values of (continuous) symp variable
# class(df_tot$crp)
crp_means <- df_tot %>%
  group_by(trial) %>%
  summarize(mean_crp = mean(crp, na.rm = TRUE))
# Merge back
df_tot <- df_tot %>% left_join(crp_means, by = "trial")
# Center comed variables
df_tot <- df_tot %>%
  mutate(crp_centered = crp - mean_crp)

crp.mort28 <- glmmTMB(mort_28 ~ 
                  + trial_f # stratified intercept
                  + (trt_centered_n -1 | trial_f) -1 # random treatment effect (and centered)
                  + age_cent_trial_1 + age_cent_trial_2 + age_cent_trial_3 # stratified prognostic factor age (and centered)
                  + age_cent_trial_4 + age_cent_trial_5 + age_cent_trial_6 + age_cent_trial_7
                  + (clinstatus_baseline_n -1 | trial_f) # random prognostic factor clinstatus_baseline within trial
                  + trt_centered_n*crp_centered # interaction term (common effect), centered
                  , data = df_tot, family = binomial)
summary(crp.mort28)

# # Using the common effect, random intercept model
# crp.mort28 <- glmmTMB(mort_28 ~ (1|trial) + clinstatus_baseline_n + age + trt*crp_centered
#                   , data = df_tot, family = binomial)
# summary(crp.mort28)

# effect by subgroup
df_tot <- df_tot %>% 
  mutate(crp_75 = case_when(crp < 75 ~ 1,
                            crp > 74 ~ 0))
table(df_tot$crp_75, useNA = "always")
mort.28.crp.b75 <- df_tot %>% 
  filter(crp_75 == 1) %>% # below 75
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.crp.b75, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.crp.a75 <- df_tot %>% 
  filter(crp_75 == 0) %>% # 75 and above
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.crp.a75, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```

# Collect all interaction effect estimates
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
interaction_df <- data.frame(
  variable = character(),
  log_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric())

# Extract and format results for the interaction term
extract_interaction <- function(model, variable_name) {
  if (inherits(model, "glmmTMB")) {
    trt_coef <- fixef(model)$cond[grep("^trt_centered_n:", names(fixef(model)$cond))]
    log_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt_centered_n:", rownames(confint(model))), ])
    se <- summary(model)$coefficients$cond[grep("^trt_centered_n:", rownames(summary(model)$coefficients$cond)), "Std. Error"]
    p_value <- summary(model)$coefficients$cond[grep("^trt_centered_n:", rownames(summary(model)$coefficients$cond)), "Pr(>|z|)"]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    log_odds_ratio = round(log_odds_ratio,3),
    ci_lower = round(ci[1],3),
    ci_upper = round(ci[2],3),
    standard_error = round(se,3),
    p_value = round(p_value,3)
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_interaction(rs.mort28, "respiratory support") 
result_list[[2]] <- extract_interaction(vb.mort28, "ventilation") 
result_list[[3]] <- extract_interaction(age.mort28, "age")
result_list[[4]] <- extract_interaction(comorb.mort28, "comorbidity") 
result_list[[5]] <- extract_interaction(comorb.count.mort28, "comorbidity count") 
result_list[[6]] <- extract_interaction(comorb.any.mort28, "any comorbidity") 
result_list[[7]] <- extract_interaction(comed.mort28, "comedication")
result_list[[8]] <- extract_interaction(vacc.ae28, "vaccination on AEs") 
result_list[[9]] <- extract_interaction(symp.mort28, "symptom duration") 
result_list[[10]] <- extract_interaction(crp.mort28, "crp") 

# Filter out NULL results and bind the results into a single data frame
interaction_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the analysis approach
interaction_df$approach <- "one-stage"

# Nicely formatted table
kable(interaction_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(interaction_df, file = "int_effects_one-stage.RData")
```

# Collect all subgroup treatment effect estimates
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
subgroup_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric(),
  n_intervention = numeric(),
  n_intervention_tot = numeric(),
  n_control = numeric(),
  n_control_tot = numeric()
)

# Function to extract subgroup treatment results
extract_subgroup_results <- function(model, variable_name, n_int, n_int_tot, n_cont, n_cont_tot) {
  if (inherits(model, "glm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "logistf")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", names(coef(model))), ])
    se <- sqrt(diag(vcov(model)))[grep("^trt", names(coef(model)))]
    p_value <- model$prob[grep("^trt", names(coef(model)))]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_intervention = n_int,
    n_intervention_tot = n_int_tot,
    n_control = n_cont,
    n_control_tot = n_cont_tot
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_subgroup_results(mort.28.vent.vb.yes, "High-flow or non-invasive, mechanical, or ECMO",
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[2,3,1]) 
result_list[[2]] <- extract_subgroup_results(mort.28.vent.vb.no, "None or low-flow oxygen",
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[1,2,2], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[1,3,2], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[1,2,1], 
                                             addmargins(table(df_tot$vbaseline, df_tot$mort_28, df_tot$trt))[1,3,1]) 
result_list[[3]] <- extract_subgroup_results(mort.28.vent.rs.2, "No oxygen",
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[2,3,1])
result_list[[4]] <- extract_subgroup_results(mort.28.vent.rs.3, "low-flow oxygen",
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[3,2,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[3,3,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[3,2,1], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[3,3,1])
result_list[[5]] <- extract_subgroup_results(mort.28.vent.rs.4, "high-flow oxygen / NIV",
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[4,2,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[4,3,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[4,2,1], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[4,3,1])
result_list[[6]] <- extract_subgroup_results(mort.28.vent.rs.5, "Mechanical ventilation / ECMO",
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[5,2,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[5,3,2], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[5,2,1], 
                                             addmargins(table(df_tot$clinstatus_baseline, df_tot$mort_28, df_tot$trt))[5,3,1]) 
result_list[[7]] <- extract_subgroup_results(mort.28.age.a70, "70 years and above",
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[2,3,1]) 
result_list[[8]] <- extract_subgroup_results(mort.28.age.b70, "below 70 years",
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[1,2,2], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[1,3,2], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[1,2,1], 
                                             addmargins(table(df_tot$age_70, df_tot$mort_28, df_tot$trt))[1,3,1]) 
result_list[[9]] <- extract_subgroup_results(mort.28.comorb.1, "No comorbidity",
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[1,2,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[1,3,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[1,2,1], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[1,3,1])
result_list[[10]] <- extract_subgroup_results(mort.28.comorb.2, "One comorbidity",
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[2,3,1])
result_list[[11]] <- extract_subgroup_results(mort.28.comorb.3, "Multiple comorbidities",
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[3,2,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[3,3,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[3,2,1], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[3,3,1])
result_list[[12]] <- extract_subgroup_results(mort.28.comorb.4, "Immunocompromised",
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[4,2,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[4,3,2], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[4,2,1], 
                                             addmargins(table(df_tot$comorb_cat, df_tot$mort_28, df_tot$trt))[4,3,1]) 
result_list[[13]] <- extract_subgroup_results(mort.28.comed.1, "No Dexa, no Tocilizumab",
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[1,2,2],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[1,3,2],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[1,2,1],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[1,3,1])
result_list[[14]] <- extract_subgroup_results(mort.28.comed.2, "Dexa and Tocilizumab",
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[2,2,2],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[2,3,2],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[2,2,1],
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[2,3,1])
result_list[[15]] <- extract_subgroup_results(mort.28.comed.3, "Dexa, but no Tocilizumab",
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[3,2,2], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[3,3,2], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[3,2,1], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[3,3,1])
result_list[[16]] <- extract_subgroup_results(mort.28.comed.4, "Tocilizumab, but no Dexa",
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[4,2,2], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[4,3,2], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[4,2,1], 
                                             addmargins(table(df_tot$comed_cat, df_tot$mort_28, df_tot$trt))[4,3,1])
result_list[[17]] <- extract_subgroup_results(ae.28.vacc.1, "vaccinated_firth",
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[2,2,2],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[2,3,2],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[2,2,1],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[2,3,1])
result_list[[18]] <- extract_subgroup_results(ae.28.vacc.0, "not vaccinated",
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[1,2,2],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[1,3,2],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[1,2,1],
                                             addmargins(table(df_tot$vacc, df_tot$mort_28, df_tot$trt))[1,3,1])
result_list[[19]] <- extract_subgroup_results(mort.28.sympdur.a10, "More than 10 days",
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[1,2,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[1,3,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[1,2,1], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[1,3,1])
result_list[[20]] <- extract_subgroup_results(mort.28.sympdur.510, "Between 5-10 days",
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[2,3,1])
result_list[[21]] <- extract_subgroup_results(mort.28.sympdur.b5, "5 days and less",
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[3,2,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[3,3,2], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[3,2,1], 
                                             addmargins(table(df_tot$sympdur_cat, df_tot$mort_28, df_tot$trt))[3,3,1])
result_list[[22]] <- extract_subgroup_results(mort.28.crp.a75, "CRP 75 and higher",
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[1,2,2], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[1,3,2], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[1,2,1], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[1,3,1])
result_list[[23]] <- extract_subgroup_results(mort.28.crp.b75, "CRP below 75",
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[2,2,2], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[2,3,2], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[2,2,1], 
                                             addmargins(table(df_tot$crp_75, df_tot$mort_28, df_tot$trt))[2,3,1])

# Filter out NULL results and bind the results into a single data frame
subgroup_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Nicely formatted table
kable(subgroup_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(subgroup_df, file = "subgroup_effects_one-stage.RData")
```
Discussion points

# Forestplot subgroup - on primary endpoint
```{r, fig.width=16, fig.height=8}
# str(subgroup_df)
# foresplot 
df_subgroup_mort28 <- subgroup_df %>% 
  filter(variable != "vaccinated_firth" & variable != "not vaccinated")

df_subgroup_mort28$inverse_variance <- 1 / df_subgroup_mort28$standard_error^2
events_i <- df_subgroup_mort28$n_intervention
tot_i <- df_subgroup_mort28$n_intervention_tot
events_c <- df_subgroup_mort28$n_control
tot_c <- df_subgroup_mort28$n_control_tot
# p_int <- df_subgroup_mort28$p_value
# p_int <- round(p_int,2)
p_int <- c("0", "", "0", "", "", "", "0", "", "0", "", "", "", "0", "", "","","0", "", "", "0","") # from the two-stage models

base_data <- tibble(mean = df_subgroup_mort28$hazard_odds_ratio,
                    lower = df_subgroup_mort28$ci_lower,
                    upper = df_subgroup_mort28$ci_upper,
                    subgroup = c("Ventilation at baseline: High-flow or non-invasive ventilation, mechanical ventilation, ECMO",
                                 "No ventilation at baseline: No or only low-flow oxygen",
                                 "No oxygen",
                                 "Only low-flow oxygen",
                                 "High-flow or non-invasive ventilation",
                                 "Mechanical ventilation or ECMO",
                                 "70 years of age or older",
                                 "Below 70 years of age",
                                 "No comorbidity",
                                 "One comorbidity",
                                 "Multiple comorbidities",
                                 "Immunocompromised",
                                 "No Dexamethasone, no Tocilizumab",
                                 "Dexamethasone and Tocilizumab",
                                 "Dexamethasone, but no Tocilizumab",
                                 "Tocilizumab, but no Dexamethasone",
                                 "Enrolment more than 10 days after symptom onset",
                                 "Enrolment between 5 and 10 days after symptom onset",
                                 "Enrolment 5 days or earlier after symptom onset",
                                 "C-reactive protein 75mg/L or more",
                                 "C-reactive protein less than 75mg/L"),
                    events_i = as.character(events_i),
                    tot_i = as.character(tot_i),
                    events_c = as.character(events_c),
                    tot_c = as.character(tot_c),
                    p_int = as.character(p_int))
summary <- tibble(mean  = 0.69,
                  lower = 0.54,
                  upper = 0.88,
                  subgroup = "Overall treatment effect",
                  summary = TRUE)
header <- tibble(subgroup = c("Treatment effect on 28-day mortality, by subgroup"),
                 events_i = c("Events int."),
                 tot_i = c("No. int."),
                 events_c = c("Events cont."),
                 tot_c = c("No. cont."),
                 p_int = c("p-int*"),
                 summary = TRUE)
mort28_fp <- bind_rows(header,base_data,summary)
font <- "sans"

# Open a pdf file
# pdf("mort28_fp.pdf", width=16, height=8)
mort28_fp %>%
  forestplot(labeltext = c(subgroup, events_i, tot_i, events_c, tot_c, p_int),
             txt_gp = fpTxtGp(label = gpar(fontfamily = font, cex=1),
                              ticks = gpar(cex=0.88),
                              summary = gpar(cex=1),
                              xlab = gpar(cex=0.88)),
             # title = "Treatment effect on mortality at day 28 by subgroup",
             is.summary = summary,
             graph.pos = 6,
             clip = c(0.1, 2),
             hrzl_lines = list("2" = gpar(lty = 2),
                               "4" = gpar(lty = 2),
                               "8" = gpar(lty = 2),
                               "10" = gpar(lty = 2),
                               "14" = gpar(lty = 2),
                               "18" = gpar(lty = 2),
                               "21" = gpar(lty = 2),
                               "23" = gpar(lty = 2)),
             xlog = FALSE,
             xticks = c(0,0.25,0.5,0.75,1,1.25,1.5),
             psize = sqrt(df_subgroup_mort28$inverse_variance),
             lty.ci = c(1),
             col = fpColors(box = "maroon4",
                            line = "maroon1",
                            summary = "magenta4",
                            hrz_lines = "gray63"),
             vertices = TRUE,
             xlab = "     Favours JAKi <-> Favours No JAKi",
             zero = 1,
             grid = structure(c(0.69), gp = gpar(lty = 2, col = "gray63")),
             graphwidth = unit(100, "mm"), colgap = unit(2.5, "mm")
             )

# x <- unit(0.3, 'npc')
# y <- unit(0.1, 'npc')
# grid.text('', x, y, gp = gpar(fontsize=9, font = 3))

# dev.off()


```
* The p-values for the interaction were obtained using a two-stage IPDMA approach, i.e. solely based on within-trial interactions ("deft"): 
First, to produce a treatment-covariate interaction estimate and its variance, a binomial regression was fitted in each trial separately, adjusted (where appropriate) for respiratory support and age, including the treatment and the treatment-covariate interaction, using restricted maximum likelihood estimation (with Firth penalisation correction in case of sparse data). 
Second, the interaction estimates were combined across trials in a random-effect model (the true interactions are assumed random across trials), using restricted maximum likelihood estimation and the confidence interval for the summary interaction derived using the Hartung-Knapp Sidik-Jonkman approach. 
Sizing of all squares are in proportion to the inverse of the variance of the estimates. 
For continuous covariates (age, symptom duration, CRP), a cut-off was chosen for descriptive purpose, but they were included as a continuous treatment-covariate interaction, assuming linearity. 
Similarly, ordinal covariates (detailed respiratory support, comorbidities) were included as a continuous treatment-covariate interaction, assuming linearity. 
