---
title: "recovery"
author: "A.Amstutz"
date: "2024-01-28"
output: 
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)
library(haven) # Read sas files
library(here)
library(kableExtra)

library(jtools) # for summ() and plot_summs
library(sjPlot) # for tab_model
library(ggplot2) # survival/TTE analyses and other graphs
library(ggsurvfit) # survival/TTE analyses
library(survival) # survival/TTE analyses
library(gtsummary) # survival/TTE analyses
library(ggfortify) # autoplot
library(tidycmprsk) # competing risk analysis
library(ordinal) # clinstatus ordinal regression
library(logistf) # Firth regression in case of rare events

library(finalfit) # missing data exploration
library(mice) # multiple imputation
library(jomo) # multiple imputation
library(mitools) # multiple imputation

# IDDO tool
# devtools::install_github("RhysPeploe/iddoverse")
library(iddoverse)
```

# Load Data
```{r, include=FALSE}
df_dm <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-01-24/DM 2024-01-11.csv")
df_in <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-01-24/IN 2024-01-11.csv")
df_lab <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-01-24/LB 2024-01-11.csv")
df_mb <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-01-24/MB 2024-01-11.csv")
df_sa <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-01-24/SA 2024-01-11.csv")
df_ds <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-02-06/DS 2024-02-06.csv")
df_ho <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-02-06/HO 2024-02-06.csv")
df_sv <- read_csv("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/RECOVERY/DATA_2024-02-06/SV 2024-02-06.csv")

```

# Define ITT population
```{r echo=TRUE}
df <- df_dm
addmargins(table(df$ARMCD, useNA = "always"))
```

# Baseline Characteristics
```{r echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>% # no missing
  rename(id_pat = USUBJID,
         country = COUNTRY
         )
df <- df %>%
  mutate(trt = case_when(ARMCD == "DB-BARI" ~ 1,
                         TRUE ~ 0))
# add trial variables
df$trial <- c("RECOVERY")
df$JAKi <- c("Baricitinib")

# sex
df <- df %>% # no missing in sex
  mutate(sex = case_when(SEX == "F" ~ "female",
                         SEX == "M" ~ "male"))
# Ethnicity
df$ethn <- df$RACE

# AGE
df$age <- as.numeric(df$AGE)
# df <- df %>% # exclude the kids (see protocol: 16y or older)
#   filter(age>15)
df %>% 
  drop_na(age) %>% 
  ggplot(aes(x = age)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Age",
       x = "Age",
       y = "Density")

# id_pat
df_in <- df_in %>% 
  rename(id_pat = USUBJID)

### Clinical score at baseline
df_baseline <- df_in %>% # get all respiratory support information for baseline
  filter(INDECOD %in% c("EXTRA-CORPOREAL MEMBRANE OXYGENATION", "NON-INVASIVE VENTILATION", "HEATED AND HUMIDIFIED HIGH FLOW OXYGEN THERAPY USING NASAL CANNULA", "ASSISTED BREATHING
         INTUBATION OF RESPIRATORY TRACT", "CONTINUOUS POSITIVE AIRWAY PRESSURE VENTILATION TREATMENT", "OXYGEN THERAPY", "INVASIVE MECHANICAL VENTILATION") |
           INMODIFY %in% c("CPAP", "OXYGEN THERAPY", "ECMO", "INVASIVE MECHANICAL VENTILATION", "NON-INVASIVE VENTILATION") |
           INTRT %in% c("Mechanical ventilation (tube/tracheo)", "Extra-corporeal membrane oxygenation", "Non-invasive ventilation (e.g. BiPAP)", "CPAP ventilation", "Requires non-invasive ventilation", "Requires inv. mech. ventilation or ECMO", "Requires oxygen", "High-flow nasal oxygen (e.g. AIRVO)", "Any assisted ventilation including IMV, NIV, ECMO", "Invasive mechanical ventilation or ECMO", "Non-invasive ventilation", "Other respiratory support", "New use of invasive mech. ventilation", "New use or inc. concentration of oxygen", "New use of non-invasive resp. support")) %>% 
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")

# recode into our score
df_baseline <- df_baseline %>% 
  mutate(clinstatus_baseline_all = case_when(INDECOD == "INTUBATION OF RESPIRATORY TRACT" | 
                                           INTRT == "Mechanical ventilation (tube/tracheo)" | 
                                           INDECOD == "EXTRA-CORPOREAL MEMBRANE OXYGENATION" | 
                                           INTRT == "Extra-corporeal membrane oxygenation" |
                                           INTRT == "Requires inv. mech. ventilation or ECMO" ~ 5,
                                         INDECOD == "NON-INVASIVE VENTILATION" | 
                                           INDECOD == "CONTINUOUS POSITIVE AIRWAY PRESSURE VENTILATION TREATMENT" | 
                                           INTRT == "CPAP ventilation" | 
                                           INTRT == "Requires non-invasive ventilation" | 
                                           INTRT == "High-flow nasal oxygen (e.g. AIRVO)" | 
                                           INTRT == "Non-invasive ventilation (e.g. BiPAP)" ~ 4,
                                         INDECOD == "HEATED AND HUMIDIFIED HIGH FLOW OXYGEN THERAPY USING NASAL CANNULA" | 
                                           INDECOD == "OXYGEN THERAPY" |
                                           INTRT == "Requires oxygen" ~ 3))
# sort by highest value within each patient 
df_baseline <- df_baseline %>%
  arrange(id_pat, desc(clinstatus_baseline_all))
# select only highest values
df_baseline <- df_baseline %>%
  group_by(id_pat) %>%
  slice_max(order_by = clinstatus_baseline_all) %>%
  ungroup()
# only keep one row per patient
df_baseline <- df_baseline %>%
  distinct(id_pat, clinstatus_baseline_all)
# merge
df <- left_join(df, df_baseline[, c("clinstatus_baseline_all", "id_pat")], by = join_by(id_pat == id_pat)) 

# table(df$clinstatus_baseline_all, df$trt, useNA = "always") ## corresponds to trial publication! 
df <- df %>% 
  mutate(clinstatus_baseline = case_when(is.na(clinstatus_baseline_all) ~ 2,
                                         TRUE ~ clinstatus_baseline_all))
# table(df$clinstatus_baseline, df$trt, useNA = "always") ## And the NAs are None (no O2 support)

### Vaccination at baseline
df_vacc <- df_in %>%
  filter(INDECOD %in% c("COVID-19 VACCINES") |
           INMODIFY %in% c("COVID-19 VACCINATION") |
           INTRT %in% c("COVID-19 vaccine")) %>% 
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
# df_vacc <- df_vacc %>%
#   distinct(id_pat) # no duplicates
df_vacc$vacc <- 1
# merge
df <- left_join(df, df_vacc[, c("vacc", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(vacc = case_when(is.na(vacc) ~ 0,
                          TRUE ~ vacc))
# table(df$vacc, df$trt, useNA = "always") ## corresponds to trial publication! 

### Co-medication at baseline
# rdv
df_rdv <- df_in %>% 
  filter(INDECOD %in% c("REMDESIVIR") |
           INMODIFY %in% c("REMDESIVIR") |
           INTRT %in% c("Remdesivir")) %>% 
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
# df_rdv <- df_rdv %>%
#   distinct(id_pat) # no duplicates
df_rdv$comed_rdv <- 1
# merge
df <- left_join(df, df_rdv[, c("comed_rdv", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_rdv = case_when(is.na(comed_rdv) ~ 0,
                          TRUE ~ comed_rdv))
# table(df$comed_rdv, df$trt, useNA = "always") ## corresponds to trial publication!

# toci, incl. planned to be given
df_toci <- df_in %>% 
  filter(INDECOD %in% c("TOCILIZUMAB", "INTERLEUKIN INHIBITORS") |
           INTRT %in% c("Tocilizumab planned to be given in next 24h", "Tocilizumab", "Tocilizumab or sarilumab")) %>%
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
# df_in %>% # these codes were only given during follow-up!
#   filter(INTRT == "Tocilizumab or sarilumab" | INDECOD == "INTERLEUKIN INHIBITORS") %>%
#   filter(VISIT == "RANDOMIZATION") %>% 
#   View()
# df_toci <- df_toci %>%
#   distinct(id_pat) # no duplicates
df_toci$comed_toci <- 1
# merge
df <- left_join(df, df_toci[, c("comed_toci", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_toci = case_when(is.na(comed_toci) ~ 0,
                          TRUE ~ comed_toci))
# table(df$comed_toci, df$trt, useNA = "always") ## corresponds to trial publication!

# cortico
df_dexa <- df_in %>% 
  filter(INDECOD %in% c("GLUCOCORTICOIDS") |
           INMODIFY %in% c("CORTICOSTEROIDS") |
           INTRT %in% c("Systemic corticosteroids (e.g. dexamethasone, prednisolone, hydrocorticone, methylprednisolone)", "Corticosteroid (dexamethasone, prednisolone, hydrocortisone, methylprednisolone)")) %>%
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
# df_in %>% # these codes were only given during follow-up!
#   filter(INTRT == "Corticosteroid (dexamethasone, prednisolone, hydrocortisone, methylprednisolone)") %>%
#   filter(VISIT == "RANDOMIZATION") %>%
#   View()
df_dexa <- df_dexa %>%
  distinct(id_pat) # no duplicates
df_dexa$comed_dexa <- 1
# merge
df <- left_join(df, df_dexa[, c("comed_dexa", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_dexa = case_when(is.na(comed_dexa) ~ 0,
                          TRUE ~ comed_dexa))
# table(df$comed_dexa, df$trt, useNA = "always") ## corresponds to trial publication!

# anticoa
df_acoa <- df_in %>% 
  filter(INDECOD %in% c("HEPARIN GROUP") |
           INTRT %in% c("Warfarin or direct oral anticoagulant","Higher dose low molecular weight heparin")) %>%
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
df_acoa <- df_acoa %>%
  distinct(id_pat) # remove duplicates
df_acoa$comed_acoa <- 1
# merge
df <- left_join(df, df_acoa[, c("comed_acoa", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_acoa = case_when(is.na(comed_acoa) ~ 0,
                          TRUE ~ comed_acoa))

# interferon
df$comed_interferon <- 0 # see protocol

# antibiotics
df_ab <- df_in %>% 
  filter(INDECOD %in% c("MACROLIDES", "AZITHROMYCIN, UNSPECIFIED FORM") |
           INMODIFY %in% c("AZITHROMYCIN") |
           INTRT %in% c("Macrolide", "Other macrolides", "Azithromycin or other macrolides", "Azithromycin")) %>%
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
# df_ab <- df_ab %>%
#   distinct(id_pat) # no duplicates
df_ab$comed_ab <- 1
# merge
df <- left_join(df, df_ab[, c("comed_ab", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_ab = case_when(is.na(comed_ab) ~ 0,
                          TRUE ~ comed_ab))

# comed_other
df_comed_other <- df_in %>% # get all comedication
  filter(INDECOD %in% c("ANTITHROMBOTIC AGENTS", "CASIRIVIMAB AND IMDEVIMAB", 
                        "COLCHICINE", "TRANSFUSION OF CONVALESCENT PLASMA", "HYDROXYCHLOROQUINE", "IMMUNOGLOBULINS",
                        "LOPINAVIR_RITONAVIR", "ASPIRIN") |
           INMODIFY %in% c("CONVALESCENT PLASMA", "IMMUNOGLOBULINS", "LOPINAVIR AND RITONAVIR",
                           "ACETYLSALICYLIC ACID", "COLCHICINE", 
                           "HYDROXYCHLOROQUINE", "BLOOD TRANSFUSION OR BLOOD PRODUCT") |
           INTRT %in% c("Antiplatelet therapy", "Venous thromboembolism prophylaxis",
                        "Standard low molecular weight heparin", "Synthetic monoclonal antibodies (REGN10933+REGN10987)", 
                        "Colchicine", "Convalescent plasma infusion",
                        "Hydroxychloroquine",
                        "Intravenous immunoglobulin", "Vasoactive drug", "Blood transfusion", "Lopinavir-ritonavir",
                        "Aspirin")) %>% 
  filter(VISIT == "RANDOMIZATION") %>% 
  filter(INOCCUR == "Y")
df_comed_other <- df_comed_other %>%
  distinct(id_pat) # remove duplicates
df_comed_other$comed_other <- 1
# merge
df <- left_join(df, df_comed_other[, c("comed_other", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comed_other = case_when(is.na(comed_other) ~ 0,
                          TRUE ~ comed_other))

# group them for the subgroup analysis, according to protocol
df <- df %>% 
  mutate(comed_cat = case_when(comed_dexa == 0 & (comed_toci == 0 | is.na(comed_toci)) ~ 1, # patients without Dexa nor Toci
                               comed_dexa == 1 & comed_toci == 1 ~ 2, # patients with Dexa and Toci
                               comed_dexa == 1 & (comed_toci == 0 | is.na(comed_toci)) ~ 3, # patients with Dexa but no Toci
                               comed_dexa == 0 & comed_toci == 1 ~ 4)) # patients with Toci but no Dexa
table(df$comed_cat, df$trt, useNA = "always")


### Comorbidity at baseline, including immunocompromised
# id_pat
df_sa <- df_sa %>% 
  rename(id_pat = USUBJID)

df_comorb <- df_sa %>% 
  filter(SACAT == "MEDICAL HISTORY" | is.na(SACAT)) %>% 
  filter(SAOCCUR == "Y")

# diabetes
df_dm <- df_comorb %>% 
  filter(SATERM %in% c("Diabetes") |
           SAMODIFY %in% c("DIABETES MELLITUS - TYPE NOT SPECIFIED") |
           SADECOD %in% c("Diabetes mellitus"))
# df_dm <- df_dm %>%
#   distinct(id_pat) # no duplicates
df_dm$comorb_dm <- 1
# merge
df <- left_join(df, df_dm[, c("comorb_dm", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comorb_dm = case_when(is.na(comorb_dm) ~ 0,
                          TRUE ~ comorb_dm))
# table(df$comorb_dm, df$trt, useNA = "always") ## corresponds to trial publication!

# heart disease (except hypertension)
df_cvd <- df_comorb %>% 
  filter(SATERM %in% c("Heart Disease") |
           SAMODIFY %in% c("CHRONIC CARDIAC DISEASE (NOT HYPERTENSION)") |
           SADECOD %in% c("Heart disease"))
# df_cvd <- df_cvd %>%
#   distinct(id_pat) # no duplicates
df_cvd$comorb_cvd <- 1
# merge
df <- left_join(df, df_cvd[, c("comorb_cvd", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comorb_cvd = case_when(is.na(comorb_cvd) ~ 0,
                          TRUE ~ comorb_cvd))
# table(df$comorb_cvd, df$trt, useNA = "always") ## corresponds to trial publication!

# lung disease
df_lung <- df_comorb %>% 
  filter(SATERM %in% c("Lung disease") |
           SAMODIFY %in% c("CHRONIC PULMONARY DISEASE (NOT ASTHMA)") |
           SADECOD %in% c("Disorder of lung"))
# df_lung <- df_lung %>%
#   distinct(id_pat) # no duplicates
df_lung$comorb_lung <- 1
# merge
df <- left_join(df, df_lung[, c("comorb_lung", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comorb_lung = case_when(is.na(comorb_lung) ~ 0,
                          TRUE ~ comorb_lung))
# table(df$comorb_lung, df$trt, useNA = "always") ## corresponds to trial publication!

# kidney disease
df_kidney <- df_comorb %>% 
  filter(SATERM %in% c("Severe kidney impairment (eGFR<30 or on dialysis)") |
           SAMODIFY %in% c("CHRONIC KIDNEY DISEASE") |
           SADECOD %in% c("Chronic kidney disease"))
# df_kidney <- df_kidney %>%
#   distinct(id_pat) # no duplicates
df_kidney$comorb_kidney <- 1
# merge
df <- left_join(df, df_kidney[, c("comorb_kidney", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comorb_kidney = case_when(is.na(comorb_kidney) ~ 0,
                          TRUE ~ comorb_kidney))
# table(df$comorb_kidney, df$trt, useNA = "always") ## corresponds to trial publication!

# liver disease
df_liver <- df_comorb %>% 
  filter(SATERM %in% c("Severe liver disease") |
           SAMODIFY %in% c("MODERATE OR SEVERE LIVER DISEASE") |
           SADECOD %in% c("Disorder of liver"))
# df_liver <- df_liver %>%
#   distinct(id_pat) # no duplicates
df_liver$comorb_liver <- 1
# merge
df <- left_join(df, df_liver[, c("comorb_liver", "id_pat")], by = join_by(id_pat == id_pat)) 
df <- df %>% 
  mutate(comorb_liver = case_when(is.na(comorb_liver) ~ 0,
                          TRUE ~ comorb_liver))
# table(df$comorb_liver, df$trt, useNA = "always") ## corresponds to trial publication!

df$comorb_obese <- NA # no info 
df$comorb_aht <- NA # no info 
df$comorb_cancer <- NA # no info 
df$comorb_autoimm <- NA # no info 
df$comorb_smoker <- NA # no info 
df$immunosupp <- NA # there is HIV, but no info re viral suppression / CD4

## any of the above 5
df <- df %>% 
  mutate(any_comorb = case_when(comorb_lung == 1 | comorb_liver == 1 | comorb_cvd == 1 |
                                  comorb_dm == 1 | comorb_kidney == 1 
                                  ~ 1,
                                comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_dm == 0 & comorb_kidney == 0
                                ~ 0))
addmargins(table(df$any_comorb, df$trt, useNA = "always"))

## group them for the subgroup analysis, according to protocol // count all pre-defined comorbidities per patient first
comorb <- df %>% 
  select(id_pat, comorb_lung, comorb_liver, comorb_cvd, comorb_dm, comorb_kidney)
comorb$comorb_count <- NA
for (i in 1:dim(comorb)[[1]]) {
  comorb$comorb_count[i] <- ifelse(
    sum(comorb[i, ] %in% c(1)) > 0,
    sum(comorb[i, ] %in% c(1)),
    NA
  )
}
comorb <- comorb %>%
  mutate(comorb_count = case_when(comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_dm == 0 & comorb_kidney == 0 ~ 0,
                                TRUE ~ comorb_count))
df <- left_join(df, comorb[, c("comorb_count", "id_pat")], by = join_by(id_pat == id_pat)) ## merge imputed variable back
df <- df %>%
  mutate(comorb_cat = case_when(comorb_count == 0 ~ 1, # no comorbidity
                                comorb_count == 1 ~ 2, # one comorbidity
                                comorb_count >1 ~ 3)) # multiple comorbidities
addmargins(table(df$comorb_cat, df$trt, useNA = "always"))


### Days with symptoms prior to randomization
df_symp <- df_sa %>% 
  filter(SATERM == "COVID-19 Symptoms" | SAMODIFY == "COVID-19 SYMPTOMS") %>% 
  filter(SAOCCUR == "Y")
# df_symp <- df_symp %>%
#   distinct(id_pat) # no duplicates
df_symp$sympdur <- df_symp$SASTDY
summary(df_symp$sympdur)
df <- left_join(df, df_symp[, c("sympdur", "id_pat")], by = join_by(id_pat == id_pat)) 
df %>% 
  drop_na(sympdur) %>% 
  ggplot(aes(x = sympdur)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "Symptom Duration",
       y = "Density")
# df <- df %>% 
#   mutate(sympdur2 = case_when(sympdur > 0 ~ NA,
#                              TRUE ~ sympdur))
# summary(df$sympdur2)
# df %>% 
#   filter(trt == 0) %>% 
#   summary()
# summary(df$sympdur, df$trt)
# unique(df_symp$SADY)

### CRP at baseline
# id_pat
df_lab <- df_lab %>% 
  rename(id_pat = USUBJID)

df_crp <- df_lab %>% # these are all at randomization and OCCUR == Yes
  filter(LBTESTCD == "CRP")
df_crp <- df_crp %>% 
  rename(crp = LBSTRESN)
# summary(df_crp$crp) # 7 missing
# df_crp <- df_crp %>%
#   distinct(id_pat) # no duplicates
df <- left_join(df, df_crp[, c("crp", "id_pat")], by = join_by(id_pat == id_pat)) ## merge to main df
df %>% 
  drop_na(crp) %>% 
  ggplot(aes(x = crp)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "CRP",
       y = "Density")


### Viremia / Undetectable VL at baseline
# id_pat
df_mb <- df_mb %>% 
  rename(id_pat = USUBJID)

df_vl_baseline <- df_mb %>% 
  filter(MBTESTCD == "SARSCOV2") %>% # viral load
  filter(VISIT == "RANDOMIZATION" | is.na(VISIT)) %>% # exclude follow-up
  filter(EPOCH == "BASELINE" | is.na(EPOCH)) # exclude follow-up

df_vl_baseline <- df_vl_baseline %>% # keep (within each patient) only value closest to baseline
  group_by(id_pat) %>%
  slice_max(order_by = MBDY) %>%
  ungroup()

df_vl_baseline <- df_vl_baseline %>% # viral load value <LOQ and/or undectectable
  mutate(vl_baseline = case_when(MBSTRESC == "POSITIVE" ~ 0,
                                 MBSTRESC == "NEGATIVE" ~ 1)) %>% 
  filter(vl_baseline == 0 | vl_baseline == 1) # remove NAs, to avoid duplicates when merging
df_vl_baseline <- df_vl_baseline %>%
  distinct(id_pat, vl_baseline) # remove duplicates with same result
# df_vl_baseline$duplicate <- duplicated(df_vl_baseline$id_pat)
# df_vl_baseline %>% # this person has pos and neg result at the same day, closest to baseline day.
#   filter(id_pat == 2261) %>% 
#   View()
df_vl_baseline <- df_vl_baseline %>% # Keep neg since had already a neg before
  filter(!(id_pat == 2261 & vl_baseline == 0))

df <- left_join(df, df_vl_baseline[, c("vl_baseline", "id_pat")], by = join_by(id_pat == id_pat)) ## merge to main df
# table(df$vl_baseline, df$trt, useNA = "always") # resolve


### Serology at baseline
df_sero <- df_mb %>% 
  filter(MBTESTCD == "SAR2IGG") %>% # viral load
  filter(VISIT == "RANDOMIZATION" | is.na(VISIT)) %>% # exclude follow-up
  filter(EPOCH == "BASELINE" | is.na(EPOCH)) # exclude follow-up

df_sero <- df_sero %>% 
  mutate(sero = case_when(MBSTRESC == "POSITIVE" ~ 1,
                          MBSTRESC == "NEGATIVE" ~ 0))
# df_sero <- df_sero %>%
#   distinct(id_pat, sero) # no duplicates

df <- left_join(df, df_sero[, c("sero", "id_pat")], by = join_by(id_pat == id_pat)) ## merge to main df


### Variant // not available
```
Discussion points:
* Symptom duration?
* ICU at baseline? (although not in publication)

* Resolve discrepancy in viral load

* What was the maximum systematic follow-up of all study participants? According to protocol: 6months.
* Deaths after discharge systematically documented? Probably not.
* Re-admissions systematically documented? Probably not.
* Any transfer to hospice occurred? Probably yes.

* Exclude the children


# Endpoints
```{r echo=TRUE}
# mort_28_dimp
df <- df %>% 
  mutate(mort_28 = case_when(DTHFL == "Y" ~ 1,
                             TRUE ~ 0))
table(df$mort_28, df$trt)



# First, keep mort_28 as complete case

# Second, use multiple imputation (see below)

# Third, apply a deterministic imputation (see notes): we use the same rules as ACTT2 => transfer that died are already accounted for, for the remaining -> assign "alive"
df <- df %>%
  mutate(mort_28_dimp = case_when(is.na(mort_28) ~ 0,
                             TRUE ~ c(mort_28)))

# (ii) Mortality at day 60
# (iii) Time to death within max. follow-up time
# (iv) New mechanical ventilation among survivors within 28 days
# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days
# (v) Clinical status at day 28
# (vi) Time to discharge or reaching discharge criteria up to day 28
# (vi) Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days.
# (vii) Viral clearance up to day 5, day 10, and day 15 (Viral load value <LOQ and/or undectectable)
# (viii) Quality of life at day 28: Not available in RECOVERY
# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28
# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
# (ix) Sens-analysis: Alternative definition/analysis of outcome: time to first (of these) adverse event, within 28 days, considering death as a competing risk (=> censor and set to 28 days)
# re-discuss
# (x) Adverse events of special interest within 28 days: a) thromboembolic events (venous thromboembolism, pulmonary embolism, arterial thrombosis), b) secondary infections (bacterial pneumonia including ventilator-associated pneumonia, meningitis and encephalitis, endocarditis and bacteremia, invasive fungal infection including pulmonary aspergillosis), c) Reactivation of chronic infection including tuberculosis, herpes simplex, cytomegalovirus, herpes zoster and hepatitis B, d) serious cardiovascular and cardiac events (including stroke and myocardial infarction), e) events related to signs of bone marrow suppression (anemia, lymphocytopenia, thrombocytopenia, pancytopenia), f) malignancy, g) gastrointestinal perforation (incl. gastrointestinal bleeding/diverticulitis), h) liver dysfunction/hepatotoxicity (grade 3 and 4)
# (xi) Adverse events, any grade and serious adverse event, excluding death, within 28 days, grouped by organ classes

# df_ae <- df %>% 
#   select(id_pat, trt, x, ae_28_list, aesi_28)
# # Save
# saveRDS(df_ae, file = "df_ae_recovery.RData")

```
Discussion points

# Define final datasets
```{r echo=TRUE}
# keep the overall set
df_all <- df
# reduce the df set to our standardized set across all trials
df <- df %>% 
  select(id_pat, trt, sex, age, trial, JAKi, 
         ethn, 
         country, icu, sympdur, 
         vacc, 
         clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney,
         any_comorb, comorb_cat, comorb_count,
         crp, 
         # sero, variant,
         vl_baseline, 
         mort_28, mort_28_dimp,
         mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev,
         vir_clear_5, vir_clear_10, vir_clear_15
         )

# export for one-stage model, i.e., add missing variables 
df_os <- df
df_os$sero <- NA
df_os$variant <- NA
# Save
saveRDS(df_os, file = "df_os_recovery.RData")
```

# Missing data plot: One-stage dataset
```{r echo=TRUE}
# Bar plot, missing data, each data point, standardized one-stage dataset
original_order <- colnames(df_os)
missing_plot <- df_os %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - standardized one-stage dataset") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
```
Discussion points
1. Missing variables:
* Baseline:
  - sero
  - variant
* Outcomes:
  - qol_28
2. Missing data in:

# Missing data: Explore for MI
```{r message=FALSE, warning=FALSE}
# keep the core df
# names(df_all)
df_core <- df_all %>%
    select(id_pat, trt, sex, age, trial, JAKi, ethn, 
           # vacc, 
         country, icu, sympdur, clinstatus_baseline, vbaseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney,
         any_comorb, comorb_cat, comorb_count,
         crp, vl_baseline, 
         # sero, variant,
         clinstatus_1, clinstatus_2, clinstatus_3, clinstatus_4, clinstatus_5, clinstatus_6, clinstatus_7, clinstatus_8, clinstatus_9, clinstatus_10, clinstatus_11, clinstatus_12, clinstatus_13, clinstatus_14, clinstatus_15, clinstatus_16, clinstatus_17, clinstatus_18, clinstatus_19, clinstatus_20, clinstatus_21, clinstatus_22, clinstatus_23, clinstatus_24, clinstatus_25, clinstatus_26, clinstatus_27, clinstatus_28, clinstatus_60, clinstatus_29, clinstatus_30, 
         clinstatus_28_imp,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         vir_clear_5, vir_clear_10, vir_clear_15,
         ae_28, ae_28_sev
         )

# Convert character variables to factors
char_vars <- c("id_pat", "sex", "trial", "JAKi", "country", "icu", "ethn", "clinstatus_baseline", "vbaseline", 
               "comed_dexa", "comed_rdv", "comed_toci", "comed_ab", "comed_acoa", "comed_interferon", "comed_other", "comed_cat",
               "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese", "comorb_smoker", "immunosupp", "any_comorb", "comorb_cat", "comorb_autoimm","comorb_cancer", "comorb_kidney", "vl_baseline", "clinstatus_28_imp", "mort_28", "mort_28_dimp", "mort_60", "death_reached", "new_mv_28", "new_mvd_28","discharge_reached", "discharge_reached_sus", "ae_28", "vir_clear_5", "vir_clear_10", "vir_clear_15")
df_core <- df_core %>%
  mutate(across(all_of(char_vars), factor))

# Bar plot, missing data, each data point, core dataset
original_order <- colnames(df_core)
missing_plot <- df_core %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
# Bar plot, missing data, each data point, core dataset, by arm
df_core_int <- df_core %>% 
  filter(trt == 1)
original_order <- colnames(df_core_int)
missing_plot <- df_core_int %>% # Intervention arm
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset, intervention") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
df_core_cont <- df_core %>% 
  filter(trt == 0)
original_order <- colnames(df_core_cont)
missing_plot <- df_core_cont %>% # Control arm
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset, control") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)

### Baseline table, by individuals with no missing data vs any missing data (or only in mort_28)
# df_core <- df_core %>% mutate(complete = ifelse(rowSums(is.na(.)) > 0, 0, 1));table(df_core$complete) # ANY missing 
df_core$resp<-ifelse(is.na(df_core$mort_28), 0, 1);table(df_core$resp) # only mort_28 missing 

# Assign variable list
vars.list <- c("resp", "age", "sympdur"
               ,"trt", "sex", "country", "icu", "ethn", "clinstatus_baseline", "vbaseline", 
               "comed_dexa", "comed_rdv", "comed_toci", "comed_ab", "comed_acoa", "comed_interferon", "comed_other", "comed_cat",
               "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese", "comorb_smoker", "immunosupp", "any_comorb", "comorb_cat", "comorb_count","comorb_autoimm","comorb_cancer", "comorb_kidney", "crp", "vl_baseline"
               , "mort_28", "mort_28_dimp", "mort_60", "death_reached","death_time", "new_mv_28", "new_mvd_28","discharge_reached", "discharge_time", "discharge_reached_sus", "discharge_time_sus", "ae_28", "ae_28_sev", "vir_clear_5", "vir_clear_10", "vir_clear_15")

# By completeness (only mort_28)
table_resp <- CreateTableOne(data = df_core, vars = vars.list[!vars.list %in% c("resp")], strata = "resp", includeNA = T, test = T, addOverall = TRUE)
# Print and display the table
capture.output(
  table_resp <- print(
    table_resp, 
    nonnormal = vars.list, 
    catDigits = 1, 
    SMD = TRUE, 
    showAllLevels = TRUE, 
    test = TRUE, 
    printToggle = FALSE, 
    missing = TRUE))
kable(table_resp, format = "markdown", table.attr = 'class="table"', caption = "By completeness (only mort_28)") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)


### Define variables to be included in imputation set
# table(df_core$vl_baseline)
df_imp <- df_core %>% 
  select("id_pat"
         , "trt", "sex", "age" , "ethn"
         , "country", "sympdur" 
         # ,"vacc" # no info
         # , "trial", "JAKi"  # only 0
         ,"clinstatus_baseline"
         # , "vbaseline" # derived
         # , "comed_rdv" # no info
         # , "comed_toci", "comed_interferon" # no info
         #,  "comed_cat", # derived
         , "comed_dexa", "comed_ab"
         # , "comed_acoa" # no info
         , "comed_other" 
         # , "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese",
         # "comorb_smoker", "immunosupp", "comorb_autoimm", "comorb_cancer", "comorb_kidney", "any_comorb",
         # "comorb_count",  
         ,"comorb_cat" # derived from above, contains most information, and needed as interaction term
         ,"crp"
         ,"vl_baseline"
         # , "sero" , "variant" # very little info
         , clinstatus_1, clinstatus_2, clinstatus_3, clinstatus_4, clinstatus_5, clinstatus_6, clinstatus_7, clinstatus_8, clinstatus_9, clinstatus_10, clinstatus_11, clinstatus_12, clinstatus_13, clinstatus_14, clinstatus_15, clinstatus_16, clinstatus_17, clinstatus_18, clinstatus_19, clinstatus_20, clinstatus_21, clinstatus_22, clinstatus_23, clinstatus_24, clinstatus_25, clinstatus_26, clinstatus_27, clinstatus_28
         # , "clinstatus_28_imp" # imputed via LOVCF above
         , "mort_28"
         # , "mort_28_dimp" # imputed deterministically
         # , "mort_60" # does not contain any additional information compared to death reached
         , "death_reached", "death_time", "new_mv_28", "new_mvd_28", "discharge_reached", "discharge_time"
         # , "discharge_reached_sus", "discharge_time_sus" # same as discharge, does not contain any addition information
         , "ae_28", "ae_28_sev", "vir_clear_5", "vir_clear_10", "vir_clear_15"
         )

# First, table and visualize missing data in various ways
# df_imp %>% 
#   ff_glimpse() # from finalfit package
df_imp %>%
  missing_plot() # from finalfit package
explanatory = c("age", 
  "clinstatus_baseline", "sex",  
  "ethn", "country", "sympdur", "comorb_cat", "comed_dexa", "comed_ab", "comed_other", "crp", "vl_baseline", "ae_28_sev")
dependent = "mort_28"
df_imp %>% # from finalfit package, missing plot
  missing_pairs(dependent, explanatory, position = "fill", )
# Second, let's explore the missingness patterns
md.pattern(df_imp[,c("mort_28", "age", 
  "clinstatus_baseline", "sex",  
  "ethn", "country", "sympdur", "comorb_cat", "comed_dexa", "comed_ab", "comed_other", "crp","vl_baseline", "ae_28_sev")], rotate.names = T)
# Third, let's explore if the variables from my substantive model plus auxiliary variables are associated with mort_28
mort28.aux <- glm(mort_28 ~ trt
            + age 
            + clinstatus_baseline
            + sex
            + ethn
            + country
            + sympdur
            + comorb_cat
            + comed_dexa
            + comed_ab
            + comed_other
            + vl_baseline
            + crp
            # + ae_28_sev
            ,family="binomial"
            ,data=df_imp)
summary(mort28.aux)
# Fourth, let's explore if they are associated with missingness of mort_28:
df_imp %>% 
  missing_compare(dependent, explanatory) %>%
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"))
# Fifth, check age
summary(df_imp$age)
hist(df_imp$age, breaks=50) # looks fine
# Sixth, check sympdur
summary(df_imp$sympdur)
hist(df_imp$sympdur, breaks=50) # skewed -> transform
df_imp$sqsympdur=sqrt(df_imp$sympdur)
hist(df_imp$sqsympdur) # looks fine
# Seventh, check crp
summary(df_imp$crp)
hist(df_imp$crp, breaks=50) # outliers
df_imp <- df_imp %>% # truncate outliers > 500
  mutate(crptrunc = case_when(crp > 500 ~ 500,
                               TRUE ~ crp))
hist(df_imp$crptrunc)
df_imp$sqcrptrunc=sqrt(df_imp$crptrunc)
hist(df_imp$sqcrptrunc) # looks fine


### Reshape to long format
# names(df_imp)
# str(df_imp)
df_imp <- df_imp %>% # rename to differentiate between baseline clinical status and follow-up clinical statuses
  rename(clinicalstatus_baseline = clinstatus_baseline)
# reshape
df_imp_long <- df_imp %>% 
  pivot_longer(cols = starts_with("clinstatus"), names_to = "time", values_to = "clinstatus")
# names(df_imp_long)
# Convert time to numeric
df_imp_long$time <- as.numeric(gsub("clinstatus_", "", df_imp_long$time))
# class(df_imp_long$time)
# summary(df_imp_long$time)
# have clinstatus as numeric
# class(df_imp_long$clinstatus)
# table(df_imp_long$clinstatus_n)
df_imp_long$clinstatus_f <- factor(df_imp_long$clinstatus, levels = 1:6)
df_imp_long$clinstatus_n <- as.numeric(df_imp_long$clinstatus_f)
# table(df_imp_long$clinstatus_n)

### We will impute separately by treatment arm, since we have to expect an effect modification between outcome x trt over time
df_imp_long_int <- df_imp_long %>% 
  filter(trt == 1)
df_imp_long_cont <- df_imp_long %>% 
  filter(trt == 0)

## Explore distribution of clinical status over time, by arm
plot_clinstat_int <- ggplot(df_imp_long_int, aes(x = time, y = clinstatus_n)) +
  stat_summary(fun = "mean", geom = "point", color = "blue", size = 3, na.rm = TRUE) +
  geom_smooth(method = "loess", se = FALSE, color = "red", na.rm = TRUE) +
  labs(title = "Distribution of Clinical Status Over Time / Intervention",
       x = "Time",
       y = "Mean Clinical Status") +
  theme_minimal()
print(plot_clinstat_int)
plot_clinstat_cont <- ggplot(df_imp_long_cont, aes(x = time, y = clinstatus_n)) +
  stat_summary(fun = "mean", geom = "point", color = "blue", size = 3, na.rm = TRUE) +
  geom_smooth(method = "loess", se = FALSE, color = "red", na.rm = TRUE) +
  labs(title = "Distribution of Clinical Status Over Time / Control",
       x = "Time",
       y = "Mean Clinical Status") +
  theme_minimal()
print(plot_clinstat_cont)
```

# Multiple imputation
```{r eval = FALSE}
#### INTERVENTION group
## jomo only accepts numeric or factors, check and adapt
# str(df_imp_long_int)
df_imp_long_int$timesq <- sqrt(df_imp_long_int$time) # see X below
attach(df_imp_long_int)
Y2<-data.frame(mort_28 # level 2 variables (baseline patient characteristics)
               , age
               , sex
               , ethn
               , country
               , comed_dexa
               , comed_ab
               , comed_other
               , comorb_cat
               , sqsympdur
               , sqcrptrunc
               , vl_baseline
               , ae_28_sev
                 )
Y<-data.frame(clinstatus_n) # level 1 variable within clustering variable
X <- cbind(1, data.frame(clinicalstatus_baseline, time, timesq)) # matrix modelling linearity of clinstatus throughout day 28
# X<-data.frame(clinicalstatus_baseline 
#               , time
#               , timesq)
clus<-data.frame(id_pat) # clustering variable (patient)
Z<-data.frame(rep(1,dim(df_imp_long_int)[1]),df_imp_long_int[,c("time")]) # random intercept and random slope
colnames(Z)<-c("const", "time") 

nimp<-50 # set number of iterations

## run jomo
# dry run
imputed_int_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=2) ## clinstatus_baseline has 7 missings
# plot(c(1:2),imputed_int_mcmc$collectbeta[1,1,1:2],type="l")
# plot(c(1:2),imputed_int_mcmc$collectcovu[5,5,1:2],type="l")
set.seed(1569)
imputed_int <- jomo(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=1000, nbetween=1000, nimp=nimp)
# nburn<-1000
# imputed_int_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=nburn)
# plot(c(1:nburn),imputed_int_mcmc$collectbeta[1,1,1:nburn],type="l")
# plot(c(1:nburn),imputed_int_mcmc$collectcovu[5,5,1:nburn],type="l")

# convert to jomo object, split imputations, and exclude original data (imputation "0")
imp.list_int <- imputationList(split(imputed_int, imputed_int$Imputation)[-1])

# checks
round(prop.table(table(imp.list_int[[1]]$`1`$mort_28, useNA = "always"))*100,1) # first imputed dataset
round(prop.table(table(imp.list_int[[1]]$`2`$mort_28, useNA = "always"))*100,1) # second imputed dataset
round(prop.table(table(df_imp_long_int$mort_28, useNA = "always"))*100,1) # original data
summary(imp.list_int[[1]]$`1`$comorb_cat)
summary(imp.list_int[[1]]$`2`$sqsympdur)
summary(imp.list_int[[1]]$`2`$sqcrptrunc)


#### CONTROL group
## jomo only accepts numeric or factors, check and adapt
# str(df_imp_long_cont)
df_imp_long_cont$timesq <- sqrt(df_imp_long_cont$time) # see X below
attach(df_imp_long_cont)
Y2<-data.frame(mort_28 # level 2 variables (baseline patient characteristics)
               , age
               , sex
               , ethn
               , country
               , comed_dexa
               , comed_ab
               , comed_other
               , comorb_cat
               , sqsympdur
               , sqcrptrunc
               , vl_baseline
               , ae_28_sev
                 )
Y<-data.frame(clinstatus_n) # level 1 variable within clustering variable
X<-data.frame(clinicalstatus_baseline # matrix modelling linearity of clinstatus throughout day 28
              , time
              , timesq)
clus<-data.frame(id_pat) # clustering variable (patient)
Z<-data.frame(rep(1,dim(df_imp_long_cont)[1]),df_imp_long_cont[,c("time")]) # random intercept and random slope
colnames(Z)<-c("const", "time") 

nimp<-50 # set number of iterations

# run jomo
set.seed(1569)
imputed_cont <- jomo(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=1000, nbetween=1000, nimp=nimp)
# nburn<-1000
# imputed_cont_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=nburn)
# plot(c(1:nburn),imputed_cont_mcmc$collectbeta[1,1,1:nburn],type="l")
# plot(c(1:nburn),imputed_cont_mcmc$collectcovu[5,5,1:nburn],type="l")

# convert to jomo object, split imputations, and exclude original data (imputation "0")
imp.list_cont <- imputationList(split(imputed_cont, imputed_cont$Imputation)[-1])

# checks
round(prop.table(table(imp.list_cont[[1]]$`1`$mort_28, useNA = "always"))*100,1) # first imputed dataset
round(prop.table(table(imp.list_cont[[1]]$`2`$mort_28, useNA = "always"))*100,1) # second imputed dataset
round(prop.table(table(df_imp_long_cont$mort_28, useNA = "always"))*100,1) # original data
summary(imp.list_cont[[1]]$`1`$comorb_cat)
summary(imp.list_cont[[1]]$`2`$sqsympdur)


#### Add trt back, change from long to wide format, and finally combine the two data frames
imputed_int$trt <- 1
imputed_int_s <- imputed_int %>% # remove imputation variables, not needed anymore
  select(trt, age, sqsympdur, mort_28, sex, comorb_cat, sqcrptrunc, clinicalstatus_baseline, clus, Imputation)
imputed_int_wide <- imputed_int_s %>% # change from long to wide format, i.e. remove duplicates within Imputation sets
  group_by(Imputation) %>%
  distinct(clus, .keep_all = TRUE) 

imputed_cont$trt <- 0 # treatment variable
imputed_cont_s <- imputed_cont %>% # remove imputation variables, not needed anymore
  select(trt, age, sqsympdur, mort_28, sex, comorb_cat, sqcrptrunc, clinicalstatus_baseline, clus, Imputation)
imputed_cont_wide <- imputed_cont_s %>% # change from long to wide format, i.e. remove duplicates within Imputation sets
  group_by(Imputation) %>%
  distinct(clus, .keep_all = TRUE) 

imputed_combined <- rbind(imputed_cont_wide, imputed_int_wide)


#### Convert combined df to jomo object, split imputations, and exclude original data (imputation "0")
imp.list <- imputationList(split(imputed_combined, imputed_combined$Imputation)[-1])


### Checks
round(prop.table(table(imp.list[[1]]$`1`$mort_28, imp.list[[1]]$`1`$trt, useNA = "always"),2)*100,1) # first imputed dataset
round(prop.table(table(imp.list[[1]]$`2`$mort_28, imp.list[[1]]$`2`$trt, useNA = "always"),2)*100,1) # second imputed dataset
round(prop.table(table(df_imp$mort_28, df_imp$trt, useNA = "always"),2)*100,1) # original data
summary(imp.list[[1]]$`1`$comorb_cat)
summary(imp.list[[1]]$`2`$sqsympdur)

```

# (i) Primary endpoint: Mortality at day 28
```{r warning=FALSE}
addmargins(table(df$mort_28, df$trt, useNA = "always"))
addmargins(table(df$mort_28_dimp, df$trt, useNA = "always"))
df$clinstatus_baseline_n <- as.numeric(df$clinstatus_baseline)

# Complete case analysis, substantive model
mort.28 <- df %>% 
  glm(mort_28 ~ trt 
      + age 
      # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# Deterministic imputation
mort.28.dimp <- df %>% 
  glm(mort_28_dimp ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.dimp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# # Multiple imputation analysis under MAR; use mitools package to fit imputed and combined data list and apply Rubin's rules
# mort.28.mi <- imp.list %>%
#   with(glm(mort_28 ~ trt
#            + age
#            + clinicalstatus_baseline
#            , family = binomial)) %>%
#         pool() %>%
#         summary(conf.int = T, exponentiate = T)
# mort.28.mi
```

# (ii) Mortality at day 60
```{r warning=FALSE}
table(df$mort_60, df$trt, useNA = "always")
mort.60 <- df %>% 
  glm(mort_60 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.60, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# (iii) Time to death within max. follow-up time
```{r warning=FALSE}
# table(df$death_reached, df$death_time, useNA = "always")
# table(df$death_reached, df$mort_60, useNA = "always")

# df %>%
#   drop_na(death_time) %>%
#   filter(death_reached == 1) %>%
#   group_by(trt) %>%
#   summarise(median = median(death_time),
#             IQR = IQR(death_time),
#             Q1 = quantile(death_time, probs = 0.25),
#             Q3 = quantile(death_time, probs = 0.75))

# time to death, by group. Kaplan-Meier estimate of conditional survival probability.
km.ttdeath.check <- with(df, Surv(death_time, death_reached))
# head(km.ttdeath.check, 100)

km.ttdeath_trt <- survfit(Surv(death_time, death_reached) ~ trt, data=df)
# summary(km.ttdeath_trt, times = 28)
ttdeath_28d_tbl <- km.ttdeath_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d survival (95% CI)**"
  )
# Nicely formatted table
kable(ttdeath_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# autoplot(km.ttdeath_trt)
survfit2(Surv(death_time, death_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable()

# testing: simple log-rank
# survdiff(Surv(death_time, death_reached) ~ trt, data = df)
# testing: cox ph
ttdeath <- df %>% 
  coxph(Surv(death_time, death_reached) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdeath_reg_tbl <- tbl_regression(ttdeath, exp = TRUE)
# Nicely formatted table
kable(ttdeath_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# (iv) New mechanical ventilation among survivors within 28 days
```{r warning=FALSE}
addmargins(table(df$new_mv_28, df$trt, useNA = "always"))
new.mv.28 <- df %>% 
  glm(new_mv_28 ~ trt 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(new.mv.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
addmargins(table(df$new_mvd_28, df$trt, useNA = "always"))
new.mvd.28 <- df %>% 
  glm(new_mvd_28 ~ trt 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(new.mvd.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# (v) Clinical status at day 28
```{r warning=FALSE}
table(df$clinstatus_28_imp, df$trt, useNA = "always")
clin.28 <- df %>% 
  clm(clinstatus_28_imp ~ trt 
      + age 
      + clinstatus_baseline
      , link= c("logit"), data=.)
# Summary and extract coefficients
coefficients_table <- summary(clin.28)$coefficients
# Calculate Odds Ratios and Confidence Intervals
odds_ratios <- exp(coefficients_table[, "Estimate"])
ci_lower <- exp(coefficients_table[, "Estimate"] - 1.96 * coefficients_table[, "Std. Error"])
ci_upper <- exp(coefficients_table[, "Estimate"] + 1.96 * coefficients_table[, "Std. Error"])
# Create a data frame to store Odds Ratios and CIs
clin.28_tbl <- data.frame(
  "Variable" = rownames(coefficients_table),
  "Odds Ratio" = odds_ratios,
  "CI Lower" = ci_lower,
  "CI Upper" = ci_upper
)
# Nicely formatted table
kable(clin.28_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# (vi) Time to discharge or reaching discharge criteria up to day 28
```{r warning=FALSE}
# Kaplan-Meier estimate of conditional discharge probability
# Censoring the deaths => Cause-specific hazards, i.e., represents the rate per unit of time of the event among those not having failed from other events. Instantaneous rate of occurrence of the given type of event in subjects who are currently event‐free. But by simply censoring the competing event, we bias in favour of comparator (if treatment leads to less deaths)
km.ttdischarge.check <- with(df, Surv(discharge_time, discharge_reached))
# head(km.ttdischarge.check, 100)
km.ttdischarge_trt <- survfit(Surv(discharge_time, discharge_reached) ~ trt, data=df)
# summary(km.ttdischarge_trt, times = 28)
ttdischarge_28d_tbl <- km.ttdischarge_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d hospitalization (95% CI)**"
  )
# Nicely formatted table
kable(ttdischarge_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
survfit2(Surv(discharge_time, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge <- df %>% 
  coxph(Surv(discharge_time, discharge_reached) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_reg_tbl <- tbl_regression(ttdischarge, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Sub-distribution hazards, i.e., represents the rate per unit of time of the event as well as the influence of competing events. Instantaneous rate of occurrence of the given type of event in subjects who have not yet experienced an event of that type.
df <- df %>% # cuminc needs a factor variable with censored patients coded as 0, the event as 1 and the competing event as 2.
  mutate(discharge_reached_comp = case_when (discharge_reached == 0 & (mort_28 == 0 | is.na(mort_28)) ~ 0,
                                             discharge_reached == 1 & (mort_28 == 0 | is.na(mort_28)) ~ 1,
                                             mort_28 == 1 ~ 2))

df$discharge_reached_comp <- as.factor(df$discharge_reached_comp) 
# table(df$discharge_reached_comp)
# table(df$discharge_reached_comp, df$discharge_time)
# table(df$discharge_reached, df$discharge_time)

# Cumulative incidence for the event=discharge (1) and the competing event=death (2)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ 1, data = df)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df) %>% 
  ggcuminc(outcome = c("1", "2")) +
  ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()
# in int only
df_int <- df %>% 
  filter(trt == 1)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_int) %>% 
  ggcuminc(outcome = c("1", "2")) +
  #ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()
# in cont only
df_cont <- df %>% 
  filter(trt == 0)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_cont) %>% 
  ggcuminc(outcome = c("1", "2")) +
  #ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()

# testing: Fine-Gray regression
ttdischarge.comp <- crr(Surv(discharge_time, discharge_reached_comp) ~ trt 
    + age 
    #+ clinstatus_baseline
    ,data = df)
ttdischarge_comp_reg_tbl <- tbl_regression(ttdischarge.comp, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_comp_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Censoring and assigned worst outcome (28d) to competing event (death) // hypothetical estimand where no-one died. (Another option could be, but we don't do it, is to exclude the deaths entirely, i.e. discharge among those that survived (but that might bias in favour of those in control that died more, i.e. healthier comparator))
survfit2(Surv(discharge_time_sens, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sens <- df %>% 
  coxph(Surv(discharge_time_sens, discharge_reached) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_sens_reg_tbl <- tbl_regression(ttdischarge.sens, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sens_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days
# Use cause-specific hazards
survfit2(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall sustained hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sus <- df %>% 
  coxph(Surv(discharge_time_sus, discharge_reached_sus) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_sus_reg_tbl <- tbl_regression(ttdischarge.sus, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sus_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```
Discussion points
1. Use F&G for sens-analysis (sustained discharge)?

# (vii) Viral clearance up to day 5, day 10, and day 15
```{r warning=FALSE}
table(df$vir_clear_5, df$trt, useNA = "always") #  (Viral load value <LOQ and/or undectectable)
# up to 5 days
vir.clear.5 <- df %>% 
  glm(vir_clear_5 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(vir.clear.5, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# up to 10 days
table(df$vir_clear_10, df$trt, useNA = "always") #  (Viral load value <LOQ and/or undectectable)
vir.clear.10 <- df %>% 
  glm(vir_clear_10 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(vir.clear.10, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# up to 15 days
table(df$vir_clear_15, df$trt, useNA = "always") #  (Viral load value <LOQ and/or undectectable)
vir.clear.15 <- df %>% 
  glm(vir_clear_15 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(vir.clear.15, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# (viii) Quality of life at day 28 
```{r warning=FALSE}

```
Discussion points
1. Not available in RECOVERY

# (ix) Adverse event(s) grade 3 or 4, or a serious adverse event(s), excluding death, by day 28
```{r warning=FALSE}
table(df$ae_28, df$trt, useNA = "always")
ae.28 <- df %>% 
  glm(ae_28 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(ae.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
table(df$ae_28_sev, df$trt, useNA = "always")
ae.28.sev <- df %>% 
  glm(ae_28_sev ~ trt 
      + age + clinstatus_baseline
      , family = "poisson", data=.)
summ(ae.28.sev, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Subgroup analysis: Ventilation requirement (proxy for disease severity) on primary endpoint
```{r warning=FALSE}
table(df$clinstatus_baseline, df$mort_28, useNA = "always")
table(df$vbaseline, df$mort_28, useNA = "always")
df$clinstatus_baseline_n <- as.numeric(df$clinstatus_baseline)

mort.28.vent <- df %>% 
  glm(mort_28 ~ trt*clinstatus_baseline_n
      + age 
      #+ clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.vent.vb <- df %>% 
  glm(mort_28 ~ trt*vbaseline
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.vb, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
mort.28.vent.vb.yes <- df %>% 
  filter(vbaseline == 1) %>% # ventilated
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.vb.yes, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.vent.vb.no <- df %>% 
  filter(vbaseline == 0) %>% # not ventilated
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.vb.no, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# class(df$clinstatus_baseline)
mort.28.vent.rs.2 <- df %>% 
  filter(clinstatus_baseline == "2") %>% # no oxygen
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.rs.2, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.vent.rs.3 <- df %>% 
  filter(clinstatus_baseline == "3") %>% # LF oxygen
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.rs.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.vent.rs.4 <- df %>% 
  filter(clinstatus_baseline == "4") %>% # HF oxygen/NIV
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.rs.4, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.vent.rs.5 <- df %>% 
  filter(clinstatus_baseline == "5") %>% # ECMO
  glm(mort_28 ~ trt
      + age 
     # + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent.rs.5, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Subgroup analysis: Age on primary endpoint
```{r warning=FALSE}
# table(df$age, df$mort_28, useNA = "always")
mort.28.age <- df %>% 
  glm(mort_28 ~ trt*age
      #+ age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.age, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
df <- df %>% 
  mutate(age_70 = case_when(age < 70 ~ 0,
                            age > 69 ~ 1))
table(df$age_70, useNA = "always")
mort.28.age.a70 <- df %>% 
  filter(age_70 == 1) %>% # 70 and above
  glm(mort_28 ~ trt
      # + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.age.a70, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.age.b70 <- df %>% 
  filter(age_70 == 0) %>% # below 70
  glm(mort_28 ~ trt
      # + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.age.b70, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Subgroup analysis: Comorbidities on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction
table(df$comorb_cat, df$mort_28, useNA = "always") 
# class(df$comorb_cat)
mort.28.comorb <- df %>%
  glm(mort_28 ~ trt*comorb_cat 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# 4 comorbidity categories as factor
df$comorb_cat_f <- as.factor(df$comorb_cat)
# table(df$comorb_cat_f, df$mort_28, useNA = "always") 
mort.28.comorb.f <- df %>% 
  glm(mort_28 ~ trt*comorb_cat_f 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.f, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# full comorbidity count
# table(df$comorb_count, df$mort_28, useNA = "always") 
mort.28.comorb.count <- df %>%
  glm(mort_28 ~ trt*comorb_count 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.count, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
mort.28.comorb.1 <- df %>% 
  filter(comorb_cat == 1) %>% # no comorbidity
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.1, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.2 <- df %>% 
  filter(comorb_cat == 2) %>% # 1 comorbidity
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.2, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.3 <- df %>% 
  filter(comorb_cat == 3) %>% # multiple comorbidities
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comorb.4 <- df %>% 
  filter(comorb_cat == 4) %>% # immunocompromised
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.4)

```

# Subgroup analysis: Concomitant COVID-19 treatment on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction
table(df$comed_cat, df$trt, useNA = "always")
# 1: patients without Dexamethasone nor Tocilizumab => JAKi effect alone
# 2: patients with Dexamethasone and Tocilizumab => JAKi effect with Dexa + Toci
# 3: patients with Dexamethasone but no Tocilizumab => JAKi effect with Dexa only
# 4: patients with Tocilizumab but no Dexamethasone (if exist) => JAKi effect with Toci only 
mort.28.comed <- df %>%
  glm(mort_28 ~ trt*comed_cat 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# comedication as ordinal factor
df$comed_cat_f <- factor(df$comed_cat, levels = 1:4)
# table(df$comed_cat_f, df$mort_28, useNA = "always") 
mort.28.comed.f <- df %>%
  glm(mort_28 ~ trt*comed_cat_f 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed.f, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
mort.28.comed.1 <- df %>% 
  filter(comed_cat == 1) %>% # without Dexamethasone nor Tocilizumab
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed.1, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.comed.3 <- df %>% 
  filter(comed_cat == 3) %>% # Dexamethasone but no Tocilizumab
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed.3, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Subgroup analysis: Vaccination on adverse events
```{r warning=FALSE}
# table(df$vacc, df$trt, useNA = "always")
# ae.28.vacc <- df %>% 
#   glm(ae_28 ~ trt*vacc
#       + age 
#       #+ clinstatus_baseline 
#       #+ comed_dexa + comed_rdv + comed_toci
#       , family = "binomial", data=.)
# summ(ae.28.vacc, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# SENS Subgroup analysis: Duration since symptom onset on primary endpoint
```{r warning=FALSE}
# table(df$sympdur, df$mort_28, useNA = "always")
mort.28.symp <- df %>% 
  glm(mort_28 ~ trt*sympdur
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.symp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
df <- df %>% 
  mutate(sympdur_cat = case_when(sympdur < 6 ~ 2,
                                 sympdur > 5 & sympdur < 11 ~ 1,
                                 sympdur > 10 ~ 0))
# table(df$sympdur_cat, useNA = "always")
# table(df$sympdur, useNA = "always")
mort.28.sympdur.a10 <- df %>% 
  filter(sympdur_cat == 0) %>% # more than 10 days
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline 
      , family = "binomial", data=.)
summ(mort.28.sympdur.a10, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.sympdur.510 <- df %>% 
  filter(sympdur_cat == 1) %>% # 5-10 days
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.sympdur.510, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.sympdur.b5 <- df %>% 
  filter(sympdur_cat == 2) %>% # 5d or less
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.sympdur.b5, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# SENS Subgroup analysis: CRP on primary endpoint
```{r warning=FALSE}
# table(df$crp, df$mort_28, useNA = "always")
mort.28.crp <- df %>% 
  glm(mort_28 ~ trt*crp
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.crp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# # truncate outliers > 500
# df <- df %>% 
#   mutate(crp_trunc = case_when(crp > 500 ~ 500,
#                                TRUE ~ crp))
# mort.28.crp.trunc <- df %>% 
#   glm(mort_28 ~ trt*crp_trunc
#       + age 
#       + clinstatus_baseline
#       , family = "binomial", data=.)
# summ(mort.28.crp.trunc, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# effect by subgroup
df <- df %>% 
  mutate(crp_75 = case_when(crp < 75 ~ 1,
                            crp > 74 ~ 0))
# table(df$crp_75, useNA = "always")
mort.28.crp.b75 <- df %>% 
  filter(crp_75 == 1) %>% # below 75
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.crp.b75, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

mort.28.crp.a75 <- df %>% 
  filter(crp_75 == 0) %>% # 75 and above
  glm(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.crp.a75, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# SENS Subgroup analysis: Variant on primary endpoint
```{r warning=FALSE}
# # table(df$variant, df$mort_28, useNA = "always")
# mort.28.var <- df %>% 
#   glm(mort_28 ~ trt*variant
#       + age 
#       + clinstatus_baseline 
#       #+ comed_dexa + comed_rdv + comed_toci
#       , family = "binomial", data=.)
# summ(mort.28.var, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```
Discussion points
1. variant not available

# Collect all treatment effect estimates across endpoints (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
result_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric()
)

# Function to extract treatment results from different model types (glm, clm, coxph and crr)
extract_trt_results <- function(model, variable_name, n_int, n_cont) {
  if (inherits(model, "glm") || inherits(model, "clm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "coxph")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "se(coef)"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "tidycrr")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- c(exp(model$tidy$conf.low[1]), exp(model$tidy$conf.high[1]))
    se <- model$tidy$std.error[1]
    p_value <- model$tidy$p.value[1]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_intervention = n_int,
    n_control = n_cont
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_trt_results(mort.28, "death at day 28",
                                        addmargins(table(df$mort_28, df$trt))[3,2], addmargins(table(df$mort_28, df$trt))[3,1]) # adj: age, clinstatus
result_list[[2]] <- extract_trt_results(mort.28.dimp, "death at day 28_dimp",
                                        addmargins(table(df$mort_28_dimp, df$trt))[3,2], addmargins(table(df$mort_28_dimp, df$trt))[3,1]) # adj: age, clinstatus
result_list[[3]] <- extract_trt_results(mort.60, "death at day 60",
                                        addmargins(table(df$mort_60, df$trt))[3,2], addmargins(table(df$mort_60, df$trt))[3,1]) # adj: age, clinstatus
result_list[[4]] <- extract_trt_results(ttdeath, "death within fup",
                                        addmargins(table(df$death_reached, df$trt))[3,2], addmargins(table(df$death_reached, df$trt))[3,1]) # adj: age, clinstatus
result_list[[5]] <- extract_trt_results(new.mv.28, "new MV within 28d",
                                        addmargins(table(df$new_mv_28, df$trt))[3,2], addmargins(table(df$new_mv_28, df$trt))[3,1]) # adj: age, clinstatus
result_list[[6]] <- extract_trt_results(new.mvd.28, "new MV or death within 28d",
                                        addmargins(table(df$new_mvd_28, df$trt))[3,2], addmargins(table(df$new_mvd_28, df$trt))[3,1]) # adj: age, clinstatus
result_list[[7]] <- extract_trt_results(clin.28, "clinical status at day 28",
                                        addmargins(table(df$clinstatus_28_imp, df$trt))[7,2], addmargins(table(df$clinstatus_28_imp, df$trt))[7,1]) # adj: age, clinstatus
result_list[[8]] <- extract_trt_results(ttdischarge, "discharge within 28 days",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], addmargins(table(df$discharge_reached, df$trt))[3,1]) # adj: age, clinstatus
result_list[[9]] <- extract_trt_results(ttdischarge.comp, "discharge within 28 days, death=comp.event",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], addmargins(table(df$discharge_reached, df$trt))[3,1]) # adj: age
result_list[[10]] <- extract_trt_results(ttdischarge.sens, "discharge within 28 days, death=hypo.event",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], addmargins(table(df$discharge_reached, df$trt))[3,1]) # adj: age, clinstatus
result_list[[11]] <- extract_trt_results(ttdischarge.sus, "sustained discharge within 28 days",
                                        addmargins(table(df$discharge_reached_sus, df$trt))[3,2], addmargins(table(df$discharge_reached_sus, df$trt))[3,1]) # adj: age, clinstatus
result_list[[12]] <- extract_trt_results(vir.clear.5, "viral clearance until day 5",
                                        addmargins(table(df$vir_clear_5, df$trt))[3,2], addmargins(table(df$vir_clear_5, df$trt))[3,1]) # adj: age, clinstatus
result_list[[13]] <- extract_trt_results(vir.clear.10, "viral clearance until day 10",
                                        addmargins(table(df$vir_clear_10, df$trt))[3,2], addmargins(table(df$vir_clear_10, df$trt))[3,1]) # adj: age, clinstatus
result_list[[14]] <- extract_trt_results(vir.clear.15, "viral clearance until day 15",
                                        addmargins(table(df$vir_clear_15, df$trt))[3,2], addmargins(table(df$vir_clear_15, df$trt))[3,1]) # adj: age, clinstatus
result_list[[15]] <- extract_trt_results(ae.28, "Any AE grade 3,4 within 28 days",
                                        addmargins(table(df$ae_28, df$trt))[3,2], addmargins(table(df$ae_28, df$trt))[3,1]) # adj: age, clinstatus
result_list[[16]] <- extract_trt_results(ae.28.sev, "AEs grade 3,4 within 28 days",
                                        addmargins(table(df$ae_28_sev, df$trt))[13,2], addmargins(table(df$ae_28_sev, df$trt))[13,1]) # adj: age, clinstatus

# Filter out NULL results and bind the results into a single data frame
result_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
result_df$trial <- "RECOVERY"
result_df$JAKi <- "Baricitinib"

# Nicely formatted table
kable(result_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(result_df, file = "trt_effects_recovery.RData")
```

# Collect all interaction estimates (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
interaction_df <- data.frame(
  variable = character(),
  log_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric()
)
# Extract and format results for the interaction term
extract_interaction <- function(model, variable_name) {
  if (inherits(model, "glm") || inherits(model, "clm")) {
      trt_coef <- coef(model)[grep("^trt:", names(coef(model)))]
      log_odds_ratio <- exp(trt_coef)
      ci <- exp(confint(model)[grep("^trt:", names(coef(model))), ])
      se <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Std. Error"]
      p_value <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Pr(>|z|)"]
  } else if (inherits(model, "logistf")) {
      trt_coef <- coef(model)[grep("^trt:", names(coef(model)))]
      log_odds_ratio <- exp(trt_coef)
      ci <- exp(confint(model)[grep("^trt:", names(coef(model))), ])
      se <- sqrt(diag(vcov(model)))[grep("^trt:", names(coef(model)))]
      p_value <- model$prob[grep("^trt:", names(coef(model)))]
  } else {
    stop("Unsupported model class")
  }
      # capture the results
      result <- data.frame(
        variable = variable_name,
        log_odds_ratio = log_odds_ratio,
        ci_lower = ci[1],
        ci_upper = ci[2],
        standard_error = se,
        p_value = p_value
      )
    return(result)
}
# Loop through
result_list <- list()

result_list[[1]] <- extract_interaction(mort.28.vent, "respiratory support") # adj: age, clinstatus
result_list[[2]] <- extract_interaction(mort.28.vent.vb, "ventilation") # adj: age, clinstatus
result_list[[3]] <- extract_interaction(mort.28.age, "age") # adj: age, clinstatus
result_list[[4]] <- extract_interaction(mort.28.comorb, "comorbidity") # adj: age, clinstatus
result_list[[5]] <- extract_interaction(mort.28.comed, "comedication") # adj: age, clinstatus
result_list[[6]] <- extract_interaction(ae.28.vacc, "vaccination on AEs") # adj: age, clinstatus
result_list[[7]] <- extract_interaction(mort.28.symp, "symptom duration") # adj: age, clinstatus
result_list[[8]] <- extract_interaction(mort.28.crp, "crp") # adj: age, clinstatus
# result_list[[x]] <- extract_interaction(mort.28.var, "variant") # variant not available

# Filter out NULL results and bind the results into a single data frame
interaction_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
interaction_df$trial <- "RECOVERY"
interaction_df$JAKi <- "Baricitinib"

# Nicely formatted table
kable(interaction_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(interaction_df, file = "int_effects_recovery.RData")
```

# Collect all subgroup treatment effect estimates
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
subgroup_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric(),
  n_intervention = numeric(),
  n_intervention_tot = numeric(),
  n_control = numeric(),
  n_control_tot = numeric()
)

# Function to extract subgroup treatment results
extract_subgroup_results <- function(model, variable_name, n_int, n_int_tot, n_cont, n_cont_tot) {
  if (inherits(model, "glm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "logistf")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", names(coef(model))), ])
    se <- sqrt(diag(vcov(model)))[grep("^trt", names(coef(model)))]
    p_value <- model$prob[grep("^trt", names(coef(model)))]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_intervention = n_int,
    n_intervention_tot = n_int_tot,
    n_control = n_cont,
    n_control_tot = n_cont_tot
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_subgroup_results(mort.28.vent.vb.yes, "High-flow or non-invasive, mechanical, or ECMO",
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,3,1]) 
result_list[[2]] <- extract_subgroup_results(mort.28.vent.vb.no, "None or low-flow oxygen",
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,3,1]) 
result_list[[3]] <- extract_subgroup_results(mort.28.vent.rs.2, "No oxygen",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,3,1])
result_list[[4]] <- extract_subgroup_results(mort.28.vent.rs.3, "low-flow oxygen",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,2,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,3,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,2,1], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,3,1])
result_list[[5]] <- extract_subgroup_results(mort.28.vent.rs.4, "high-flow oxygen / NIV",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,2,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,3,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,2,1], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,3,1])
result_list[[6]] <- extract_subgroup_results(mort.28.vent.rs.5, "Mechanical ventilation / ECMO",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,2,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,3,2], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,2,1], 
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,3,1]) 
result_list[[7]] <- extract_subgroup_results(mort.28.age.a70, "70 years and above",
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,3,1]) 
result_list[[8]] <- extract_subgroup_results(mort.28.age.b70, "below 70 years",
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,3,1]) 
result_list[[9]] <- extract_subgroup_results(mort.28.comorb.1, "No comorbidity",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,3,1])
result_list[[10]] <- extract_subgroup_results(mort.28.comorb.2, "One comorbidity",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,3,1])
result_list[[11]] <- extract_subgroup_results(mort.28.comorb.3, "Multiple comorbidities",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,3,1])
result_list[[12]] <- extract_subgroup_results(mort.28.comorb.4, "Immunocompromised_firth",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,3,1]) 
result_list[[13]] <- extract_subgroup_results(mort.28.comed.1, "No Dexa, no Tocilizumab",
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,3,1])

others?


result_list[[14]] <- extract_subgroup_results(mort.28.comed.3, "Dexa, but no Tocilizumab",
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,3,1])
result_list[[15]] <- extract_subgroup_results(ae.28.vacc.1, "vaccinated",
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,2,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,3,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,2,1],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,3,1])
result_list[[16]] <- extract_subgroup_results(ae.28.vacc.0, "not vaccinated",
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,3,1])
result_list[[17]] <- extract_subgroup_results(mort.28.sympdur.a10, "More than 10 days",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,3,1])
result_list[[18]] <- extract_subgroup_results(mort.28.sympdur.510, "Between 5-10 days",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,3,1])
result_list[[19]] <- extract_subgroup_results(mort.28.sympdur.b5, "5 days and less",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,2,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,3,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,2,1], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,3,1])
result_list[[20]] <- extract_subgroup_results(mort.28.crp.a75, "CRP 75 and higher",
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,3,1])
result_list[[21]] <- extract_subgroup_results(mort.28.crp.b75, "CRP below 75",
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,3,1])

# Filter out NULL results and bind the results into a single data frame
subgroup_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
subgroup_df$trial <- "RECOVERY"
subgroup_df$JAKi <- "Baricitinib"

# Nicely formatted table
kable(subgroup_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(subgroup_df, file = "subgroup_effects_recovery.RData")
```
