---
title: "TOFACOV"
author: "A.Amstutz"
date: "2023-08-30"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)
library(haven) # Read sas files
library(here)
library(kableExtra)

library(jtools) # for summ() and plot_summs
library(sjPlot) # for tab_model
library(ggplot2) # survival/TTE analyses and other graphs
library(ggsurvfit) # survival/TTE analyses
library(survival) # survival/TTE analyses
library(gtsummary) # survival/TTE analyses
library(ggfortify) # autoplot
library(tidycmprsk) # competing risk analysis
library(ordinal) # clinstatus ordinal regression
library(mosaic) # OR for 0.5-corrected 2x2 table in case of rare events
library(logistf) # Firth regression in case of rare events

library(finalfit) # missing data exploration
library(mice) # multiple imputation
library(jomo) # multiple imputation
library(mitools) # multiple imputation
```

# Load Data
```{r, include=FALSE}
df <- read_xlsx("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/TOFACOV/JAK inhibitor metaanalysis TOFACOV2 NEW.xlsx")
```

# Baseline Characteristics
```{r echo=TRUE}
df$trial <- c("TOFACOV")
df$JAKi <- c("Tofacitinib")
df <- df %>%
  rename(id_pat = ID)
df <- df %>% # no missing in sex
  mutate(sex = case_when(sex == 1 ~ "female",
                         sex == 2 ~ "male"))

df %>% 
  drop_na(age) %>% 
  ggplot(aes(x = age)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Age",
       x = "Age",
       y = "Density")

# no missing in sex and age and ethnicity, no duplicates in ID, trt according to publication, add country
df$country <- c("Italy")
df$icu <- 0 # no icu patients at enrolment

# Days with symptoms prior to randomization
df %>% # no missings
  drop_na(sympdur) %>% 
  ggplot(aes(x = sympdur)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "Symptom Duration",
       y = "Density")

# Severity of COVID-19 with respect to respiratory support at randomisation // tofacov only recruited mild (2) and moderate (3) covid-19, noone with any kind of ventilation (4 above)
df$clinstatus_baseline <- factor(df$clinstatus_baseline, levels = 1:6) ## no missing data
addmargins(table(df$clinstatus_baseline, df$trt, useNA = "always"))

df <- df %>% 
  mutate(vbaseline = case_when(clinstatus_baseline == "2" | clinstatus_baseline == "3" ~ 0,
                                clinstatus_baseline == "4" | clinstatus_baseline == "5" ~ 1))

# Co-medication at baseline
df$comed_toci <- 0 # see publication: "It is noteworthy that no patients have received tocilizumab. Concomitant use of tofacitinib and tocilizumab was prohibited by protocol"
df$comed_interferon <- 0 # no interferon used

## group them for the subgroup analysis, according to protocol
df <- df %>% 
  mutate(comed_cat = case_when(comed_dexa == 0 & comed_toci == 0 ~ 1, # patients without Dexamethasone nor Tocilizumab
                               comed_dexa == 1 & comed_toci == 0 ~ 2, # patients with Dexamethasone but no Tocilizumab
                               comed_dexa == 1 & comed_toci == 1 ~ 3, # patients with Dexamethasone and Tocilizumab
                               comed_dexa == 0 & comed_toci == 1 ~ 4)) # patients with Tocilizumab but no Dexamethasone (if exist)

# Comorbidity at baseline, including immunocompromised // no immunosupp in tofacov and few comorbidities
df$comorb_cancer <- NA
df$comorb_autoimm <- NA
df$comorb_kidney <- NA
df <- df %>% # 18 missing - as in publication
  mutate(any_comorb = case_when(comorb_lung == 1 | comorb_liver == 1 | comorb_cvd == 1 |
                                  comorb_aht == 1 | comorb_dm == 1 | comorb_obese == 1 | comorb_smoker == 1
                                | immunosupp == 1 | comorb_cancer == 1 | comorb_autoimm == 1 | comorb_kidney == 1 
                                  ~ 1,
                                comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 & comorb_cancer == 0 & comorb_autoimm == 0 & comorb_kidney == 0
                                ~ 0))
# the remaining missing have no evidence for comorbidity -> recode as 0
df <- df %>% 
  mutate(any_comorb = case_when(is.na(any_comorb) ~ 0,
                                TRUE ~ any_comorb))
## group them for the subgroup analysis, according to protocol // count all pre-defined comorbidities per patient first
comorb <- df %>% 
  select(id_pat, comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp, comorb_kidney, comorb_autoimm, comorb_cancer)
comorb$comorb_count <- NA
for (i in 1:dim(comorb)[[1]]) {
  comorb$comorb_count[i] <- ifelse(
    sum(comorb[i, ] %in% c(1)) > 0,
    sum(comorb[i, ] %in% c(1)),
    NA
  )
}
comorb <- comorb %>% 
  mutate(comorb_count = case_when(comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 & comorb_cancer == 0 & comorb_autoimm == 0 & comorb_kidney == 0 ~ 0,
                                TRUE ~ comorb_count))
# the remaining missing have no evidence for comorbidity -> recode as 0
comorb <- comorb %>% 
  mutate(comorb_count = case_when(is.na(comorb_count) ~ 0,
                                TRUE ~ comorb_count))
df <- left_join(df, comorb[, c("comorb_count", "id_pat")], by = join_by(id_pat == id_pat)) ## merge imputed variable back
df <- df %>% # no missing
  mutate(comorb_cat = case_when(immunosupp == 1 ~ 4, # immunocompromised
                                comorb_count == 0 ~ 1, # no comorbidity
                                comorb_count == 1 & (immunosupp == 0 | is.na(immunosupp)) ~ 2, # one comorbidity
                                comorb_count >1 & (immunosupp == 0 | is.na(immunosupp)) ~ 3)) # multiple comorbidities
df <- df %>%
  mutate(comorb_any = case_when(comorb_count == 0 ~ 0, # no comorbidity
                                comorb_count >0 ~ 1)) # any comorbidity

# CRP
df$crp <- as.numeric(df$`crp mg/L`) ## 1 missing
df %>% 
  drop_na(crp) %>% 
  ggplot(aes(x = crp)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of CRP",
       x = "CRP",
       y = "Density")

# Vaccination
# table(df$vacc, useNA = "always") # no missing

# Viremia
# Variant
# Serology

```

# Endpoints
```{r echo=TRUE}
# time to event data
df$death_d <- as.numeric(df$death_date - df$randdate) 
df$discharge_d <- as.numeric(df$discharge_date - df$randdate)

# (i) Primary outcome: Mortality at day 28
df <- df %>% # corresponds to publication: only 1 death (in tofa arm)
  mutate(mort_28 = case_when(death_d <29 ~ 1, 
                             TRUE ~ 0 # Tofacov hat complete follow-up, only 1 died, all others were alive at day 28
                             )) 
df$mort_28_dimp <- df$mort_28 # to be consistent with other trials


# (ii) Mortality at day 60
df$mort_60 <- df$mort_28 # max fup time in TOFACOV was 28 days; thus mort_60 imputed from mort_28


# (iii) Time to death within max. follow-up time
df$death_reached <- df$mort_28
df <- df %>% # no missing and those that were discharged and afterwards have correct time to event data
  mutate(death_time = case_when(death_d >=0 ~ c(death_d), # time to death, if no time to death, then...
                                TRUE ~ 28)) # time to death censoring data (there were no withdrawals/LTFU)
# max follow-up time in ACTT2 was 28 days => no restriction of time window to 60 days.


# (iv) New mechanical ventilation among survivors within 28 days. TOFACOV only included clinstatus 2 and 3.
df <- df %>% # the 1 that died is NA => correct
  mutate(new_mv_28 = case_when(mort_28 == 0 & (clinstatus_2 == 5 | clinstatus_7 == 5 | clinstatus_14 == 5) ~ 1,
                               mort_28 == 0 ~ 0))

# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
df <- df %>% # no missing anymore => correct
  mutate(new_mvd_28 = case_when(new_mv_28 == 1 | mort_28 == 1 ~ 1,
                                new_mv_28 == 0 | mort_28 == 0 ~ 0))


# (v) Clinical status at day 28
df <- df %>% # They called and followed all patients at day 28 and recorded no new serious adverse events, re-admissions or deaths besides the one that died on day 26
  mutate(clinstatus_28 = case_when(mort_28 == 1 ~ 6,
                                   discharge_d <29 ~ 1, # discharged alive / reached discharge criteria within 28d
                                   clinstatus_14 == 3 ~ 3, # these ones are still hosp at day 28
                                   clinstatus_14 == 4 ~ 4,
                                   clinstatus_14 == 5 ~ 5)) 
df$clinstatus_28 <- factor(df$clinstatus_28, levels = 1:6) # no missing
df$clinstatus_28_imp <- df$clinstatus_28 # equal to clinstatus_28_imp

# (vi) Time to discharge or reaching discharge criteria up to day 28
df <- df %>%
  mutate(discharge_reached = case_when(discharge_d <29 ~ 1,
                                       TRUE ~ 0))
df <- df %>% 
  mutate(discharge_time = case_when(discharge_d >=0 ~ c(discharge_d), # time to discharge, if no time to discharge, then...
                                    death_d >=0 ~ death_d)) # time to discharge censoring time
df <- df %>% # restrict to max fup time 28d
  mutate(discharge_time = case_when(discharge_time >28 ~ 28,
                                    TRUE ~ discharge_time))
df <- df %>% # add 28d for those that died
  mutate(discharge_time_sens = case_when(mort_28 == 1 ~ 28,
                                    TRUE ~ discharge_time))

# (vi) Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days. There were no re-admissions within 28d
df$discharge_reached_sus <- df$discharge_reached
df$discharge_time_sus <- df$discharge_time


# (vii) Viral clearance up to day 5, day 10, and day 15 (Viral load value <LOQ and/or undectectable): Not available in TOFACOV
# (viii) Quality of life at day 28: Not available in TOFAVOC


# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28
table(df$ae_28, df$trt, useNA = "always")
# table(df$ae_28, df$mort_28, useNA = "always") # corresponds to publication and excludes the death

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
df$ae_28_sev <- df$ae_28 # there were only 1 AE grade 3/4 per person 

# (ix) Sens-analysis: Alternative definition/analysis of outcome: time to first (of these) adverse event, within 28 days, considering death as a competing risk (=> censor and set to 28 days)
# time to first ae not available

# (x) Adverse events of special interest within 28 days: a) thromboembolic events (venous thromboembolism, pulmonary embolism, arterial thrombosis), b) secondary infections (bacterial pneumonia including ventilator-associated pneumonia, meningitis and encephalitis, endocarditis and bacteremia, invasive fungal infection including pulmonary aspergillosis), c) Reactivation of chronic infection including tuberculosis, herpes simplex, cytomegalovirus, herpes zoster and hepatitis B, d) serious cardiovascular and cardiac events (including stroke and myocardial infarction), e) events related to signs of bone marrow suppression (anemia, lymphocytopenia, thrombocytopenia, pancytopenia), f) malignancy, g) gastrointestinal perforation (incl. gastrointestinal bleeding/diverticulitis), h) liver dysfunction/hepatotoxicity (grade 3 and 4)
df_aesi <- df %>% # these were >5 ULN 
  mutate(aesi = case_when(ae_28_spec == "Pulmonary embolism" ~ "thrombo",
                             ae_28_spec == "ALT increase" ~ "hepatox")) %>% 
  filter(!is.na(aesi)) %>% 
  filter(ae_28 == 1) %>% 
  select(id_pat, trt, ae_28_spec, aesi) %>% 
  rename(ae = ae_28_spec)
# Save
saveRDS(df_aesi, file = "df_aesi_tofacov.RData")


# (xi) Adverse events, any grade and serious adverse event, excluding death, within 28 days, grouped by organ classes
df_ae <- df %>% # these were >5 ULN 
  mutate(aesi = case_when(ae_28_spec == "Pulmonary embolism" ~ "thrombo",
                             ae_28_spec == "ALT increase" ~ "hepatox")) %>%
  filter(ae_28 == 1) %>% 
  select(id_pat, trt, ae_28_spec, aesi) %>% 
  rename(ae = ae_28_spec)
# Save
saveRDS(df_ae, file = "df_ae_tofacov.RData")
```
Discussion points
1. Outcome definition "new_mv_28" differs to original trial publication (included new non-invasive ventilations, i.e. clinstatus == 4)

# Define final datasets
```{r echo=TRUE}
# keep the overall set
df_all <- df
# reduce the df set to our standardized set across all trials
df <- df %>% 
  select(id_pat, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline, trial,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney,
         any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, 
         # sero, vl_baseline, variant,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev
         # vir_clear_5, vir_clear_10, vir_clear_15
         )
# export for one-stage model, i.e., add missing variables 
df_os <- df
df_os$sero <- NA
df_os$vl_baseline <- NA
df_os$variant <- NA
df_os$vir_clear_5 <- NA
df_os$vir_clear_10 <- NA
df_os$vir_clear_15 <- NA
# Save
saveRDS(df_os, file = "df_os_tofacov.RData")
```

# Missing data plot: One-stage dataset
```{r echo=TRUE}
# Bar plot, missing data, each data point, standardized one-stage dataset
original_order <- colnames(df_os)
missing_plot <- df_os %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - standardized one-stage dataset") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
```
Discussion points
1. Missing variables:
* Baseline:
  - variant
  - sero
  - vl_baseline
  - clarify comorbidity categories (autoimmun, CKD, cancer)
* Outcomes:
  - vir_clear_5, vir_clear_10, vir_clear_15
2. Missing data:
- 1 NA in crp 
- NAs in new_mv_28: Not part of denominator

# Missing data: Explore for MI
```{r message=FALSE, warning=FALSE}
# keep the core df
# names(df_all)
df_core <- df_all %>% 
  select(id_pat, JAKi, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline, vbaseline, trial,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         comorb_autoimm, comorb_cancer, comorb_kidney,
         any_comorb, comorb_cat, comorb_any, comorb_count,
         crp, 
         # sero, vl_baseline, variant,
         clinstatus_2, clinstatus_7, clinstatus_14,
         mort_28, mort_28_dimp, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_time_sens, discharge_reached_sus, discharge_time_sus,
         ae_28, ae_28_sev
         # vir_clear_5, vir_clear_10, vir_clear_15
         )

# Convert character variables to factors
char_vars <- c("id_pat", "sex", "trial", "JAKi", "ethn", "vacc", "country", "icu", "clinstatus_baseline", "vbaseline", 
               "comed_dexa", "comed_rdv", "comed_toci", "comed_ab", "comed_acoa", "comed_interferon", "comed_other", "comed_cat",
               "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese", "comorb_smoker", "immunosupp", "comorb_autoimm", "comorb_cancer", "comorb_kidney", "any_comorb", "comorb_cat", "comorb_any", "clinstatus_28_imp", "mort_28", "mort_28_dimp", "mort_60", "death_reached", "new_mv_28", "new_mvd_28","discharge_reached", "discharge_reached_sus", "ae_28")
df_core <- df_core %>%
  mutate(across(all_of(char_vars), factor))

# Bar plot, missing data, each data point, core dataset
original_order <- colnames(df_core)
missing_plot <- df_core %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
# Bar plot, missing data, each data point, core dataset, by arm
df_core_int <- df_core %>% 
  filter(trt == 1)
original_order <- colnames(df_core_int)
missing_plot <- df_core_int %>% # Intervention arm
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset, intervention") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
df_core_cont <- df_core %>% 
  filter(trt == 0)
original_order <- colnames(df_core_cont)
missing_plot <- df_core_cont %>% # Control arm
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data - core dataset, control") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)

#### only CRP missing -> MI
### Baseline table, by individuals with no missing data vs any missing data (or only in crp)
df_core$resp<-ifelse(is.na(df_core$crp), 0, 1);table(df_core$resp) # only mort_28 missing 

# Assign variable list
vars.list <- c("resp", "age", "sympdur", "trt", "sex", "trial", "JAKi", "ethn", "vacc", "country", "icu", "sympdur", "clinstatus_baseline", "vbaseline", "comed_dexa", "comed_rdv", "comed_toci", "comed_ab", "comed_acoa", "comed_interferon", "comed_other", "comed_cat",
               "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese", "comorb_smoker", "immunosupp", "any_comorb", "comorb_cat", "comorb_any", "comorb_count","comorb_autoimm","comorb_cancer", "comorb_kidney", "crp"
               , "mort_28", "mort_28_dimp", "mort_60", "death_reached","death_time", "new_mv_28", "new_mvd_28","discharge_reached", "discharge_time", "discharge_reached_sus", "discharge_time_sus", "ae_28", "ae_28_sev")

# By completeness (only mort_28)
table_resp <- CreateTableOne(data = df_core, vars = vars.list[!vars.list %in% c("resp")], strata = "resp", includeNA = T, test = T, addOverall = TRUE)
# Print and display the table
capture.output(
  table_resp <- print(
    table_resp, 
    nonnormal = vars.list, 
    catDigits = 1, 
    SMD = TRUE, 
    showAllLevels = TRUE, 
    test = TRUE, 
    printToggle = FALSE, 
    missing = TRUE))
kable(table_resp, format = "markdown", table.attr = 'class="table"', caption = "By completeness (only crp)") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)


### Define variables to be included in imputation set
# table(df_core$comorb_cat, useNA = "always")
df_imp <- df_core %>% 
  select("id_pat"
         , "trt", "sex", "age" 
         # , "ethn" # no info
         # , "country" # no info
         , "sympdur" 
         , "vacc"
         # , "trial", "JAKi"  # only 0
         , "clinstatus_baseline"
         # , "vbaseline" # derived
         , "comed_rdv"
         # , "comed_toci", "comed_interferon" # no info
         #,  "comed_cat", # derived
         , "comed_dexa", "comed_ab"
         , "comed_acoa" # no info
         # , "comed_other" # no info
         # , "comorb_lung", "comorb_liver", "comorb_cvd", "comorb_aht", "comorb_dm", "comorb_obese",
         # "comorb_smoker", "immunosupp", "comorb_autoimm", "comorb_cancer", "comorb_kidney", "any_comorb",
         # "comorb_count", 
         # "comorb_any", 
         ,"comorb_cat" # derived from above, contains most information, and needed as interaction term
         ,"crp"
         # ,"vl_baseline"
         # , "sero" , "variant" # no info
         , clinstatus_2, clinstatus_7, clinstatus_14, clinstatus_28_imp
         # , "clinstatus_28_imp" # imputed via LOVCF above
         , "mort_28"
         # , "mort_28_dimp" # imputed deterministically
         # , "mort_60" # does not contain any additional information compared to death reached
         , "death_reached", "death_time", "new_mv_28", "new_mvd_28", "discharge_reached", "discharge_time"
         # , "discharge_reached_sus", "discharge_time_sus" # same as discharge, does not contain any addition information
         , "ae_28", "ae_28_sev"
         # , "vir_clear_5", "vir_clear_10", "vir_clear_15"
         )

# First, table and visualize missing data in various ways
# df_imp %>% 
#   ff_glimpse() # from finalfit package
df_imp %>%
  missing_plot() # from finalfit package
explanatory = c("age", 
  "clinstatus_baseline", "sex", "vacc",  
  "sympdur", "comorb_cat", "comed_rdv", "comed_dexa", "comed_ab", "comed_acoa", "mort_28", "ae_28_sev")
dependent = "crp"
df_imp %>% # from finalfit package, missing plot
  missing_pairs(dependent, explanatory, position = "fill", )
# Second, let's explore the missingness patterns
md.pattern(df_imp[,c("age", "crp",
  "clinstatus_baseline", "sex", "vacc",  
  "sympdur", "comorb_cat", "comed_rdv", "comed_dexa", "comed_ab", "comed_acoa", "mort_28", "ae_28_sev")], rotate.names = T)
# Third, let's explore if the variables from my substantive model plus auxiliary variables are associated with crp
mort28.aux <- lm(crp ~
            + age 
            + clinstatus_baseline
            + sex
            + vacc
            + sympdur
            + comorb_cat
            + comed_dexa
            + comed_ab
            + comed_acoa
            + comed_rdv
            + mort_28
            + clinstatus_28_imp
            + ae_28_sev
            ,family="binomial"
            ,data=df_imp)
summary(mort28.aux)
# Fourth, let's explore if they are associated with missingness of crp:
df_imp %>% 
  missing_compare(dependent, explanatory) %>%
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"))
# Fifth, check age
summary(df_imp$age)
hist(df_imp$age, breaks=50) # looks fine
# Sixth, check sympdur
summary(df_imp$sympdur)
hist(df_imp$sympdur, breaks=50)
# Seventh, check crp
summary(df_imp$crp)
hist(df_imp$crp, breaks=50) # outliers
# df_imp <- df_imp %>% # truncate outliers > 500
#   mutate(crptrunc = case_when(crp > 500 ~ 500,
#                                TRUE ~ crp))
# hist(df_imp$crptrunc)
# df_imp$sqcrptrunc=sqrt(df_imp$crptrunc)
# hist(df_imp$sqcrptrunc) # looks fine


### Reshape to long format
# names(df_imp)
# str(df_imp)
df_imp <- df_imp %>% # rename to differentiate between baseline clinical status and follow-up clinical statuses
  rename(clinicalstatus_baseline = clinstatus_baseline)

# reshape
df_imp$clinstatus_28_imp <- as.numeric(df_imp$clinstatus_28_imp)
df_imp_long <- df_imp %>% 
  pivot_longer(cols = starts_with("clinstatus"), names_to = "time", values_to = "clinstatus")
names(df_imp_long)

# Convert time to numeric
df_imp_long$time <- as.numeric(gsub("clinstatus_", "", df_imp_long$time))
# class(df_imp_long$time)
# summary(df_imp_long$time)
# have clinstatus as numeric
df_imp_long$clinstatus_f <- factor(df_imp_long$clinstatus, levels = 1:6)

### We will impute separately by treatment arm, since we have to expect an effect modification between outcome x trt over time
df_imp_long_int <- df_imp_long %>% 
  filter(trt == 1)
df_imp_long_cont <- df_imp_long %>% 
  filter(trt == 0)

## Explore distribution of clinical status over time, by arm
plot_clinstat_int <- ggplot(df_imp_long_int, aes(x = time, y = clinstatus)) +
  stat_summary(fun = "mean", geom = "point", color = "blue", size = 3, na.rm = TRUE) +
  geom_smooth(method = "loess", se = FALSE, color = "red", na.rm = TRUE) +
  labs(title = "Distribution of Clinical Status Over Time / Intervention",
       x = "Time",
       y = "Mean Clinical Status") +
  theme_minimal()
print(plot_clinstat_int)
plot_clinstat_cont <- ggplot(df_imp_long_cont, aes(x = time, y = clinstatus)) +
  stat_summary(fun = "mean", geom = "point", color = "blue", size = 3, na.rm = TRUE) +
  geom_smooth(method = "loess", se = FALSE, color = "red", na.rm = TRUE) +
  labs(title = "Distribution of Clinical Status Over Time / Control",
       x = "Time",
       y = "Mean Clinical Status") +
  theme_minimal()
print(plot_clinstat_cont)
```

# Multiple imputation // only CRP is missing: No MI for CRP, as per protocol
```{r eval = FALSE}
# #### INTERVENTION group
# ## jomo only accepts numeric or factors, check and adapt
# # str(df_imp_long_int)
# df_imp_long_int$timesq <- sqrt(df_imp_long_int$time) # see X below
# attach(df_imp_long_int)
# Y2<-data.frame(mort_28 # level 2 variables (baseline patient characteristics)
#                , age
#                , sex
#                , ethn
#                , vacc
#                , comed_dexa
#                , comed_ab
#                , comed_other
#                , comorb_cat
#                , sympdur
#                , sqcrptrunc
#                  )
# Y<-data.frame(clinstatus) # level 1 variable within clustering variable
# X<-data.frame(clinicalstatus_baseline # matrix modelling linearity of clinstatus throughout day 28
#               , time
#               , timesq)
# clus<-data.frame(id_pat) # clustering variable (patient)
# Z<-data.frame(rep(1,dim(df_imp_long_int)[1]),df_imp_long_int[,c("time")]) # random intercept and random slope
# colnames(Z)<-c("const", "time") 
# 
# nimp<-50 # set number of iterations
# 
# ## run jomo
# # dry run
# imputed_int_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=2) ## clinstatus_baseline has 7 missings
# # plot(c(1:2),imputed_int_mcmc$collectbeta[1,1,1:2],type="l")
# # plot(c(1:2),imputed_int_mcmc$collectcovu[5,5,1:2],type="l")
# set.seed(1569)
# imputed_int <- jomo(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=1000, nbetween=1000, nimp=nimp)
# # nburn<-1000
# # imputed_int_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=nburn)
# # plot(c(1:nburn),imputed_int_mcmc$collectbeta[1,1,1:nburn],type="l")
# # plot(c(1:nburn),imputed_int_mcmc$collectcovu[5,5,1:nburn],type="l")
# 
# # convert to jomo object, split imputations, and exclude original data (imputation "0")
# imp.list_int <- imputationList(split(imputed_int, imputed_int$Imputation)[-1])
# 
# # checks
# round(prop.table(table(imp.list_int[[1]]$`1`$mort_28, useNA = "always"))*100,1) # first imputed dataset
# round(prop.table(table(imp.list_int[[1]]$`2`$mort_28, useNA = "always"))*100,1) # second imputed dataset
# round(prop.table(table(df_imp_long_int$mort_28, useNA = "always"))*100,1) # original data
# summary(imp.list_int[[1]]$`1`$comorb_cat)
# summary(imp.list_int[[1]]$`2`$sqsympdur)
# summary(imp.list_int[[1]]$`2`$sqcrptrunc)
# 
# 
# #### CONTROL group
# ## jomo only accepts numeric or factors, check and adapt
# # str(df_imp_long_cont)
# df_imp_long_cont$timesq <- sqrt(df_imp_long_cont$time) # see X below
# attach(df_imp_long_cont)
# Y2<-data.frame(mort_28 # level 2 variables (baseline patient characteristics)
#                , age
#                , sex
#                , ethn
#                , country
#                , comed_dexa
#                , comed_ab
#                , comed_other
#                , comorb_cat
#                , sqsympdur
#                , sqcrptrunc
#                , vl_baseline
#                , ae_28_sev
#                  )
# Y<-data.frame(clinstatus_n) # level 1 variable within clustering variable
# X<-data.frame(clinicalstatus_baseline # matrix modelling linearity of clinstatus throughout day 28
#               , time
#               , timesq)
# clus<-data.frame(id_pat) # clustering variable (patient)
# Z<-data.frame(rep(1,dim(df_imp_long_cont)[1]),df_imp_long_cont[,c("time")]) # random intercept and random slope
# colnames(Z)<-c("const", "time") 
# 
# nimp<-50 # set number of iterations
# 
# # run jomo
# set.seed(1569)
# imputed_cont <- jomo(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=1000, nbetween=1000, nimp=nimp)
# # nburn<-1000
# # imputed_cont_mcmc<-jomo.MCMCchain(Y=Y, Y2=Y2, X=X, Z=Z, clus=clus, nburn=nburn)
# # plot(c(1:nburn),imputed_cont_mcmc$collectbeta[1,1,1:nburn],type="l")
# # plot(c(1:nburn),imputed_cont_mcmc$collectcovu[5,5,1:nburn],type="l")
# 
# # convert to jomo object, split imputations, and exclude original data (imputation "0")
# imp.list_cont <- imputationList(split(imputed_cont, imputed_cont$Imputation)[-1])
# 
# # checks
# round(prop.table(table(imp.list_cont[[1]]$`1`$mort_28, useNA = "always"))*100,1) # first imputed dataset
# round(prop.table(table(imp.list_cont[[1]]$`2`$mort_28, useNA = "always"))*100,1) # second imputed dataset
# round(prop.table(table(df_imp_long_cont$mort_28, useNA = "always"))*100,1) # original data
# summary(imp.list_cont[[1]]$`1`$comorb_cat)
# summary(imp.list_cont[[1]]$`2`$sqsympdur)
# 
# 
# #### Add trt back, change from long to wide format, and finally combine the two data frames
# imputed_int$trt <- 1
# imputed_int_s <- imputed_int %>% # remove imputation variables, not needed anymore
#   select(trt, age, sqsympdur, mort_28, sex, comorb_cat, sqcrptrunc, clinicalstatus_baseline, clus, Imputation)
# imputed_int_wide <- imputed_int_s %>% # change from long to wide format, i.e. remove duplicates within Imputation sets
#   group_by(Imputation) %>%
#   distinct(clus, .keep_all = TRUE) 
# 
# imputed_cont$trt <- 0 # treatment variable
# imputed_cont_s <- imputed_cont %>% # remove imputation variables, not needed anymore
#   select(trt, age, sqsympdur, mort_28, sex, comorb_cat, sqcrptrunc, clinicalstatus_baseline, clus, Imputation)
# imputed_cont_wide <- imputed_cont_s %>% # change from long to wide format, i.e. remove duplicates within Imputation sets
#   group_by(Imputation) %>%
#   distinct(clus, .keep_all = TRUE) 
# 
# imputed_combined <- rbind(imputed_cont_wide, imputed_int_wide)
# 
# 
# #### Convert combined df to jomo object, split imputations, and exclude original data (imputation "0")
# imp.list <- imputationList(split(imputed_combined, imputed_combined$Imputation)[-1])
# 
# 
# ### Checks
# round(prop.table(table(imp.list[[1]]$`1`$mort_28, imp.list[[1]]$`1`$trt, useNA = "always"),2)*100,1) # first imputed dataset
# round(prop.table(table(imp.list[[1]]$`2`$mort_28, imp.list[[1]]$`2`$trt, useNA = "always"),2)*100,1) # second imputed dataset
# round(prop.table(table(df_imp$mort_28, df_imp$trt, useNA = "always"),2)*100,1) # original data
# summary(imp.list[[1]]$`1`$comorb_cat)
# summary(imp.list[[1]]$`2`$sqsympdur)

```

# (i) Primary outcome: Mortality at day 28
```{r}
# adjusted for baseline patient characteristics (age, respiratory support at baseline (ordinal scale 1-3 vs 4-5), dexamethasone use at baseline (y/n), remdesivir use at baseline (y/n), anti-IL-6 use at baseline (y/n)).
addmargins(table(df$mort_28, df$trt, useNA = "always"))
mort.28 <- df %>% 
  glm(mort_28 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.dimp <- df %>% 
  glm(mort_28_dimp ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.dimp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Create a 2x2 contingency table
# The oddsRatio function from the mosaic package calculates the odds ratio for a 2 x 2 contingency table and a confidence interval for the each estimate. x should be a matrix, data frame or table. "Successes" should be located in column 1 of x, and the treatment of interest should be located in row 2. The odds ratio is calculated as (Odds row 2) / (Odds row 1). The confidence interval is calculated from the log(OR) and back-transformed.
tbl <- tibble(
  Event = c(0.5, 1.5), # add the 0.5 correction
  NoEvent = c(58.5, 57.5) # add the 0.5 correction
)
oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
mort.28.corr <- oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
# Firth regression
mort.28.firth <- df %>% 
  logistf(mort_28 ~ trt 
      + age 
      + clinstatus_baseline 
      , data=.)
mort.28.dimp.firth <- df %>% 
  logistf(mort_28_dimp ~ trt 
      + age 
      + clinstatus_baseline 
      , data=.)
# Summary and extract coefficients
coefficients_table <- summary(mort.28.firth)$coefficients
trt_coef <- coef(mort.28.firth)["trt"]
hazard_odds_ratio <- exp(trt_coef)
ci <- c(exp(mort.28.firth$ci.lower["trt"]), exp(mort.28.firth$ci.upper["trt"]))
# Create a data frame to store Odds Ratios and CIs
mort.28.firth_tbl <- data.frame(
  # "Variable" = rownames(coefficients_table),
  "Odds Ratio" = hazard_odds_ratio,
  "CI Lower" = ci[1],
  "CI Upper" = ci[2])
# Nicely formatted table
kable(mort.28.firth_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
Discussion points
1. Rare event correction
-- Add 0.5 correction to crosstab, calculate ORs and then inverse variance pooling in second stage?
-- 2x2 directly into Mantel-Haenszel across several trials (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5297998/)?
-- Firth regression according to R.Riley IPDMA Handbook (page 99): "Traditionally 0.5 is the value added, but Sweeting et al. suggest that a ‘treatment arm’ continuity correction is more appropriate, which adds 1/(sample size of the opposite treatment group) to the number of event and non-events. In the IPD context, a similar approach is to add two extra participants to each group in a trial if it has zero events in either of the groups; one of the added participants has the event and the other does not have the event in each group. Then, a weighted regression analysis can be performed to analyse the extended IPD, with all participants weighted equally except the four added participants, who are given a weight according to Sweeting correction (i.e. 1/(sample size of the opposite treat- ment group)). However, this approach becomes problematic when adjusting for prognostic factors or extending to non-binary variables. For this reason, a more general approach is to adopt Firth regression, which is a penalisation method that reduces small sample bias for non-linear models such as logistic regression, and resolves problems related to separation. Alternatively, researchers may revert to a one-stage IPD meta-analysis approach and placing random effects on parameters (rather than stratifying parameters by trial) so that estimation of trial-specific terms are avoided."

# (i.i) Covariate adjustment for primary endpoint: Mortality at day 28
```{r warning=FALSE}
# unadjusted estimator for the (absolute) risk difference
with(df, table(trt, mort_28))
mort.28.prop.test <- prop.test(x = with(df, table(trt, mort_28)))
# print(mort.28.prop.test)
# # Estimate
-diff(mort.28.prop.test$estimate)
# # Confidence Interval
mort.28.prop.test$conf.int
# # P-Value
mort.28.prop.test$p.value

# Covariate-Adjusted Analysis
# Fit the `glm` object
# Same as Complete case analysis, substantive model // but don't use piping, otherwise problem in margins::margins
df_mort28_comp <- df %>% filter(!is.na(mort_28))
mort.28.cov.adj <-
  glm(formula = mort_28 ~ trt + age + clinstatus_baseline,
      data = df_mort28_comp,
      family = binomial(link = "logit")
      )
# Print a summary of the `glm` object
summary(mort.28.cov.adj)
# Predict Pr{Y = 1 | Z = 1, X} // equals: E(Y|Z=1,X)
pr_y1_z1 <-
  predict(
    object = mort.28.cov.adj,
    newdata =
      df_mort28_comp %>%
      dplyr::mutate(
        trt = 1
      ),
    type = "response"
  )
# Predict Pr{Y = 1 | Z = 0, X} // equals: E(Y|Z=0,X)
pr_y1_z0 <-
  predict(
    object = mort.28.cov.adj,
    newdata =
      df_mort28_comp %>%
      dplyr::mutate(
        trt = 0
      ),
    type = "response"
  )

# Estimate RD
adj_mean = mean(pr_y1_z1) - mean(pr_y1_z0)
print(adj_mean)
# Standard Error RD
# The variance/standard error can be calculted as 1/n times the sample variance of:
# Z/P(Z=1)*[Y-E(Y|Z=1,X)] + E(Y|Z=1,X) - ((1-Z)/(1-P(1=Z))*[Y-E(Y|Z=0,X)] + E(Y|Z=0,X))
p_arm = mean(df_mort28_comp$trt==1)
adj_se = sqrt(
  var((df_mort28_comp$trt==1)/p_arm * (df_mort28_comp$mort_28 - pr_y1_z1) + pr_y1_z1 -
      ((df_mort28_comp$trt==0)/(1-p_arm) * (df_mort28_comp$mort_28-pr_y1_z0) + pr_y1_z0))/
    nrow(df_mort28_comp))
print(adj_se)
# Confidence Interval
c(adj_mean-qnorm(0.975)*adj_se, adj_mean+qnorm(0.975)*adj_se)

# Or, we can obtain the standard error of the estimate two ways. The first way is using the margins::margins() command, using the robust standard errors from sandwich::vcovHC // The second way to obtain these would be the bias corrected and accelerated (BCa) non-parametric bootstrap
# You’ll see that we now have a standard error, p-value under the hypothesis that the marginal effect is 0, and a 95% Confidence Interval for the estimate. 

library(sandwich)
library(margins)
mort.28.cov.adj.ame <-
  margins::margins(
    model = mort.28.cov.adj,
    # Specify treatment variable
    variables = "trt",
    # Convert to outcome scale, not link scale
    type = "response",
    # Obtain robust standard errors
    vcov = sandwich::vcovHC(x = mort.28.cov.adj, type = "HC3")
  )
summary(object = mort.28.cov.adj.ame, level = 0.95)
mort.28.ame <- summary(object = mort.28.cov.adj.ame, level = 0.95)
# replace with manually calculated values above
mort.28.ame$AME <- adj_mean
mort.28.ame$SE <- adj_se
mort.28.ame$lower <- c(adj_mean-qnorm(0.975)*adj_se)
mort.28.ame$upper <- c(adj_mean+qnorm(0.975)*adj_se)
mort.28.ame$p <- NA

```

# (ii) Mortality at day 60
```{r}
table(df$mort_60, df$trt, useNA = "always")
mort.60 <- df %>% 
  glm(mort_60 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.60, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Create a 2x2 contingency table
# The oddsRatio function from the mosaic package calculates the odds ratio for a 2 x 2 contingency table and a confidence interval for the each estimate. x should be a matrix, data frame or table. "Successes" should be located in column 1 of x, and the treatment of interest should be located in row 2. The odds ratio is calculated as (Odds row 2) / (Odds row 1). The confidence interval is calculated from the log(OR) and back-transformed.
tbl <- tibble(
  Event = c(0.5, 1.5), # add the 0.5 correction
  NoEvent = c(58.5, 57.5) # add the 0.5 correction
)
oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
mort.60.corr <- oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
# Firth regression
mort.60.firth <- df %>% 
  logistf(mort_60 ~ trt 
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.60.firth)
```

# (iii) Time to death within max. follow-up time
```{r}
# KM curve
survfit2(Surv(death_time, death_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# # testing: cox ph
# ttdeath <- df %>% 
#   coxph(Surv(death_time, death_reached) ~ trt 
#         + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
#         , data =.)
# ttdeath_reg_tbl <- tbl_regression(ttdeath, exp = TRUE)
# # Nicely formatted table
# kable(ttdeath_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
#   kable_styling(bootstrap_options = "striped", full_width = FALSE)

library(coxphf)
ttdeath.firth <- df %>%
  coxphf(Surv(death_time, death_reached) ~ trt
        # + age
        # + clinstatus_baseline
        # + comed_dexa + comed_rdv + comed_toci
        , data =.)
summary(ttdeath.firth)

```
Discussion points
1. Result illogical. R.Riley IPDMA handbook, page 102: "As for binary outcomes, when there are few events in some trials adaptions of Firth’s correction are important to reduce small sample bias in the estimated treatment effect."

# (iv) New mechanical ventilation among survivors within 28 days
```{r}
table(df$new_mv_28, df$trt, useNA = "always")
new.mv.28 <- df %>% 
  glm(new_mv_28 ~ trt 
       + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(new.mv.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Create a 2x2 contingency table
# The oddsRatio function from the mosaic package calculates the odds ratio for a 2 x 2 contingency table and a confidence interval for the each estimate. x should be a matrix, data frame or table. "Successes" should be located in column 1 of x, and the treatment of interest should be located in row 2. The odds ratio is calculated as (Odds row 2) / (Odds row 1). The confidence interval is calculated from the log(OR) and back-transformed.
tbl <- tibble(
  Event = c(2.5, 0.5), # add the 0.5 correction, intervention second
  NoEvent = c(56.5, 57.5) # add the 0.5 correction, intervention second
)
oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
new.mv.28.corr <- oddsRatio(tbl, conf.level = 0.95, digits = 3, verbose = TRUE)
# Firth regression
new.mv.28.firth <- df %>% 
  logistf(new_mv_28 ~ trt 
      + age 
      + clinstatus_baseline
      , data=.)
summary(new.mv.28.firth)

# (iv) Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
table(df$new_mvd_28, df$trt, useNA = "always")
new.mvd.28 <- df %>% 
  glm(new_mvd_28 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(new.mvd.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# (v) Clinical status at day 28
```{r warning=FALSE}
table(df$clinstatus_28_imp, df$trt, useNA = "always")
clin.28 <- df %>% 
  clm(clinstatus_28_imp ~ trt 
      + age 
      + clinstatus_baseline
      , link= c("logit"), data=.)
# Summary and extract coefficients
coefficients_table <- summary(clin.28)$coefficients
# Calculate Odds Ratios and Confidence Intervals
odds_ratios <- exp(coefficients_table[, "Estimate"])
ci_lower <- exp(coefficients_table[, "Estimate"] - 1.96 * coefficients_table[, "Std. Error"])
ci_upper <- exp(coefficients_table[, "Estimate"] + 1.96 * coefficients_table[, "Std. Error"])
# Create a data frame to store Odds Ratios and CIs
clin.28_tbl <- data.frame(
  "Variable" = rownames(coefficients_table),
  "Odds Ratio" = odds_ratios,
  "CI Lower" = ci_lower,
  "CI Upper" = ci_upper
)
# Nicely formatted table
kable(clin.28_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# (vi) Time to discharge or reaching discharge criteria up to day 28
```{r warning=FALSE}
# Kaplan-Meier estimate of conditional discharge probability
# Just censoring => Cause-specific hazards
km.ttdischarge.check <- with(df, Surv(discharge_time, discharge_reached))
head(km.ttdischarge.check, 100)
km.ttdischarge_trt <- survfit(Surv(discharge_time, discharge_reached) ~ trt, data=df)
# KM curve
survfit2(Surv(discharge_time, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge <- df %>% 
  coxph(Surv(discharge_time, discharge_reached) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_reg_tbl <- tbl_regression(ttdischarge, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Sub-distribution hazards, i.e., represents the rate per unit of time of the event as well as the influence of competing events. Instantaneous rate of occurrence of the given type of event in subjects who have not yet experienced an event of that type.
df <- df %>% # cuminc needs a factor variable with censored patients coded as 0, the event as 1 and the competing event as 2.
  mutate(discharge_reached_comp = case_when (discharge_reached == 0 & (mort_28 == 0 | is.na(mort_28)) ~ 0,
                                             discharge_reached == 1 & (mort_28 == 0 | is.na(mort_28)) ~ 1,
                                             mort_28 == 1 ~ 2))
df$discharge_reached_comp <- as.factor(df$discharge_reached_comp)
table(df$discharge_reached_comp)
# Cumulative incidence for competing risks
cuminc(Surv(discharge_time, discharge_reached_comp) ~ 1, data = df)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df) %>%
  ggcuminc(outcome = c("1", "2")) +
  ylim(c(0, 1)) +
  labs(
    x = "Days"
  ) +
  add_confidence_interval() +
  add_risktable()
# in int only
df_int <- df %>% 
  filter(trt == 1)
cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_int) %>% 
  ggcuminc(outcome = c("1", "2")) +
  #ylim(c(0, 1)) + 
  labs(
    x = "Days"
  ) + 
  add_confidence_interval() +
  add_risktable()
# in cont only
# df_cont <- df %>% 
#   filter(trt == 0)
# cuminc(Surv(discharge_time, discharge_reached_comp) ~ trt, data = df_cont) %>% 
#   ggcuminc(outcome = c("1", "2")) +
#   #ylim(c(0, 1)) + 
#   labs(
#     x = "Days"
#   ) + 
#   add_confidence_interval() +
#   add_risktable()
# testing: Fine-Gray regression
ttdischarge.comp <- crr(Surv(discharge_time, discharge_reached_comp) ~ trt
    + age
    #+ clinstatus_baseline
    ,data = df)
ttdischarge_comp_reg_tbl <- tbl_regression(ttdischarge.comp, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_comp_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Censoring and assigned worst outcome (28d) to competing event (death) // hypothetical estimand
survfit2(Surv(discharge_time_sens, discharge_reached) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sens <- df %>% 
  coxph(Surv(discharge_time_sens, discharge_reached) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_sens_reg_tbl <- tbl_regression(ttdischarge.sens, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sens_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days
# Use cause-specific hazards
km.ttdischarge_sus_trt <- survfit(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df)
ttdischarge_sus_28d_tbl <- km.ttdischarge_sus_trt %>% 
  tbl_survfit(
    times = 28,
    label_header = "**28-d sustained hospitalization (95% CI)**"
  )
# Nicely formatted table
kable(ttdischarge_sus_28d_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
#KM curve
survfit2(Surv(discharge_time_sus, discharge_reached_sus) ~ trt, data=df) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall sustained hospitalization probability"
  ) + 
  add_confidence_interval() +
  add_risktable()
# testing: cox ph
ttdischarge.sus <- df %>% 
  coxph(Surv(discharge_time_sus, discharge_reached_sus) ~ trt 
        + age + clinstatus_baseline
        , data =.)
ttdischarge_sus_reg_tbl <- tbl_regression(ttdischarge.sus, exp = TRUE)
# Nicely formatted table
kable(ttdischarge_sus_reg_tbl, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# (vii) Viral clearance up to day 5, day 10, and day 15
```{r warning=FALSE}

```
Discussion points
1. Not available

# (viii) Quality of life at day 28 
```{r warning=FALSE}

```
Discussion points
1. Not available

# (ix) Adverse event(s) grade 3 or 4, or a serious adverse event(s), excluding death, by day 28
```{r warning=FALSE}
table(df$ae_28, df$trt, useNA = "always") 
ae.28 <- df %>% 
  glm(ae_28 ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(ae.28, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
table(df$ae_28_sev, df$trt, useNA = "always") 
ae.28.sev <- df %>% 
  glm(ae_28_sev ~ trt 
      + age + clinstatus_baseline
      , family = "binomial", data=.)
summ(ae.28.sev, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)

```

# Subgroup analysis: Ventilation requirement (proxy for disease severity) on primary endpoint
```{r warning=FALSE}
table(df$clinstatus_baseline, df$mort_28, useNA = "always") # only 2 - 3 included
table(df$vbaseline, df$mort_28, useNA = "always")
# class(df$clinstatus_baseline)
df$clinstatus_baseline_n <- as.numeric(df$clinstatus_baseline)

mort.28.vent <- df %>% 
  glm(mort_28 ~ trt*clinstatus_baseline_n
      + age 
      #+ clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.vent, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth regression
mort.28.vent.firth <- df %>% 
  logistf(mort_28 ~ trt*clinstatus_baseline_n
      + age 
      #+ clinstatus_baseline
      , data=.)
summary(mort.28.vent.firth)
# vbaseline
# mort.28.vent.vb.firth <- df %>%
#   logistf(mort_28 ~ trt*vbaseline
#       + age
#       #+ clinstatus_baseline
#       #+ comed_dexa + comed_rdv + comed_toci
#       , data=.)
# summary(mort.28.vent.vb.firth)

# effect by subgroup
mort.28.vent.vb.no <- df %>%
  filter(vbaseline == 0) %>% # not ventilated
  logistf(mort_28 ~ trt
      + age
     # + clinstatus_baseline
      , data=.)
summary(mort.28.vent.vb.no)

# class(df$clinstatus_baseline)
mort.28.vent.rs.2 <- df %>%
  filter(clinstatus_baseline == "2") %>% # no oxygen
  logistf(mort_28 ~ trt
      + age
     # + clinstatus_baseline
      , data=.)
summary(mort.28.vent.rs.2)

mort.28.vent.rs.3 <- df %>%
  filter(clinstatus_baseline == "3") %>% # LF oxygen
  logistf(mort_28 ~ trt
      + age
     # + clinstatus_baseline
      , data=.)
summary(mort.28.vent.rs.3)

```
Discussion points
1. Firth regression works for clinstatus_baseline but not for vbaseline

# Subgroup analysis: Age on primary endpoint
```{r warning=FALSE}
mort.28.age <- df %>% 
  glm(mort_28 ~ trt*age
      # + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.age, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth regression
mort.28.age.firth <- df %>% 
  logistf(mort_28 ~ trt*age
      #+ age 
      + clinstatus_baseline 
      , data=.)
summary(mort.28.age.firth)

# effect by subgroup // adapted to cut-off 65, see deft plot!
df <- df %>% 
  mutate(age_70 = case_when(age < 65 ~ 0,
                            age > 64 ~ 1))

mort.28.age.a70.firth <- df %>%
  filter(age_70 == 1) %>% # 70 and above
  logistf(mort_28 ~ trt
     # + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.age.a70.firth)
tab_model(mort.28.age.a70.firth)

mort.28.age.b70.firth <- df %>% 
  filter(age_70 == 0) %>% # below 70
  logistf(mort_28 ~ trt
      # + age 
      + clinstatus_baseline 
      , data=.)
summary(mort.28.age.b70.firth)
tab_model(mort.28.age.b70.firth)
```

# Subgroup analysis: Comorbidities on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction
table(df$comorb_cat, df$mort_28, useNA = "always") 
# class(df$comorb_cat)
mort.28.comorb <- df %>%
  glm(mort_28 ~ trt*comorb_cat 
      #+ age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# table(df$comorb_cat, df$mort_28, df$trt, useNA = "always") ### too few events!
# Firth
mort.28.comorb.firth <- df %>%
  logistf(mort_28 ~ trt*comorb_cat
      + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.firth)
# 4 comorbidity categories as factor
df$comorb_cat_f <- as.factor(df$comorb_cat)
# table(df$comorb_cat_f, df$mort_28, useNA = "always") 
mort.28.comorb.f <- df %>% 
  glm(mort_28 ~ trt*comorb_cat_f 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.f, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# full comorbidity count
# table(df$comorb_count, df$mort_28, useNA = "always") 
mort.28.comorb.count <- df %>%
  glm(mort_28 ~ trt*comorb_count 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comorb.count, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
mort.28.comorb.count.firth <- df %>%
  logistf(mort_28 ~ trt*comorb_count 
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.count.firth)
# any comorbidity
# table(df$comorb_any, df$mort_28, useNA = "always") 
mort.28.comorb.any.firth <- df %>%
  logistf(mort_28 ~ trt*comorb_any 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summary(mort.28.comorb.any.firth)
# comorbidities without immunosuppressed
# table(df$comorb_cat, df$mort_28, useNA = "always")
df <- df %>% 
  mutate(comorb_noimmuno = case_when(comorb_cat == 4 ~ NA,
                              TRUE ~ comorb_cat))
# table(df$comorb_noimmuno, df$mort_28, useNA = "always")
mort.28.comorb.noimmuno.firth <- df %>%
  logistf(mort_28 ~ trt*comorb_noimmuno 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summary(mort.28.comorb.noimmuno.firth)


# effect by subgroup
mort.28.comorb.1 <- df %>% 
  filter(comorb_cat == 1) %>% # no comorbidity
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.1)

mort.28.comorb.2 <- df %>% 
  filter(comorb_cat == 2) %>% # 1 comorbidity
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.2)

mort.28.comorb.3 <- df %>% 
  filter(comorb_cat == 3) %>% # multiple comorbidities
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comorb.3)

# mort.28.comorb.4 <- df %>%
#   filter(comorb_cat == 4) %>% # immunocompromised
#   logistf(mort_28 ~ trt
#      # + age
#      # + clinstatus_baseline
#       , data=.)
# summ(mort.28.comorb.4, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# Subgroup analysis: Concomitant COVID-19 treatment on primary endpoint
```{r warning=FALSE}
# 4 comorbidity categories as numeric/continuous, i.e., linear interaction

table(df$comed_cat, df$trt, useNA = "always")
# 1: patients without Dexamethasone nor Tocilizumab => JAKi effect alone
# 2: patients with Dexamethasone but no Tocilizumab => JAKi effect with Dexa only
# 3: patients with Dexamethasone and Tocilizumab => JAKi effect with Dexa + Toci
# 4: patients with Tocilizumab but no Dexamethasone (if exist) => JAKi effect with Toci only 
mort.28.comed <- df %>%
  glm(mort_28 ~ trt*comed_cat 
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.comed, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth
mort.28.comed.firth <- df %>%
  logistf(mort_28 ~ trt*comed_cat
      + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.comed.firth)

# effect by subgroup
mort.28.comed.1 <- df %>% 
  filter(comed_cat == 1) %>% # without Dexamethasone nor Tocilizumab
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comed.1)

mort.28.comed.2 <- df %>% 
  filter(comed_cat == 2) %>% # with Dexamethasone but no Tocilizumab
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.comed.2)
```

# Subgroup analysis: Vaccination on adverse events
```{r warning=FALSE}
table(df$vacc, df$ae_28, useNA = "always") # none of the vaccinated had an adverse event
ae.28.vacc <- df %>% 
  glm(ae_28 ~ trt*vacc
      #+ age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(ae.28.vacc, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth
ae.28.vacc.firth <- df %>%
  logistf(ae_28 ~ trt*vacc
      + age
      + clinstatus_baseline
      , data=.)
summary(ae.28.vacc.firth)

# effect by subgroup
# class(df$vacc)
ae.28.vacc.1 <- df %>%
  filter(vacc == 1) %>% # vaccinated
  logistf(ae_28 ~ trt
      + age
      + clinstatus_baseline
      , data=.)
summary(ae.28.vacc.1)

ae.28.vacc.0 <- df %>% 
  filter(vacc == 0) %>% # not vaccinated
  glm(ae_28 ~ trt
      + age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(ae.28.vacc.0, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
```

# SENS Subgroup analysis: Duration since symptom onset on primary endpoint
```{r warning=FALSE}
# table(df$sympdur, df$mort_28, useNA = "always")
mort.28.symp <- df %>% 
  glm(mort_28 ~ trt*sympdur
      #+ age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.symp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth
mort.28.symp.firth <- df %>%
  logistf(mort_28 ~ trt*sympdur
      + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.symp.firth)

# effect by subgroup
df <- df %>% 
  mutate(sympdur_cat = case_when(sympdur < 6 ~ 2,
                                 sympdur > 5 & sympdur < 11 ~ 1,
                                 sympdur > 10 ~ 0))
table(df$sympdur_cat, useNA = "always")
# table(df$sympdur, useNA = "always")
mort.28.sympdur.a10 <- df %>%
  filter(sympdur_cat == 0) %>% # more than 10 days
  logistf(mort_28 ~ trt
      + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.sympdur.a10)

mort.28.sympdur.510 <- df %>% 
  filter(sympdur_cat == 1) %>% # 5-10 days
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.sympdur.510)

mort.28.sympdur.b5 <- df %>% 
  filter(sympdur_cat == 2) %>% # 5d or less
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.sympdur.b5)
```

# SENS Subgroup analysis: CRP on primary endpoint
```{r warning=FALSE}
# table(df$crp, df$mort_28, useNA = "always")
mort.28.crp <- df %>% 
  glm(mort_28 ~ trt*crp
      #+ age 
      + clinstatus_baseline
      , family = "binomial", data=.)
summ(mort.28.crp, exp = T, confint = T, model.info = T, model.fit = F, digits = 2)
# Firth
mort.28.crp.firth <- df %>%
  logistf(mort_28 ~ trt*crp
      + age
      + clinstatus_baseline
      , data=.)
summary(mort.28.crp.firth)

# effect by subgroup
df <- df %>% 
  mutate(crp_75 = case_when(crp < 75 ~ 1,
                            crp > 74 ~ 0))
# table(df$crp_75, useNA = "always")
mort.28.crp.b75 <- df %>% 
  filter(crp_75 == 1) %>% # below 75
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.crp.b75)

mort.28.crp.a75 <- df %>% 
  filter(crp_75 == 0) %>% # 75 and above
  logistf(mort_28 ~ trt
      + age 
      + clinstatus_baseline
      , data=.)
summary(mort.28.crp.a75)
```

# SENS Subgroup analysis: Variant on primary endpoint
```{r warning=FALSE}

```
Discussion points
1. No data on variant available

# Collect all treatment effect estimates across endpoints (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
result_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric(),
  n_int = numeric(),
  n_cont = numeric(),
  e_int = numeric(),
  e_cont = numeric()
)
# Function to extract treatment results from different model types (glm, clm, coxph, crr, logistf)
extract_trt_results <- function(model, variable_name, n_int, n_cont, e_int, e_cont) {
  if (inherits(model, "glm") || inherits(model, "clm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "logistf") || inherits(model, "coxphf")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- c(exp(model$ci.lower["trt"]), exp(model$ci.upper["trt"]))
    se <- sqrt(diag(vcov(model)))["trt"]
    p_value <- model$prob["trt"]
  } else if (inherits(model, "coxph")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "se(coef)"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "tidycrr")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- c(exp(model$tidy$conf.low[1]), exp(model$tidy$conf.high[1]))
    se <- model$tidy$std.error[1]
    p_value <- model$tidy$p.value[1]
  } else if (inherits(model, "summary.margins")) {
    hazard_odds_ratio <- model$AME ### CAVE: this is not an HR or OR, but a marginal RD
    ci <- c(model$lower, model$upper)
    se <- model$SE
    p_value <- model$p
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_int = n_int,
    n_cont = n_cont,
    e_int = e_int,
    e_cont = e_cont
  )
  return(result)
}

# Loop through
result_list <- list()

result_list[[1]] <- extract_trt_results(mort.28.firth, "death at day 28_firth",
                                        addmargins(table(df$mort_28, df$trt))[3,2], 
                                        addmargins(table(df$mort_28, df$trt))[3,1],
                                        addmargins(table(df$mort_28, df$trt))[2,2],
                                        addmargins(table(df$mort_28, df$trt))[2,1])
result_list[[2]] <- extract_trt_results(mort.28.dimp.firth, "death at day 28_dimp_firth",
                                        addmargins(table(df$mort_28_dimp, df$trt))[3,2], 
                                        addmargins(table(df$mort_28_dimp, df$trt))[3,1],
                                        addmargins(table(df$mort_28_dimp, df$trt))[2,2], 
                                        addmargins(table(df$mort_28_dimp, df$trt))[2,1])
result_list[[3]] <- extract_trt_results(mort.28.ame, "death at day 28_marginal",
                                        addmargins(table(df$mort_28, df$trt))[3,2], 
                                        addmargins(table(df$mort_28, df$trt))[3,1],
                                        addmargins(table(df$mort_28, df$trt))[2,2], 
                                        addmargins(table(df$mort_28, df$trt))[2,1])
result_list[[4]] <- extract_trt_results(mort.60.firth, "death at day 60_firth",
                                        addmargins(table(df$mort_60, df$trt))[3,2], 
                                        addmargins(table(df$mort_60, df$trt))[3,1],
                                        addmargins(table(df$mort_60, df$trt))[2,2], 
                                        addmargins(table(df$mort_60, df$trt))[2,1])
result_list[[5]] <- extract_trt_results(ttdeath.firth, "death within fup_firth",
                                        addmargins(table(df$death_reached, df$trt))[3,2], 
                                        addmargins(table(df$death_reached, df$trt))[3,1],
                                        addmargins(table(df$death_reached, df$trt))[2,2], 
                                        addmargins(table(df$death_reached, df$trt))[2,1])
result_list[[6]] <- extract_trt_results(new.mv.28.firth, "new MV within 28d_firth",
                                        addmargins(table(df$new_mv_28, df$trt))[3,2], 
                                        addmargins(table(df$new_mv_28, df$trt))[3,1],
                                        addmargins(table(df$new_mv_28, df$trt))[2,2], 
                                        addmargins(table(df$new_mv_28, df$trt))[2,1])
result_list[[7]] <- extract_trt_results(new.mvd.28, "new MV or death within 28d",
                                        addmargins(table(df$new_mvd_28, df$trt))[3,2], 
                                        addmargins(table(df$new_mvd_28, df$trt))[3,1],
                                        addmargins(table(df$new_mvd_28, df$trt))[2,2], 
                                        addmargins(table(df$new_mvd_28, df$trt))[2,1])
result_list[[8]] <- extract_trt_results(clin.28, "clinical status at day 28",
                                        addmargins(table(df$clinstatus_28_imp, df$trt))[7,2], 
                                        addmargins(table(df$clinstatus_28_imp, df$trt))[7,1],
                                        NA,
                                        NA)
result_list[[9]] <- extract_trt_results(ttdischarge, "discharge within 28 days",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[3,1],
                                        addmargins(table(df$discharge_reached, df$trt))[2,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[2,1])
result_list[[10]] <- extract_trt_results(ttdischarge.comp, "discharge within 28 days, death=comp.event",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[3,1],
                                        addmargins(table(df$discharge_reached, df$trt))[2,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[2,1])
result_list[[11]] <- extract_trt_results(ttdischarge.sens, "discharge within 28 days, death=hypo.event",
                                        addmargins(table(df$discharge_reached, df$trt))[3,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[3,1],
                                        addmargins(table(df$discharge_reached, df$trt))[2,2], 
                                        addmargins(table(df$discharge_reached, df$trt))[2,1])
result_list[[12]] <- extract_trt_results(ttdischarge.sus, "sustained discharge within 28 days",
                                         addmargins(table(df$discharge_reached_sus, df$trt))[3,2], 
                                         addmargins(table(df$discharge_reached_sus, df$trt))[3,1],
                                         addmargins(table(df$discharge_reached_sus, df$trt))[2,2], 
                                         addmargins(table(df$discharge_reached_sus, df$trt))[2,1])
result_list[[13]] <- extract_trt_results(ae.28, "any AE grade 3,4 within 28 days",
                                         addmargins(table(df$ae_28, df$trt))[3,2], 
                                         addmargins(table(df$ae_28, df$trt))[3,1],
                                         addmargins(table(df$ae_28, df$trt))[2,2], 
                                         addmargins(table(df$ae_28, df$trt))[2,1])
result_list[[14]] <- extract_trt_results(ae.28.sev, "AEs grade 3,4 within 28 days",
                                         addmargins(table(df$ae_28_sev, df$trt))[3,2], 
                                         addmargins(table(df$ae_28_sev, df$trt))[3,1],
                                         NA,
                                         NA)

# Filter out NULL results and bind the results into a single data frame
result_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
result_df$trial <- "TOFACOV"
result_df$JAKi <- "Tofacitinib"

# ### Add the rare event results
# ## Add the results from the 0.5-corrected models
# # mort.28.corr
# # summary(mort.28.corr)
# hazard_odds_ratio <- mort.28.corr[["OR"]]
# ci_lower <- c(0.1218) # how to extract directly from the object?
# ci_upper <- c(76.48) # how to extract directly from the object?
# standard_error <- (ci_upper - ci_lower) / (2 * 1.96)
# z_score <- (log(hazard_odds_ratio) - log(1)) / standard_error
# p_value <- 2 * (1 - pnorm(abs(z_score)))
# new_row <- data.frame(
#     variable = "death at day 28_0.5-corr",
#     hazard_odds_ratio = hazard_odds_ratio,
#     ci_lower = ci_lower,
#     ci_upper = ci_upper,
#     standard_error = standard_error,
#     p_value = p_value,
#     n_intervention = addmargins(table(df$mort_28, df$trt))[3,2], 
#     n_control = addmargins(table(df$mort_28, df$trt))[3,1],
#     trial = "TOFACOV",
#     JAKi = "Tofacitinib")
# result_df <- rbind(result_df, new_row) # no adj
# 
# # mort.60.corr
# # summary(mort.60.corr)
# hazard_odds_ratio <- mort.60.corr[["OR"]]
# ci_lower <- c(0.1218) # how to extract directly from the object?
# ci_upper <- c(76.48) # how to extract directly from the object?
# standard_error <- (ci_upper - ci_lower) / (2 * 1.96)
# z_score <- (log(hazard_odds_ratio) - log(1)) / standard_error
# p_value <- 2 * (1 - pnorm(abs(z_score)))
# new_row <- data.frame(
#     variable = "death at day 60_0.5-corr",
#     hazard_odds_ratio = hazard_odds_ratio,
#     ci_lower = ci_lower,
#     ci_upper = ci_upper,
#     standard_error = standard_error,
#     p_value = p_value,
#     n_intervention = addmargins(table(df$mort_60, df$trt))[3,2], 
#     n_control = addmargins(table(df$mort_60, df$trt))[3,1],
#     trial = "TOFACOV",
#     JAKi = "Tofacitinib")
# result_df <- rbind(result_df, new_row) # no adj
# 
# # new.mv.28.corr
# # summary(new.mv.28.corr)
# hazard_odds_ratio <- new.mv.28.corr[["OR"]]
# ci_lower <- c(0.009228) # how to extract directly from the object?
# ci_upper <- c(4.185) # how to extract directly from the object?
# standard_error <- (ci_upper - ci_lower) / (2 * 1.96)
# z_score <- (log(hazard_odds_ratio) - log(1)) / standard_error
# p_value <- 2 * (1 - pnorm(abs(z_score)))
# new_row <- data.frame(
#     variable = "new MV within 28d_0.5-corr",
#     hazard_odds_ratio = hazard_odds_ratio,
#     ci_lower = ci_lower,
#     ci_upper = ci_upper,
#     standard_error = standard_error,
#     p_value = p_value,
#     n_intervention = addmargins(table(df$new_mv_28, df$trt))[3,2], 
#     n_control = addmargins(table(df$new_mv_28, df$trt))[3,1],
#     trial = "TOFACOV",
#     JAKi = "Tofacitinib")
# result_df <- rbind(result_df, new_row) # no adj

# Nicely formatted table
kable(result_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(result_df, file = "trt_effects_tofacov.RData")
```

# Collect all interaction estimates (stage one)
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
interaction_df <- data.frame(
  variable = character(),
  log_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric()
)
# Extract and format results for the interaction term
extract_interaction <- function(model, variable_name) {
  if (inherits(model, "glm") || inherits(model, "clm")) {
      trt_coef <- coef(model)[grep("^trt:", names(coef(model)))]
      log_odds_ratio <- exp(trt_coef)
      ci <- exp(confint(model)[grep("^trt:", names(coef(model))), ])
      se <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Std. Error"]
      p_value <- summary(model)$coefficients[grep("^trt:", names(coef(model))), "Pr(>|z|)"]
  } else if (inherits(model, "logistf")) {
      trt_coef <- coef(model)[grep("^trt:", names(coef(model)))]
      log_odds_ratio <- exp(trt_coef)
      ci <- exp(confint(model)[grep("^trt:", names(coef(model))), ])
      se <- sqrt(diag(vcov(model)))[grep("^trt:", names(coef(model)))]
      p_value <- model$prob[grep("^trt:", names(coef(model)))]
  } else {
    stop("Unsupported model class")
  }
      # capture the results
      result <- data.frame(
        variable = variable_name,
        log_odds_ratio = log_odds_ratio,
        ci_lower = ci[1],
        ci_upper = ci[2],
        standard_error = se,
        p_value = p_value
      )
    return(result)
}
# Loop through
result_list <- list()

result_list[[1]] <- extract_interaction(mort.28.vent.firth, "respiratory support_firth")
# result_list[[2]] <- extract_interaction(mort.28.vent.vb.firth, "ventilation_firth") # does not converge
result_list[[3]] <- extract_interaction(mort.28.age.firth, "age_firth")
result_list[[4]] <- extract_interaction(mort.28.comorb.firth, "comorbidity_firth")
result_list[[5]] <- extract_interaction(mort.28.comorb.count.firth, "comorbidity_count_firth")
result_list[[6]] <- extract_interaction(mort.28.comorb.any.firth, "comorbidity_any_firth")
result_list[[7]] <- extract_interaction(mort.28.comorb.noimmuno.firth, "comorbidity_noimmuno_firth")
result_list[[8]] <- extract_interaction(mort.28.comed.firth, "comedication_firth")
result_list[[9]] <- extract_interaction(ae.28.vacc.firth, "vaccination on AEs_firth")
result_list[[10]] <- extract_interaction(mort.28.symp.firth, "symptom duration_firth")
result_list[[11]] <- extract_interaction(mort.28.crp.firth, "crp_firth")

# Filter out NULL results and bind the results into a single data frame
interaction_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
interaction_df$trial <- "TOFACOV"
interaction_df$JAKi <- "Tofacitinib"

# Nicely formatted table
kable(interaction_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(interaction_df, file = "int_effects_tofacov.RData")
```

# Collect all subgroup treatment effect estimates
```{r message=FALSE, warning=FALSE}
# Empty data frame to store the results
subgroup_df <- data.frame(
  variable = character(),
  hazard_odds_ratio = numeric(),
  ci_lower = numeric(),
  ci_upper = numeric(),
  standard_error = numeric(),
  p_value = numeric(),
  n_intervention = numeric(),
  n_intervention_tot = numeric(),
  n_control = numeric(),
  n_control_tot = numeric()
)

# Function to extract subgroup treatment results
extract_subgroup_results <- function(model, variable_name, n_int, n_int_tot, n_cont, n_cont_tot) {
  if (inherits(model, "glm")) {
    trt_coef <- coef(model)["trt"]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)["trt", ])
    se <- summary(model)$coefficients["trt", "Std. Error"]
    p_value <- summary(model)$coefficients["trt", "Pr(>|z|)"]
  } else if (inherits(model, "logistf")) {
    trt_coef <- coef(model)[grep("^trt", names(coef(model)))]
    hazard_odds_ratio <- exp(trt_coef)
    ci <- exp(confint(model)[grep("^trt", names(coef(model))), ])
    se <- sqrt(diag(vcov(model)))[grep("^trt", names(coef(model)))]
    p_value <- model$prob[grep("^trt", names(coef(model)))]
  } else {
    stop("Unsupported model class")
  }
  # capture the results
  result <- data.frame(
    variable = variable_name,
    hazard_odds_ratio = hazard_odds_ratio,
    ci_lower = ci[1],
    ci_upper = ci[2],
    standard_error = se,
    p_value = p_value,
    n_intervention = n_int,
    n_intervention_tot = n_int_tot,
    n_control = n_cont,
    n_control_tot = n_cont_tot
  )
  return(result)
}

# Loop through
result_list <- list()

# result_list[[1]] <- extract_subgroup_results(mort.28.vent.vb.yes, "High-flow or non-invasive, mechanical, or ECMO",
#                                              addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,2,2], 
#                                              addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,3,2], 
#                                              addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,2,1], 
#                                              addmargins(table(df$vbaseline, df$mort_28, df$trt))[2,3,1]) 
result_list[[2]] <- extract_subgroup_results(mort.28.vent.vb.no, "None or low-flow oxygen_firth",
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$vbaseline, df$mort_28, df$trt))[1,3,1])
result_list[[3]] <- extract_subgroup_results(mort.28.vent.rs.2, "No oxygen_firth",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,2,2],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,3,2],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,2,1],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[2,3,1])
result_list[[4]] <- extract_subgroup_results(mort.28.vent.rs.3, "low-flow oxygen_firth",
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,2,2],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,3,2],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,2,1],
                                             addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[3,3,1])
# result_list[[5]] <- extract_subgroup_results(mort.28.vent.rs.4, "high-flow oxygen / NIV",
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,2,2], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,3,2], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,2,1], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[4,3,1])
# result_list[[6]] <- extract_subgroup_results(mort.28.vent.rs.5, "Mechanical ventilation / ECMO",
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,2,2], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,3,2], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,2,1], 
#                                              addmargins(table(df$clinstatus_baseline, df$mort_28, df$trt))[5,3,1]) 
result_list[[7]] <- extract_subgroup_results(mort.28.age.a70.firth, "70 years and above_firth",
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,2,2],
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,3,2],
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,2,1],
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[2,3,1])
result_list[[8]] <- extract_subgroup_results(mort.28.age.b70.firth, "below 70 years_firth",
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$age_70, df$mort_28, df$trt))[1,3,1]) 
result_list[[9]] <- extract_subgroup_results(mort.28.comorb.1, "No comorbidity_firth",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[1,3,1])
result_list[[10]] <- extract_subgroup_results(mort.28.comorb.2, "One comorbidity_firth",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[2,3,1])
result_list[[11]] <- extract_subgroup_results(mort.28.comorb.3, "Multiple comorbidities_firth",
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,2,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,3,2], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,2,1], 
                                             addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[3,3,1])
# result_list[[12]] <- extract_subgroup_results(mort.28.comorb.4, "Immunocompromised",
#                                              addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,2,2], 
#                                              addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,3,2], 
#                                              addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,2,1], 
#                                              addmargins(table(df$comorb_cat_f, df$mort_28, df$trt))[4,3,1]) 
result_list[[13]] <- extract_subgroup_results(mort.28.comed.1, "No Dexa, no Tocilizumab_firth",
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[1,3,1])
result_list[[14]] <- extract_subgroup_results(mort.28.comed.2, "Dexa, but no Tocilizumab_firth",
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$comed_cat, df$mort_28, df$trt))[2,3,1])
result_list[[15]] <- extract_subgroup_results(ae.28.vacc.1, "vaccinated_firth",
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,2,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,3,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,2,1],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[2,3,1])
result_list[[16]] <- extract_subgroup_results(ae.28.vacc.0, "not vaccinated",
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$vacc, df$mort_28, df$trt))[1,3,1])
result_list[[17]] <- extract_subgroup_results(mort.28.sympdur.a10, "More than 10 days_firth",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,2,2],
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,3,2],
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,2,1],
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[1,3,1])
result_list[[18]] <- extract_subgroup_results(mort.28.sympdur.510, "Between 5-10 days_firth",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[2,3,1])
result_list[[19]] <- extract_subgroup_results(mort.28.sympdur.b5, "5 days and less_firth",
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,2,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,3,2], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,2,1], 
                                             addmargins(table(df$sympdur_cat, df$mort_28, df$trt))[3,3,1])
result_list[[20]] <- extract_subgroup_results(mort.28.crp.a75, "CRP 75 and higher_firth",
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,2,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,3,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,2,1], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[1,3,1])
result_list[[21]] <- extract_subgroup_results(mort.28.crp.b75, "CRP below 75_firth",
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,2,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,3,2], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,2,1], 
                                             addmargins(table(df$crp_75, df$mort_28, df$trt))[2,3,1])

# Filter out NULL results and bind the results into a single data frame
subgroup_df <- do.call(rbind, Filter(function(x) !is.null(x), result_list))

# Add the trial name and JAKi
subgroup_df$trial <- "TOFACOV"
subgroup_df$JAKi <- "Tofacitinib"

# Nicely formatted table
kable(subgroup_df, format = "markdown", table.attr = 'class="table"') %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Save
saveRDS(subgroup_df, file = "subgroup_effects_tofacov.RData")
```
