---
title: "TOFACOV"
author: "A.Amstutz"
date: "2023-08-30"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

# Load packages
```{r load packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(tableone)

library(jtools) # for summ() and plot_summs
library(sjPlot) # for tab_model
library(ggsurvfit) # survival/TTE analyses
library(survival) # survival/TTE analyses
library(ggplot2) # survival/TTE analyses and other graphs
library(ggfortify) # autoplot

```

# Load Data
```{r, include=FALSE}
df <- read_xlsx("/Users/amstutzal/Library/CloudStorage/OneDrive-usb.ch/Dokumente - JAKi IPDMA data source management/General/TOFACOV/JAK inhibitor metaanalysis TOFACOV2 NEW.xlsx")

```

# Baseline Characteristics
```{r echo=TRUE}
df <- df %>%
  rename(id_pat = ID)
df <- df %>% # no missing in 
  mutate(sex = case_when(sex == 1 ~ "female",
                         sex == 2 ~ "male"))
# no missing in sex and age and ethnicity, no duplicates in ID, trt according to publication, add country
df$country <- c("Italy")
df$icu <- 0 # no icu patients at enrolment

# Days with symptoms prior to randomization
df %>% # no missings
  drop_na(sympdur) %>% 
  ggplot(aes(x = sympdur)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of Symptom Duration",
       x = "Symptom Duration",
       y = "Density")

# Severity of COVID-19 with respect to respiratory support at randomisation // tofacov only recruited mild (2) and moderate (3) covid-19, noone with any kind of ventilation (4 above)
df$clinstatus_baseline <- factor(df$clinstatus_baseline, levels = 1:6) ## no missing data
addmargins(table(df$clinstatus_baseline, df$trt, useNA = "always"))

# Co-medication at baseline
df$comed_toci <- 0 # see publication: "It is noteworthy that no patients have received tocili- zumab. Concomitant use of tofacitinib and tocilizumab was prohibited by protocol"
df$comed_interferon <- 0 # no interferon used

## group them for the subgroup analysis, according to protocol
df <- df %>% 
  mutate(comed_cat = case_when(comed_dexa == 0 & comed_toci == 0 ~ 1, # patients without Dexamethasone nor Tocilizumab
                               comed_dexa == 1 & comed_toci == 1 ~ 2, # patients with Dexamethasone and Tocilizumab
                               comed_dexa == 1 & comed_toci == 0 ~ 3, # patients with Dexamethasone but no Tocilizumab
                               comed_dexa == 0 & comed_toci == 1 ~ 4)) # patients with Tocilizumab but no Dexamethasone (if exist)

# Comorbidity at baseline, including immunocompromised // no immunosupp in tofacov
df <- df %>% # no missing
  mutate(any_comorb = case_when(comorb_lung == 1 | comorb_liver == 1 | comorb_cvd == 1 |
                                  comorb_aht == 1 | comorb_dm == 1 | comorb_obese == 1 | comorb_smoker == 1
                                | immunosupp == 1
                                  ~ 1,
                                comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0
                                ~ 0))
## group them for the subgroup analysis, according to protocol // count all pre-defined comorbidities per patient first
comorb <- df %>% 
  select(id_pat, comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp)
comorb$comorb_count <- NA
for (i in 1:dim(comorb)[[1]]) {
  comorb$comorb_count[i] <- ifelse(
    sum(comorb[i, ] %in% c(1)) > 0,
    sum(comorb[i, ] %in% c(1)),
    NA
  )
}
comorb <- comorb %>% 
  mutate(comorb_count = case_when(comorb_lung == 0 & comorb_liver == 0 & comorb_cvd == 0 &
                                  comorb_aht == 0 & comorb_dm == 0 & comorb_obese == 0 & comorb_smoker == 0
                                & immunosupp == 0 ~ 0,
                                TRUE ~ comorb_count))
df <- left_join(df, comorb[, c("comorb_count", "id_pat")], by = join_by(id_pat == id_pat)) ## merge imputed variable back
df <- df %>% # no missing
  mutate(comorb_cat = case_when(immunosupp == 1 ~ 4, # immunocompromised
                                comorb_count == 0 ~ 1, # no comorbidity
                                comorb_count == 1 ~ 2, # one comorbidity
                                comorb_count >1 & (immunosupp == 0 | is.na(immunosupp)) ~ 3)) # multiple comorbidities

# CRP
df$crp <- as.numeric(df$`crp mg/L`) ## 1 missing
df %>% 
  drop_na(crp) %>% 
  ggplot(aes(x = crp)) +
  geom_density(fill = "blue", color = "black") +
  labs(title = "Density Plot of CRP",
       x = "CRP",
       y = "Density")

# Vaccination
# table(df$vacc, useNA = "always") # no missing

# Viremia
# Variant
# Serology

```
Discussion points


# Endpoints
```{r echo=TRUE}
# time to event data
df$death_d <- as.numeric(df$death_date - df$randdate) 
df$discharge_d <- as.numeric(df$discharge_date - df$randdate)

# (i) Primary outcome: Mortality at day 28
df <- df %>% # corresponds to publication: only 1 death (in tofa arm)
  mutate(mort_28 = case_when(death_d <29 ~ 1, 
                             TRUE ~ 0)) 

# (ii) Mortality at day 60
df$mort_60 <- df$mort_28 # max fup time in TOFACOV was 28 days; thus mort_60 imputed from mort_28


# (iii) Time to death within max. follow-up time
df$death_reached <- df$mort_28
df <- df %>% # no missing and those that were discharged and afterwards have correct time to event data
  mutate(death_time = case_when(death_d >=0 ~ c(death_d), # time to death, if no time to death, then...
                                TRUE ~ 28)) # time to death censoring data (there were no withdrawals/LTFU)


# (iv) New mechanical ventilation among survivors within 28 days. TOFACOV only included clinstatus 2 and 3.
df <- df %>% # the 1 that died is NA => correct
  mutate(new_mv_28 = case_when(mort_28 == 0 & (clinstatus_2 == 5 | clinstatus_7 == 5 | clinstatus_14 == 5) ~ 1,
                               mort_28 == 0 ~ 0))

# (iv) Sens-analysis: Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
df <- df %>% # no missing anymore => correct
  mutate(new_mvd_28 = case_when(new_mv_28 == 1 | mort_28 == 1 ~ 1,
                                new_mv_28 == 0 | mort_28 == 0 ~ 0))


# (v) Clinical status at day 28
df <- df %>% # They called and followed all patients at day 28 and recorded no new serious adverse events, re-admissions or deaths besides the one that died on day 26
  mutate(clinstatus_28 = case_when(mort_28 == 1 ~ 6,
                                   discharge_d <29 ~ 1, # discharged alive / reached discharge criteria within 28d
                                   clinstatus_14 == 3 ~ 3, # these ones are still hosp at day 28
                                   clinstatus_14 == 4 ~ 4,
                                   clinstatus_14 == 5 ~ 5)) 
df$clinstatus_28 <- factor(df$clinstatus_28, levels = 1:6) # no missing
df$clinstatus_28_imp <- df$clinstatus_28 # equal to clinstatus_28_imp

# (vi) Time to discharge or reaching discharge criteria up to day 28 // Patients who died prior to day 28 are assumed not having reached discharge, i.e. counted as 28 days (as someone who has been censored on day 28). 
df <- df %>%
  mutate(discharge_reached = case_when(discharge_d <29 ~ 1,
                                       TRUE ~ 0))
df <- df %>% 
  mutate(discharge_time = case_when(discharge_d >=0 ~ c(discharge_d), # time to discharge, if no time to discharge, then...
                                    death_d >=0 ~ death_d)) # time to discharge censoring time
df <- df %>% # add 28d for those that died
  mutate(discharge_time = case_when(mort_28 == 1 ~ 28,
                                    TRUE ~ discharge_time))
df <- df %>% # restrict to max fup time 28d
  mutate(discharge_time = case_when(discharge_time >28 ~ 28,
                                    TRUE ~ discharge_time))


# (vi) Sens-analysis: Alternative definition/analysis of outcome: time to sustained discharge within 28 days. There were no re-admissions within 28d
df$discharge_reached_sus <- df$discharge_reached
df$discharge_time_sus <- df$discharge_time


# (vii) Viral clearance up to day 5, day 10, and day 15 (Viral load value <LOQ and/or undectectable): Not available in TOFACOV
# (viii) Quality of life at day 28: Not available in TOFAVOC


# (ix) Participants with an adverse event grade 3 or 4, or a serious adverse event, excluding death, by day 28
# table(df$ae_28, df$trt, useNA = "always")
# table(df$ae_28, df$mort_28, useNA = "always") # corresponds to publication and excludes the death

# (ix) Sens-analysis: Alternative definition/analysis of outcome: incidence rate ratio (Poisson regression) -> AE per person by d28
df$ae_28_sev <- df$ae_28 # there were only 1 AE grade 3/4 per person

# (ix) Sens-analysis: Alternative definition/analysis of outcome: time to first (of these) adverse event, within 28 days, considering death as a competing risk (=> censor and set to 28 days)
# time to first ae not available

# (x) Adverse events of special interest within 28 days: a) thromboembolic events (venous thromboembolism, pulmonary embolism, arterial thrombosis), b) secondary infections (bacterial pneumonia including ventilator-associated pneumonia, meningitis and encephalitis, endocarditis and bacteremia, invasive fungal infection including pulmonary aspergillosis), c) Reactivation of chronic infection including tuberculosis, herpes simplex, cytomegalovirus, herpes zoster and hepatitis B, d) serious cardiovascular and cardiac events (including stroke and myocardial infarction), e) events related to signs of bone marrow suppression (anemia, lymphocytopenia, thrombocytopenia, pancytopenia), f) malignancy, g) gastrointestinal perforation (incl. gastrointestinal bleeding/diverticulitis), h) liver dysfunction/hepatotoxicity (grade 3 and 4)
df <- df %>% 
  mutate(aesi_28 = case_when(ae_28_spec == "Pulmonary embolism" ~ "thromboembolic events",
                             ae_28_spec == "ALT increase" ~ "liver dysfunction/hepatotoxicity (grade 3 and 4)")) # these were >5 ULN

# (xi) Adverse events, any grade and serious adverse event, excluding death, within 28 days, grouped by organ classes
df <- df %>% 
  mutate(ae_28_list = case_when(ae_28_spec == "Pulmonary embolism" ~ "thromboembolic events",
                             ae_28_spec == "ALT increase" ~ "liver dysfunction/hepatotoxicity (grade 3 and 4)",
                             ae_28_spec == "Pneumomediastinum" ~ "pneumomediastinum"))

```
Discussion points
1) Get time to first adverse event?
2) Outcome definition "new_mv_28" differs to original trial publication (included new non-invasive ventilations, i.e. clinstatus == 4)


# Multiple imputation using chained equation
```{r}

```
Discussion points
1)


# Define final dataset, set references, summarize missing data and variables
```{r echo=TRUE}
# keep the overall set
df_all <- df
# reduce the df set to our standardized set across all trials
df <- df %>% 
  select(id_pat, trt, sex, age, ethn, country, icu, sympdur, vacc, clinstatus_baseline,
         comed_dexa, comed_rdv, comed_toci, comed_ab, comed_acoa, comed_interferon, comed_other,
         comed_cat,
         comorb_lung, comorb_liver, comorb_cvd, comorb_aht, comorb_dm, comorb_obese, comorb_smoker, immunosupp,
         any_comorb, comorb_cat,
         crp, 
         # sero, vl_baseline, variant,
         mort_28, mort_60, death_reached, death_time,
         new_mv_28, new_mvd_28,
         clinstatus_28_imp,
         discharge_reached, discharge_time, discharge_reached_sus, discharge_time_sus,
         ae_28_sev, aesi_28, ae_28_list,
         # ae_reached, ae_time, 
         # vir_clear_5, vir_clear_10, vir_clear_15, 
         # qol_28
         )
## set references, re-level
str(df)
# df <- df %>% 
#   mutate(Treatment = relevel(Treatment, "no JAK inhibitor"))

# Create a bar plot to visualize missing values in each column
original_order <- colnames(df)
missing_plot <- df %>%
  summarise_all(~ mean(is.na(.))) %>%
  gather() %>%
  mutate(key = factor(key, levels = original_order)) %>%
  ggplot(aes(x = key, y = value)) +
  geom_bar(stat = "identity") +
  labs(x = "Columns", y = "Proportion of Missing Values", title = "Missing Data Visualization") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(0, 1)
print(missing_plot)
```
Discussion points
1) Missing variables:
  Baseline:
  - variant
  - viremia
  - sero
  - vl_baseline
  Outcomes:
  - vir_clear_5, vir_clear_10, vir_clear_15
  - ae_reached, ae_time
  - qol_28
2) Missing data:
- 1 NA in crp 
- NAs in new_mv_death, aesi_28, ae_28_list: Not part of denominator


# (i) Primary outcome: Mortality at day 28
```{r}
# adjusted for baseline patient characteristics (age, respiratory support at baseline (ordinal scale 1-3 vs 4-5), dexamethasone use at baseline (y/n), remdesivir use at baseline (y/n), anti-IL-6 use at baseline (y/n)).
table(df$mort_28, df$trt, useNA = "always")
mort.28 <- df %>% 
  glm(mort_28 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.28, exp = T, confint = T, model.info = T, model.fit = T, digits = 2)
```
Discussion points
1) respiratory support at baseline (ordinal scale 1-3 vs 4-5 OR leave it as it is)?


# (ii) Mortality at day 60
```{r}
table(df$mort_60, df$trt, useNA = "always")
mort.60 <- df %>% 
  glm(mort_60 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(mort.60, exp = T, confint = T, model.info = T, model.fit = T, digits = 2)
```
Discussion points
1) 


# (iii) Time to death within max. follow-up time
```{r}
table(df$death_reached, df$death_time, useNA = "always")
table(df$death_reached, df$mort_60, useNA = "always")

df %>%
  drop_na(death_time) %>%
  filter(death_reached == 1) %>%
  group_by(trt) %>%
  summarise(median = median(death_time),
            IQR = IQR(death_time),
            Q1 = quantile(death_time, probs = 0.25),
            Q3 = quantile(death_time, probs = 0.75)) 

# time to death
km.ttdeath.check <- with(df, Surv(death_time, death_reached))
head(km.ttdeath.check, 100)
km.ttdeath <- survfit(Surv(death_time, death_reached) ~ 1, data=df)
summary(km.ttdeath, times = c(5,10,20,28,30,40,60))
plot(km.ttdeath, xlab="Days", main = 'Kaplan Meyer Plot') #base graphics
autoplot(km.ttdeath)
km.ttdeath_trt <- survfit(Surv(death_time, death_reached) ~ trt, data=df)
# autoplot(km.ttdeath_trt)

ttdeath <- df %>% 
  coxph(Surv(death_time, death_reached) ~ trt 
        + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
        , data =.)

tab_model(ttdeath)
# summ(ttdeath, exp = T, confint = T, model.info = T, model.fit = T, digits = 2)
```
Discussion points
1) 


# (iv) New mechanical ventilation among survivors within 28 days
```{r}
table(df$new_mv_28, df$trt, useNA = "always")
new.mv.28 <- df %>% 
  glm(new_mv_28 ~ trt 
       + age + clinstatus_baseline 
       + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(new.mv.28, exp = T, confint = T, model.info = T, model.fit = T, digits = 2)

# (iv) Sens-analysis: Alternative definition/analysis: New mechanical ventilation OR death within 28 days => include all in denominator. 
table(df$new_mvd_28, df$trt, useNA = "always")
new.mvd.28 <- df %>% 
  glm(new_mvd_28 ~ trt 
      + age + clinstatus_baseline + comed_dexa + comed_rdv + comed_toci
      , family = "binomial", data=.)
summ(new.mvd.28, exp = T, confint = T, model.info = T, model.fit = T, digits = 2)

```
Discussion points
1) 
